### 2019-01-24 abstract.tex
ssr: Added tex and bib files -- initial versions

\vspace*{-0.05in}
\begin{center}
\textbf{Abstract}~~ \textcolor{red}{\textbf{(To be revised)}}
\end{center}

\smallskip

Several researchers have studied 
a class of Boolean functions, 
called \textbf{nested canalyzing functions} (NCFs),
which are used to model certain biological 
phenomena \cite{Kauffman-1969,Kauffman-etal-2003,Kauffman-etal-2004}.
We study the relationships between NCFs and symmetric 
Boolean functions. Our results ...

### 2019-01-24 intro.tex
ssr: Added tex and bib files -- initial versions

\section{Introduction} 
\label{sec:intro}

\subsection{Canalyzing and Nested Canalyzing Functions}
\label{sse:ncf_def}

\begin{definition}\label{def:canalyzing}
Let $X = $ $\{x_1, x_2, \ldots, x_n\}$ be a set of $n$  Boolean variables.
A Boolean function $f(x_1, x_2, \ldots, x_n)$ over $X$ is \textbf{canalyzing}
if there is a variable $x_i \in X$ and values $a$, $b$ $\in \{0,1\}$ such that 
$f(x_1, \ldots, x_{i-1}, a, x_{i+1}, \ldots, x_n)$ $~=~ b$, regardless of the values
assigned to the variables in $X - \{x_i\}$.
\end{definition}

\medskip

\noindent
\textbf{Example 1:}~ Consider the function 
$f(x_1, x_2, x_3) ~=~ x_1 \vee (x_2 \wedge x_3)$.
This function is canalyzing since if we set to $x_1 = 1$,
$f(1, x_2, x_3) ~=~ 1$, regardless of the values assigned to $x_2$ and $x_3$. \qed

\medskip

A class of Boolean functions, called \textbf{nested canalyzing functions} (NCFs),
was introduced in \cite{Kauffman-etal-2003} to model the behavior of
certain biological systems.
We follow the presentation in \cite{Layne-2011} in defining 
such a Boolean function. 
(For a Boolean value $b$,~ the complement is denoted by $\overline{b}$.)

\begin{definition}\label{def:nested_canalyzing}
Let $X = $ $\{x_1, x_2, \ldots, x_n\}$ denote a set of $n$  Boolean variables.
Let $\pi$ be a permutation of $\{1, 2, \ldots, n\}$.
A Boolean function $f(x_1, x_2, \ldots, x_n)$ over $X$ is \textbf{nested canalyzing}
in the variable order $x_{\pi(1)}, x_{\pi(2)}, \ldots, x_{\pi(n)}$ with
\textbf{canalyzing values} $a_1, a_2, \ldots, a_n$ and 
\textbf{canalyzed values} $b_1, b_2, \ldots, b_n$ 
if $f$ can be expressed in the following form:
\[
f(x_1, x_2, \ldots, x_n) ~=~ 
   \begin{cases}
       \:b_1 & \mathrm{if~~} x_{\pi(1)} ~=~ a_1 \\
       \:b_2 & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~}
            x_{\pi(2)} ~=~ a_2 \\
       \:\vdots & \vdots \\
       \:b_n & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots~
             x_{\pi(n-1)} ~\neq~ a_{n-1} \mathrm{~~and~~} x_{\pi(n)} ~=~ a_n \\
       \:\overline{b_n} & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots~
            x_{\pi(n)} ~\neq~ a_n \\
   \end{cases}
\]
\end{definition}

\medskip
%% \clearpage
For convenience, we will use a computational notation to represent NCFs.
For $1 \leq i \leq n$, line $i$ of our representation has the following form:

\medskip

\noindent
\hspace*{1.1in} $x_{\pi(i)}:~ a_i ~~\longrightarrow~~ b_i$

\medskip

\noindent 
We say that $x_{\pi(i)}$ is the \textbf{canalyzing variable} that is
\textbf{tested} in line $i$, 
with $a_i$ and $b_i$ denoting respectively the canalyzing and 
canalyzed values in line $i$ as before,~ $1 \leq i \leq n$.
We refer to each such line as a \textbf{rule}.
When none of the conditions ``$x_{\pi(i)} ~=~ a_i$" 
is satisfied, we have line $n+1$ with the ``Default" rule
for which the canalyzed value is~ $\overline{b_n}$:

\medskip

\noindent
\hspace*{1.1in} Default:~ $\overline{b_n}$

\medskip
\noindent
In the remainder of this paper, we will refer to the above specification
of an NCF as the \textbf{simplified representation} and assume
(without loss of generality) that each NCF is specified in this manner.
The simplified representation provides the following 
convenient computational view of an NCF.~
Lines defining an NCF are 
considered sequentially in a top-down manner.
The computation stops at the first line where the 
specified condition is satisfied, and the value of the function
is the canalyzed value on that line. 
We now present an example of an NCF using the two representations
mentioned above.

\medskip
\noindent
\textbf{Example 2:}~ Consider the function 
$f(x_1, x_2, x_3) ~=~ x_1 \vee (x_2 \wedge x_3)$.
This function is nested canalyzing using the identity permutation $\pi$ on $\{1,2,3\}$
with canalyzing values $1,0,0$ and canalyzed values $1, 0, 0$.
We first show how this function can be expressed using the syntax of
Definition~\ref{def:nested_canalyzing}.

\[
f(x_1, x_2, x_3) ~=~ 
   \begin{cases}
       \:1 & \mathrm{if~~} x_{1} ~=~ 1 \\
       \:0 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~}
            x_{2} ~=~ 0 \\
       \:0 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~}
            x_{2} ~\neq~ 0 \mathrm{~~and~~} x_{3} = 0 \\
       \:1 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~}
            x_{2} ~\neq~ 0 \mathrm{~~and~~} x_{3} ~\neq~ 0 \\
   \end{cases}
\]

\medskip
\noindent
A simplified representation of the same function is as follows.

\bigskip

\noindent
\begin{tabular}{ll}
\hspace*{1.1in} & $x_1:~$  $1 ~\longrightarrow~ 1$ \\ [1ex]
\hspace*{1.1in} & $x_2:~$  $0 ~\longrightarrow~ 0$ \\ [1ex]
\hspace*{1.1in} & $x_3:~$  $0 ~\longrightarrow~ 0$ \\ [1ex]
\hspace*{1.1in} & Default:~ $1$ \\
\end{tabular}

\noindent
\qed

\noindent
\underline{\textsf{Generalized NCFs:}}~
For some problems, we will also consider a more general form
of NCFs defined below.

\begin{definition}\label{def:generalized ncf}
A {\bf generalized NCF} is a function represented as either a constant
or an NCF representation of a subset (not necessarily proper) 
of the function's variables.
\end{definition}

\noindent
\textbf{Example 3:}~ For $b \in \{0,1\}$, the constant function which takes 
on the value $b$ for every input is a generalized NCF since 
it can be represented by the following rule:

\smallskip

\hspace*{1.1in} Default:~ $b$

\smallskip

\noindent
The following generalized NCF specification for 
a function $f(x_1, x_2, x_3, x_4)$ indicates that the function
depends only on variables $x_1$ and $x_3$.

\medskip

\noindent
\begin{tabular}{ll}
\hspace*{1.1in} & $x_1:~$  $0 ~\longrightarrow~ 1$ \\ [1ex]
\hspace*{1.1in} & $x_3:~$  $1 ~\longrightarrow~ 0$ \\ [1ex]
\hspace*{1.1in} & Default:~ $1$ \\
\end{tabular}

\noindent
\qed

\bigskip

\noindent
\textbf{Note:} If a generalized NCF is not a constant, 
there is an assignment to its variables
such that the value of the function is 0,
and there is an assignment to its variables
such that the value of the function is 1.

%\noindent

\subsection{Summary of Results}
\label{sse:contrib}

Our focus is on the relationship between NCFs and symmetric
Boolean functions.

\subsection{Related Work}

Many researchers have studied mathematical properties of NCFs and
pointed out the importance of NCFs in modeling biological
phenomena (e.g., \cite{Kauffman-etal-2003,Kauffman-etal-2004,Layne-2011,
Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}).

\medskip

\noindent
\textbf{Equivalence and Implication Problems.}~
We note that for general Boolean formulas, the implication and equivalence problems
are \cnp-hard \cite{GJ-1979}.

\bigskip

\noindent
\textbf{Sensitivity of Boolean Functions.}~
There is extensive literature on the sensitivity of various classes
of Boolean functions
(e.g., \cite{Buhrman-etal-2002,Nisan-etal-1994,Zhang-2011}).
For a discussion on how the sensitivity of a Boolean function provides
an indication of its stability, the reader is referred to \cite{Kauffman-etal-2004,
Layne-2011,Layne-etal-2012}.
Observations regarding relationships between sensitivities of Boolean functions
and their computational complexity are presented
in \cite{Buhrman-etal-2002,Nisan-etal-1994}.
Li and Adeyeye \cite{Li-etal-2012} present lower and upper bounds on the
sensitivity of any NCF.
Li et al. \cite{Li-etal-2011,Li-etal-2013} conjectured that the average
sensitivity of any NCF is strictly less than  $4/3$.
This conjecture was proved by Klotz et al. \cite{Klotz-etal-2013} by establishing
a tight upper bound on the average sensitivity of any NCF.
Their methods rely on Fourier analysis of Boolean functions
\cite{Odonnell-2014}.
In \cite{Stearns-etal-2018}, 
we presented a characterization of NCFs with maximum average sensitivity.
This characterization allowed us to present an alternative proof 
of the upper bound in \cite{Klotz-etal-2013} and derive a closed
form expression for the number of such NCFs.

### 2019-01-24 symmetry_properties.tex
ssr: Added tex and bib files -- initial versions

\section{Symmetry of Nested Canalyzing Functions}
\label{sse:ncf_and_symmetry}

A pair of variables of a Boolean function $f(x_1, x_2, \ldots, x_n)$ is
said to be \textbf{symmetric} if their values can be interchanged without
affecting the value of the function.
This form of interchange symmetry partitions the set of variables into a set of 
 \textbf{symmetry groups}, where the members of each group are pairwise symmetric.
Function $f$ is said to be
$r$-\textbf{symmetric} if it has at most $r$ symmetry groups.
In this case, the value of $f$
depends only on how many of the variables in each symmetry group have the value 1.
We say that $f$ is \textbf{properly} $r$-\textbf{symmetric} if 
it is $r$-symmetric, but not $(r-1)$-symmetric.

In the literature, a 1-symmetric function $f$ is referred to simply 
as a \textbf{symmetric} function. 
Thus, if $f$ is a symmetric function, then for any
input $(a_1, a_2, \ldots, a_n)$ to $f$, where $a_i \in \{0,1\}$ for 
$1 \leq i \leq n$, and any permutation $\pi$ of $\{1, 2, \ldots, n\}$,
$f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.

\begin{theorem}\label{thm:ncf_symmetric}
The only symmetric canalyzing functions are OR, AND, NOR, NAND, 
the constant function 0, and the constant function 1.

The only symmetric NCFs are OR, AND, NOR, and NAND.
\end{theorem}
\noindent
\textbf{Proof:}~
Suppose symmetric function $f$ is a canalyzing function. Then there is a variable $x$,
and values $a$ and $b$ such that whenever $x = a$, function $f$ has value $b$.
Since $f$ is symmetric, $f$  has value $b$ whenever any of its variables have value $a$.
Thus, if at least one of its variables has value $a$, then $f$ has value $b$.
If $f$ has value $\bar{b}$ when none of its variables has value $a$,
the four possible combinations of values for $a$ and $b$ correspond to the four functions
OR, AND, NOR, and NAND.
If $f$ has value $b$ when none of its variables has value $a$, then $f$ is the constant function $b$.

Suppose that $f$ is a NCF.
Since constant functions are not NCFs, 
$f$ has value $\bar{b}$ when none of its variables has value $a$.
The four possible combinations of values for $a$ and $b$ correspond to the four functions
OR, AND, NOR, and NAND, each of which is a NCF.
\QED

\begin{theorem}\label{thm:ncf_symmetric_variables}
A pair of variables of a NCF are symmetric iff
in the normalized representation of the NCF
they occur in the same layer and have the same canalyzing value.
\end{theorem}
\noindent
\textbf{Proof:}~
Suppose that a pair of variables occur in the same layer of a canalyzing representation, 
and the lines containing these variables have the same canalyzing value.
Then in any input assignment, the values of these two variables can be interchanged,
and the evaluation of the lines in that layer will have the same effect.
Thus, the variables are symmetric.

Now, let $f$ be the normalized NCF representation for the function.
For purposes of notational simplicity, 
wlog we assume that the canalyzing variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.
Let  $x_j$ and $x_k$ be a pair of variables that occur in different layers of $f$,
or occur in the same layer, but with different canalyzing values.
Wlog, assume that $j < k$, so that the line for $x_j$ occurs above the line for $x_k$.

Suppose that line $i$ of $f$, $1 \leq i \leq n$, is

\noindent
$(x_i : a_i \rightarrow b_i)$. \\
Consider the following input assignment $\alpha = (c_1, c_2, \ldots,  c_n)$ 
to the variables of $f$.

\noindent
For $1 \leq i < j$, set $c_i = \overline{a_i}$. \\
For $i = j$, set $c_j =a_j$. \\
For $i = k$, set $c_k =\overline{a_j}$. \\
For $j < i < k$, and for  $k < i \leq n$, 
if $b_i = b_j$, then set $c_i =  \overline{a_i}$,
else set $c_i =  a_i$.

Note that $f(\alpha) = b_j$,
since the variables in all lines above line $j$ have the complement of their canalyzing value,
and variable $x_j$ has its canalyzing value.

Now let $\alpha'$ be the assignment that is obtained from $\alpha$ by interchanging the values
of variables $x_j$ and $x_k$,
so that after the interchange variable $x_j$ has value $\overline{a_j}$
and variable $x_k$ has value $a_j$.
We will show that 
$f(\alpha') = \overline{b_j}$,
so variables $x_j$ and $x_k$ are not symmetric.

In line $k$ of $f$, canalyzing value $a_k$ is either the same or the complement of canalyzing value $a_j$,
and canalyzed output $b_k$ is either the same or the complement of canalyzed output $b_j$,
Thus, there are four possible cases for the form of line $k$.
We now consider each of the four cases.

{\bf Case 1:} $(x_k : a_j \rightarrow b_j)$. \\
Since line $k$ has the same canalyzing value and canalyzed output as line $j$,
line $k$ must occur in a lower layer than that of line $j$.
Thus, there is at least one line between line $j$ and line $k$ 
for which the canalyzed output is $\overline{b_j}$.
Let line $q$ be the first such line.
Note that for all $i$ such that $1 \leq i < q$, 
the value of variable $x_i$ in assignment $\alpha'$ does not match the canalyzing value of line $i$,
but the value of variable $x_q$ does match the canalyzing value of line $q$.
Since canalyzed output $b_q$ equals $\overline{b_j}$,
we have that $f(\alpha') = \overline{b_j}$.

{\bf Case 2:} $(x_k : a_j \rightarrow \overline{b_j})$. \\
Let line $q$ be the first line such that $j < q \leq k$ and
the canalyzed output of line $q$ is $\overline{b_j}$.
Note that $q$ might possibly equal $k$.
For all $i$ such that $1 \leq i < q$, 
the value of variable $x_i$ in assignment $\alpha'$ does not match the canalyzing value of line $i$,
but the value of variable $x_q$ does match the canalyzing value of line $q$.
Since canalyzed output $b_q$ equals $\overline{b_j}$,
we have that $f(\alpha') = \overline{b_j}$.

{\bf Case 3:} $(x_k : \overline{a_j} \rightarrow b_j)$. \\
Suppose there is a line below line $j$ for which the canalyzed output is $\overline{b_j}$.
Let $q$ be the first such line.
Note that for all $i$ such that $1 \leq i < q$, 
the value of variable $x_i$ in assignment $\alpha'$ does not match the canalyzing value of line $i$,
but the value of variable $x_q$ does match the canalyzing value of line $q$.
Since canalyzed output $b_q$ equals $\overline{b_j}$,
we have that $f(\alpha') = \overline{b_j}$.

Now suppose there is no line below line $j$ for which the canalyzed output is $\overline{b_j}$.
Then lines $j$ and $k$ both occur in the last layer of $f$.
Thus, for every line $i$, $1 \leq i \leq n$, 
the value of variable $x_i$ in assignment $\alpha'$ does not match the canalyzing value of line $i$.
Consequently, $f(\alpha') = \overline{b_j}$, the complement of the canalyzed output in the last layer of $f$.

{\bf Case 4:} $(x_k : \overline{a_j} \rightarrow \overline{b_j})$. \\
Suppose there is a line, other than line $k$, below line $j$ for which the canalyzed output is $\overline{b_j}$.
Let $q$ be the first such line.
Note that for all $i$ such that $1 \leq i < q$, 
the value of variable $x_i$ in assignment $\alpha'$ does not match the canalyzing value of line $i$,
but the value of variable $x_q$ does match the canalyzing value of line $q$.
Since canalyzed output $b_q$ equals $\overline{b_j}$,
we have that $f(\alpha') = \overline{b_j}$.

Now suppose that line $k$ is the only line below line $j$ for which the canalyzed output is $\overline{b_j}$.
Since line $j$ has canalyzed output $b_j$, line $k$ is the only line in its layer.
Since $f$ is normalized, the last layer of $f$ contains at least two lines.
Thus, there is a last layer following the layer containing line $k$.
The canalyzed output of all the lines in this last layer is $b_j$.
Thus, for every line $i$, $1 \leq i \leq n$, 
the value of variable $x_i$ in assignment $\alpha'$ does not match the canalyzing value of line $i$.
Consequently, $f(\alpha') = \overline{b_j}$, the complement of the canalyzed output in the last layer of $f$.
\QED

\begin{theorem}\label{thm:ncf_r_symmetric}
Suppose the normalized representation of a given NCF contains 
$r_1$ layers with only one distinct canalyzing value,
and $r_2$ layers with two distinct canalyzing values.
Then the function is properly $(r_1 + 2 r_2)$-symmetric.
\end{theorem}
\noindent
\textbf{Proof:}~
From Theorem \ref{thm:ncf_symmetric_variables}, 
any pair of variables occurring in the same layer, with the same canalyzing value,
are symmetric.
Thus the function is at most $(r_1 + 2 r_2)$-symmetric.
Moreover, any pair of variables from different layers, or with different canalyzing values
are not symmetric, so there are at least $(r_1 + 2 r_2)$ symmetry groups.
\QED

\begin{corollary}\label{cor:ncf_not_rsymm}
For every $n \geq 2$, there is an $n$-variable NCF that is not $n-1$ symmetric.
\end{corollary}

\begin{corollary}\label{cor:ncf_r_symmetric_layers}
A NCF with a normalized nested canalyzing representation consisting of $q$ layers is $2q$-symmetric,
and is not $(q-1)$-symmetric.

A $r$-symmetric NCF has a normalized nested canalyzing representation with at most $r-1$ layers,
and has an unnormalized nested canalyzing representation with at most $r$ layers.
\end{corollary}

Generalized forms of symmetry have been considered \cite{KS-2000}
in which a Boolean function is symmetric with respect to some permutation,
which can be more general than the permutations corresponding to symmetry groups.
We say that a Boolean function $f(x_1, x_2, \ldots, x_n)$ 
is \textbf{strongly asymmetric}
if for any permutation $\pi$ of $\{1, 2, \ldots, n\}$
\emph{except} the identity permutation,
there exists an input $(a_1, a_2, \ldots,  a_n)$ 
to $f$ such that $f(a_1, a_2, \ldots, a_n)$ $\neq$  
$f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
In general, there are functions with $n$ variables that are properly $n$-symmetric,
but not strongly asymmetric \cite{KS-2000}.

We now show that any $n$ variable NCF that is properly $n$-symmetric
is also strongly asymmetric.
As a consequence, there are NCFs with $n$ variables
which are properly $n$-symmetric.

\begin{theorem}\label{thm:ncf_strong_asymmetry}
A NCF with $n$ variables is strongly asymmetric iff
it is properly $n$-symmetric.
\end{theorem}

\noindent
\textbf{Proof:}~ 
If a NCF is $(n-1)$-symmetric, then consider a permutation than interchanges two variables 
from a symmetry group with at least two members.
The value of the function is invariant under any such permutation,
so the function is not strongly asymmetric.

Now consider a $n$ variable NCF that is properly $n$-symmetric.
Let $f$ be the normalized NCF representation for the function.
For purposes of notational simplicity, 
wlog we assume that the canalyzing variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.
Let $\pi$ be any permutation of $\{1, 2, \ldots, n\}$
\emph{except} the identity permutation.
We will construct an assignment $(c_1, c_2, \ldots,  c_n)$ 
to the variables of $f$ such that $f(c_1, c_2, \ldots,  c_n)$ $\neq$  
$f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)})$.

Let $k$ be the smallest index such that $k \neq \pi(k)$.
Since $\pi$ is a permutation, $k < n$.
Overall, for $1 \leq i < k$, $\pi(i) = i$;
 $\pi(k) > k$;
 and for $k < i \leq n$, $\pi(i) \geq k$.

Suppose that line $i$ of $f$, $1 \leq i \leq n$, is 
\noindent
$(x_i : a_i \rightarrow b_i)$. \\
Assignment $(c_1, c_2, \ldots,  c_n)$ is constructed as follows.

\noindent
For $1 \leq i < k$, set $c_i = \overline{a_i}$. \\
For $i = k$, set $c_k =a_k$. \\
For $i > k$, let $i' = \pi^{-1}(i)$, so that $\pi(i') = i$.
Thus, after the permutation of values, variable $x_{i'}$ will have value $c_i$.
If $b_{i'} = b_k$, then set $c_i =  \overline{a_{i'}}$,
else set $c_i =  a_{i'}$.

Since $c_k$ matches the canalyzing value of line $k$ of $f$,
and $c_i$ does not match the canalyzing value of any other earlier line $i$, $1 \leq i < k$, 
we have that $f(c_1, c_2, \ldots,  c_n) = b_k$.

Now consider $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)})$.
We first note that for $1 \leq i < k$, 
$c_{\pi(i)}$ does not match the canalyzing value of line $i$.

Let $k' = \pi^{-1}(k)$, so that $\pi(k') = k$.
Canalyzing value $a_{k'}$ is either the same or the complement of $a_k$,
and canalyzed output $b_{k'}$ is either the same or the complement of $b_k$,
Thus, there are four possible cases for the form of line $k'$ of $f$.
We will show that in each of the four cases,
$f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.

{\bf Case 1:} $(x_{k'} : a_k \rightarrow b_k)$. \\
Since line $k'$ has the same canalyzing value and canalyzed output as line $k$,
and $f$ is properly $n$-symmetric,
line $k'$ must occur in a lower layer than that of line $k$.
Thus, there is at least one line between line $k$ and line $k'$ for which the canalyzed output is $\overline{b_k}$.
Let line $q$ be the first such line.
Note that for all $i$ such that $1 \leq i < q$, 
$c_{\pi(i)}$ does not match the canalyzing value of line $i$,
but $c_{\pi(q)}$ does match the canalyzing value of line $q$.
Since canalyzed output $b_q$ equals $\overline{b_k}$,
we have that $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.

{\bf Case 2:} $(x_{k'} : a_k \rightarrow \overline{b_k})$. \\
Let line $q$ be the first line such that $k < q \leq k'$ and
the canalyzed output of line $q$ is $\overline{b_k}$.
Note that for all $i$ such that $1 \leq i < q$, 
$c_{\pi(i)}$ does not match the canalyzing value of line $i$,
so $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.

{\bf Case 3:} $(x_{k'} : \overline{a_k} \rightarrow b_k)$. \\
Suppose there is a line below line $k$ for which the canalyzed output is $\overline{b_k}$.
Let $q$ be the first such line.
Note that for all $i$ such that $1 \leq i < q$, 
$c_{\pi(i)}$ does not match the canalyzing value of line $i$,
but $c_{\pi(q)}$ does match the canalyzing value of line $q$.
Thus, $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.

Now suppose there is no line below line $k$ for which the canalyzed output is $\overline{b_k}$.
Then line $k$ occurs in the last layer of $f$.
Since $f$ is both normalized and properly $n$-symmetric,
this last layer contains exactly two lines.
Thus $k = n-1$ and $k' = n$.
Note that for all $i$, 
$c_{\pi(i)}$ does not match the canalyzing value of line $i$.
Thus, $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.

{\bf Case 4:} $(x_{k'} : \overline{a_k} \rightarrow \overline{b_k})$. \\
Suppose there is a line, other than line $k'$, below line $k$ for which the canalyzed output is $\overline{b_k}$.
Let $q$ be the first such line.
Note that for all $i$ such that $1 \leq i < q$, 
$c_{\pi(i)}$ does not match the canalyzing value of line $i$,
but $c_{\pi(q)}$ does match the canalyzing value of line $q$.
Thus, $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.

Now suppose that line $k'$ is the only line below line $k$ for which the canalyzed output is $\overline{b_k}$.
Since line $k$ has canalyzed output $b_k$, line $k'$ is the only line in its layer.
Since $f$ is normalized, there is a last layer following the layer containing line $k'$.
Thus, for all $i$, 
$c_{\pi(i)}$ does not match the canalyzing value of line $i$,
so $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.
\QED

Examples of properly $n$-symmetric $n$-variable NCFs
satisfying the conditions of Theorem \ref{thm:ncf_r_symmetric}
are the following. For each $n > 1$, let $f_n$ be the function defined by the following formula.

$$x_1 \vee \bar{x}_2 ( x_3 \vee \bar{x}_4 (\cdots  ) )$$

For instance,
$$f_6 = x_1 \vee \bar{x}_2 ( x_3 \vee \bar{x}_4 ( x_5 \vee \bar{x}_6  ) ),$$
and
$$f_7 = x_1 \vee \bar{x}_2 ( x_3 \vee \bar{x}_4 ( x_5 \vee \bar{x}_6   x_7) ).$$

Function $f_n(x_1, x_2, \ldots, x_n)$ is the NCF corresponding to the following NCF representation:

\medskip
\noindent
\hspace*{0.5in}
$x_1:~ (1 \rightarrow 1)$ \\
\hspace*{0.5in}
$x_2:~ (1 \rightarrow 0)$ \\
\hspace*{0.5in}
$x_3:~ (1 \rightarrow 1)$ \\
\hspace*{0.75in}
$\vdots$

\noindent
If $n$ is odd, the last line is

\noindent
\hspace*{0.5in}
$x_n:~ (1 \rightarrow 1)$

\noindent
and if $n$ is even, the last line is

\noindent
\hspace*{0.5in}
$x_n:~ (1 \rightarrow 0)$. \\

Note that if $x_i = 0$ for all $i$, $1 \leq i \leq n$, then the
value of $f_n$ is 0 if $n$ is odd and 1 if $n$ is even.
Also, note that the above representation is unnormalized. To normalize the representation,
the last line would be changed to have canalyzing value 0, and the same canalyzed output as the preceding line.

\iffalse
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%\begin{theorem}\label{thm:ncf_asymmetry}
\begin{theorem}\label{thm:ncf_asymmetry}.
Consider the set of variables $x_i$ for $1 \leq i \leq n$ and 
let $f(x_1, x_2, \ldots, x_n)$ be the NCF defined by the following sequence
of lines:

\medskip
\noindent
\hspace*{0.5in}
$x_1:~ (1 \rightarrow 1)$ \\
\hspace*{0.5in}
$x_2:~ (1 \rightarrow 0)$ \\
\hspace*{0.5in}
$x_3:~ (1 \rightarrow 1)$ \\
\hspace*{0.75in}
$\vdots$

\noindent
If $n$ is odd, the last line is

\noindent
\hspace*{0.5in}
$x_n:~ (1 \rightarrow 1)$

\noindent
and if $n$ is even, the last line is

\noindent
\hspace*{0.5in}
$x_n:~ (1 \rightarrow 0)$. \\

\noindent
Let $\pi$ be any permutation of $\{1, 2, \ldots, n\}$
\emph{except} the identity permutation.
Then there exists an input $(a_1, a_2, \ldots,  a_n)$ 
to $f$ such that $f(a_1, a_2, \ldots, ,a_n)$ $\neq$  
$f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
\end{theorem}

\noindent
\textbf{Proof:}~ 
Note that if $x_i = 0$ for all $i$, $1 \leq i \leq n$, then the
value of the function $f$ is 0 if $n$ is odd and 1 if $n$ is even.

First assume there is an odd $i$ such that $\pi(i)$ is even. 
Consider the input where $a_i$ is 1 and all others are 0. 
It can be verified that $f(a_1, a_2, \ldots, a_n)$ = 1 while
$f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$ = 0.

\smallskip
Now assume the permutation $\pi$ is such that for each odd $i$, $\pi(i)$ is odd 
and for each even $i$, $\pi(i)$ is even.
Let $k$ be the smallest index such that $k \neq \pi(k)$.
Assume that $k$  is odd. 
Construct the input $(a_1, a_2, \ldots,  a_n)$ to $f$ as follows: 
\begin{enumerate}
\item Set $a_k$ to 1 and $a_i$ to 0 for all other odd $i$. 
\item For even $i$, set $a_i$ to 0 if $i < k$ and to 1 otherwise. 
\end{enumerate}
We have the following claim.

\smallskip
\noindent
\textbf{Claim:}~ For the input $(a_1, a_2, \ldots, a_n)$, 
$f(a_1, a_2, \ldots, a_n)$ = 1 and
$f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$ = 0.

\smallskip
\noindent
\textbf{Proof of claim:}~ Note that the permutation
on the even indices maps 1's to 1's and leaves 0's in place.  
Before the permutation,  
$x_j = 0$ for $1 \leq j < k$ and $x_k = 1$,
so the canalyzing variable is $x_k$ and the function value is 1.
After the permutation, 
$x_j = a_j = 0$ for $1 \leq j < k$,
$x_k = 0$, and $x_{k+1} = 1$;
so the canalyzing variable is $x_{k+1}$ and the function value is 0.
This establishes the claim.

A similar proof can be given when $k$ is even and this completes
the proof of Theorem~\ref{thm:ncf_asymmetry}.
\QED

\noindent
The following result is a consequence of the above theorem.

\begin{corollary}\label{cor:ncf_not_rsymm}
For every $n \geq 2$, there is an $n$-variable NCF that is not $n-1$ symmetric.
\end{corollary}

\noindent
\textbf{Proof:}~ 
Let $f$ be the Boolean function defined in Theorem~\ref{thm:ncf_asymmetry}.
We will show by contradiction that $f$ is not $r$-symmetric for any $r < n$.

Suppose $f$ is $r$-symmetric for some $r < n$.
Then, by definition, there is a partition of the set $X$ =
$\{x_1, x_2, \ldots, x_n\}$ of variables into at most $r$ subsets such that
the value of $f$ depends only on the number of variables with value 1
in each subset.
Since $r < n$, this partition has a subset, say $X_j$, 
such that $|X_j| \geq 2$.
Let $X_j = \{x_{j_1}, x_{j_2}, \ldots, x_{j_k}\}$, where $k \geq 2$.
Assume without loss of generality that $j_1 < j_2 < \ldots < j_k$.
Define a permutation $\pi$ of $\{1, 2, \ldots, n\}$ as follows. 
\begin{enumerate}
\item For all $i$ such that $x_i \not\in X_j$, $\pi(i) = i$.
\item $\pi(j_{\ell})$ = $j_{\ell+1}$ for $1 \leq \ell_t \leq k-1$
      and $\pi(j_{k})$ = $j_1$. 
\end{enumerate}
Note that $\pi$ is not the identity permutation.
Thus, by Theorem~\ref{thm:ncf_asymmetry},
there is an input $(a_1, a_2, \ldots, a_n)$ to $f$ such that
such that
\begin{equation}\label{eqn:f_asymmetry}
f(a_1, a_2, \ldots, a_n) ~\neq~  f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})
\end{equation}
However, from the definition of permutation $\pi$, it follows
that the number of 1's in any subset of the partition 
is the same in both the inputs
$(a_1, a_2, \ldots, a_n)$ and
$(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
Therefore, since $f$ is $r$-symmetric, we must have
\begin{equation}\label{eqn:f_contra}
f(a_1, a_2, \ldots, a_n) ~=~ f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)}).
\end{equation}
Equation~(\ref{eqn:f_contra}) contradicts 
Equation~(\ref{eqn:f_asymmetry}), and the result follows. \QED
\fi

\iffalse
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
$$x_1 \vee \bar{x}_2 ( x_3 \vee \bar{x}_4 (\cdots  ) )$$

For instance,
$$f_6 = x_1 \vee \bar{x}_2 ( x_3 \vee \bar{x}_4 ( x_5 \vee \bar{x}_6  ) ),$$
and
$$f_7 = x_1 \vee \bar{x}_2 ( x_3 \vee \bar{x}_4 ( x_5 \vee \bar{x}_6   x_7) ).$$

Function $f_n$ is a NCF, 
and can be representing as a sequence where the variables appear in lexicographic order,
each $a_i$ is 1, and the $b_i$'s alternate between 1 and 0.
For instance, $f_6$ can be represented as follows:

\noindent
$(x_1: 1 \rightarrow 1 )$ \\
$(x_2: 1 \rightarrow 0 )$ \\
$(x_3: 1 \rightarrow 1 )$ \\
$(x_4: 1 \rightarrow 0 )$ \\
$(x_5: 1 \rightarrow 1 )$ \\
$(x_6: 1 \rightarrow 0 )$ \\

We now argue that $f_n$ is not $n-1$ symmetric.
Consider the classes of $f_n$, represented as an $r$-symmetric function for some $r$.
If we compare the assignment where a given odd-numbered variable $x_{2i+1}$ is the only variable equal to 1,
and the assignment where a given even-numbered variable $x_{2j}$ is the only variable equal to 1,
we can see that these assignments produce different values for $f_n$,
so variables $x_{2i+1}$ and $x_{2j}$ must belong to different classes.
If we compare the assignment where an odd-numbered variable $x_{2i+1}$  
and the even-numbered variable $x_{2i+2}$ 
are the only variables equal to 1,
and the assignment where an odd-numbered variable of the form $x_{2k+1}$, $k >i$,
and $x_{2i+2}$  are the only variables equal to 1,
we can see that these assignments produce different values for $f_n$,
so $x_{2i+1}$ and $x_{2k+1}$ must belong to different classes.
Thus, each odd-numbered variable is the only member of its class.
If we compare the assignment where an even-numbered variable $x_{2i}$  
and the odd-numbered variable $x_{2i+1}$ 
are the only variables equal to 1,
and the assignment where an even-numbered variable of the form $x_{2k}$, $k >i$,
and $x_{2i+1}$  are the only variables equal to 1,
we can see that these assignments produce different values for $f_n$,
so $x_{2i}$ and $x_{2k}$ must belong to different classes.
Thus, each even-numbered variable is the only member of its class.
\QED

%%The following result shows another form of asymmetry for the function
%%defined in the above theorem.  (This form of asymmetry may be stronger
%%than the one established in the above theorem.)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fi

### 2019-01-25 preliminaries.tex
ssr

\section{Other Definitions and Preliminary Results}
\label{sec:prelim}

\subsection{Overview}

Here, we present some definitions and preliminary results which enable
us to define a normalized representation for NCFs.
This representation plays an important role in several
problems considered in later sections.

\medskip

Recall that any NCF $f$ with $n$ variables,
denoted by $x_1$, $x_2$, $\ldots$, $x_n$,~
is specified using a variable ordering 
$x_{\pi(1)}$, $x_{\pi(2)}$, $\ldots$,  $x_{\pi(n)}$,
where $\pi$ is a permutation of $\{1, 2, \ldots, n\}$.
We will refer to this as the \textbf{lexicographic} ordering
of the variables in the specification.

\medskip

Throughout this section, we assume that an NCF $f$ with $n$
variables is specified using the simplified representation with $n$ lines. 
Unless otherwise mentioned, the ``Default" line (which is assigned the number $n+1$)
in the simplified representation of an NCF is \emph{not} considered
in the results stated in this section.

\subsection{Layers and Normalized Representation}
\label{sse:ncf_layer}

We begin with a two-part observation from \cite{Stearns-etal-2018}.

\begin{observation} \label{obs:ncf_transformations}
Let $f$ be an NCF with $n$ variables specified using $n$ lines
in the simplified representation. 
\begin{description}
\item{(a)} 
For any $q \geq 2$ and for any $i$, $1 \leq i \leq n-q+1$,
if the $q$ consecutive lines $i$, $i+1$, $\ldots$, $i+q-1$~
have the same canalyzed value, then the function remains
unchanged if these $q$ lines are permuted in any order
without changing the other lines.

\item{(b)} 
Suppose lines $n-1$ and $n$ in the specification of $f$ 
have complementary canalyzed outputs.
Then, the function remains unchanged 
if the canalyzing value and canalyzed output in line $n$
are both complemented. 
(Here, the value on the ``Default" line is also complemented.)
\QED
\end{description}
\end{observation}

\noindent
Li et al. \cite{Li-etal-2013} have defined the concept of a 
{\bf layer} of an NCF in terms of
an algebraic representation of the NCF as an extended monomial.
For our purposes, we use the following definition based on the
simplified representation of NCFs given in Section~\ref{sec:intro}.

\begin{definition}
\label{def:layer}
A {\bf layer} of a NCF representation is a maximal length sequence 
of lines with the same canalyzed output.
\end{definition}
%%Note that the ``Default" line in the representation of an NCF is
%%\emph{not} considered in the definition of a layer.

\noindent
\textbf{Example 4:}~ Consider the following function $f$ of six variables
$x_1$, $x_2$, $\ldots$, $x_6$.

\bigskip

\noindent
\begin{tabular}{ll}
\hspace*{1.1in} & $x_1:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
\hspace*{1.1in} & $x_2:~$  $0 ~\longrightarrow~ 1$ \\ [0.5ex]
\hspace*{1.1in} & $x_3:~$  $0 ~\longrightarrow~ 0$ \\ [0.5ex]
\hspace*{1.1in} & $x_4:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
\hspace*{1.1in} & $x_5:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
\hspace*{1.1in} & $x_6:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
\hspace*{1.1in} & Default:~ $1$ \\
\end{tabular}

\medskip
\noindent
This function has 4 layers, with 
Layer~1 consisting of the lines with $x_1$ and $x_2$, 
Layer~2 consisting of the lines with $x_3$ and $x_4$, 
Layer~3 consisting of the line with $x_5$ and 
Layer~4 consisting of the line with $x_6$. \qed

\medskip
From Part~(a) of Observation~\ref{obs:ncf_transformations},
it follows that the lines within the same layer of an NCF
representation can be permuted without changing the function.
Part~(b) of Observation~\ref{obs:ncf_transformations} points
out that for any NCF $f$ with $n \geq 2$ variables, there is a simplified 
representation in which lines $n-1$ and $n$ are in the same layer.

%% It can be verified that the number of layers in the 
%% extended monomial formula for a NCF (as defined in \cite{Li-etal-2013})
%% is the same as the number of layers in the normalized representation.

\medskip
We can now define the notion of a \textbf{normalized} 
representation for NCFs.

\begin{definition}
\label{def:normalized}
Let $f$ be an NCF with $n \geq 2$ variables specified using
the simplified representation with 
lexicographic ordering given by  
$\langle x_{\pi(1)}, x_{\pi(2)}, x_{\pi(n)}\rangle$.
We say that the representation is {\bf normalized} if
both of the following conditions hold.
\begin{enumerate}
  \item Lines $n-1$ and $n$ have the same canalyzed output ~and
  \item each pair of consecutive lines
        with the same canalyzed output occur in the lexicographic order 
        of their canalyzing variables; that is, if two consecutive lines are \\
%\medskip
\noindent
\hspace*{1.1in} $x_i$:~ $a_i \longrightarrow b$ \\
\hspace*{1.1in} $x_j$:~ $a_j \longrightarrow b$

%\medskip
\noindent
then $x_i$ precedes $x_j$ in the lexicographic ordering.
\end{enumerate}
\end{definition}

\noindent
\textbf{Example 5:}~

\medskip
\noindent
\textbf{Part 1:}~
Consider the NCF $f$ shown in Example~4 with
the variable ordering $x_1$, $x_2$, $x_3$, $x_4$, $x_5$, $x_6$.
This representation is \emph{not} normalized since lines 5 and 6
have different canalyzed values.
If we change line 6 to ``$x_6$:~ 0 $\longrightarrow$ 1" and the
value on the ``Default" line to 0, then we obtain 
a normalized representation.

\medskip
\noindent
\textbf{Part 2:}~ Consider the normalized version of Example~4 
produced in Part~1 above. 
In that representation, if we exchange Lines 1 and 2, 
by Part~(a) of Observation~\ref{obs:ncf_transformations},
the function does not change. 
However, the resulting representation is \emph{not} normalized
since the line with $x_2$ precedes the one 
with $x_1$. \qed

\medskip
From Observation \ref{obs:ncf_transformations},
we note that in polynomial time, for any NCF representation with more than one line,
we can first modify the last line if necessary so that it has the same canalyzed output as the preceding line,
and then we can lexicographically sort
each maximal sequence of consecutive lines with the same canalyzed output,
thereby obtaining an equivalent normalized NCF representation.
We state this formally below.

\begin{observation}\label{obs:normalization_poly}
Given an NCF representation for a function $f$, a normalized
representation for $f$ can be obtained in polynomial time. \QED
\end{observation}

\subsection{Closure Properties of NCFs}

\begin{observation}\label{complement}
The set of NCFs and the set of generalized NCFs
are each closed under complement.
Moreover, given an NCF or a generalized NCF representation $f$, 
an NCF or generalized NCF $f'$ representing the complement of $f$ 
can be constructed in linear time.
\end{observation}

\noindent
\textbf{Proof:}~ We construct
representation $f'$ by complementing 
every canalyzed output in representation $f$.  \QED

\medskip
We also observe that the set of NCFs is not closed under OR,~ XOR~ or AND.
To see this for OR and XOR,  note that the functions 
$f_1 = x_1 \, \bar{x}_2$ and $f_2 = \bar{x}_1 \, x_2$
are each NCFs, but their OR and XOR is the same Boolean function,
and this Boolean function is not  a canalyzing function of either variable.
To see this for AND, note that the functions 
$f_1 = x_1 \vee \bar{x}_2$ and $f_2 = \bar{x}_1 \vee x_2$
are each NCFs, but their AND is not  a canalyzing function of either variable.
We now define a new operation (projection) on Boolean functions 
and show through an explicit construction that
NCFs are closed under this operation.
This construction is used in Section~\ref{sse:gen_ncf_implication}
where we present an efficient algorithm for the implication problem
for generalized NCFs.

\begin{definition}\label{def:projection}
Given a Boolean function $f$, a variable $x$, and a Boolean value $a$,
the {\bf projection} of $f$ on $x = a$,
denoted by $f_{x = a}$, is the function on the remaining variables
whose value on any assignment $\alpha$ to these variables
is the value of $f$ when $\alpha$ is extended to a complete assignment for $f$
by setting the variable $x$ to the value $a$.
\end{definition}

\noindent
We now show that generalized NCFs are closed under projection.

\begin{lemma}\label{lem:projection}
Given a generalized NCF representation $f$, a variable $x$, and a
Boolean value $a$, a generalized NCF representing the projection
$f_{x = a}$ can be constructed in linear time.
\end{lemma}

\noindent
\textbf{Proof:}~ We use a simple case analysis.
We will assume that the layers of the representation of $f$ have
been identified and that they are numbered 
from top to bottom, starting with 1.

\smallskip

\noindent
{\bf Case 1}:~ Variable $x$ does not occur in $f$.
Then  $f_{x = a}$ is the same as $f$.

\smallskip

\noindent
{\bf Case 2}:~ For some value $b$, $f$ contains the line~ 
``$x:~ a \longrightarrow b$".~
We have two subcases.
\begin{description}
\item {(i)} If this line occurs in the first layer of $f$, 
then $f_{x = a}$ is the constant $b$.

\item{(ii)} If this line occurs in a layer of $f$ other than the first layer, 
$f_{x = a}$ is obtained from $f$ by deleting 
the layer containing this line and all subsequent layers.
(Note that in this case, the preceding layer has canalyzed output $\bar{b}$.)
\end{description}

%%\smallskip

\noindent
{\bf Case 3}:~ For some value $b$, $f$ contains the line~
``$x:~ \bar{a} \longrightarrow b$".~
We have three subcases here.
\begin{description}
\item{(i)}
If this line occurs in a layer with more than one line, 
or if this line occurs in a layer that is not the last layer,
then $f_{x = a}$ is obtained from $f$ by deleting this line.
\item{(ii)}
If this line is the last line in $f$, 
and there is no other line in $f$ whose canalyzed output is $b$,
then $f_{x = a}$ is  the constant $\bar{b}$.
\item{(iii)}
If this line is the last line in $f$, is the only line is its layer, 
and there is at least one other line in $f$ whose canalyzed output is $b$,
then $f_{x = a}$ is obtained from $f$ by deleting the last two layers.
\end{description}
\QED

### 2019-01-25 intro.tex
ssr

@@
 
+\iffalse
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \noindent
@@
 such that the value of the function is 1.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\fi
+
+\subsection{Use of NCFs in Modeling Biological Systems}
+
+The term \textbf{canalization} was first used by
+Conrad Waddington (a British geneticist)
+in 1942 \cite{Waddington-1942}
+to express the ability of a species
+to produce members with very similar observable features over a period
+of time even when there are significant changes in the environment
+during that period.
+Since that time, the term ``canalization'' has been used to describe
+the stability of a biological system with changes
+in external conditions.
+In 1969, Kauffman \cite{Kauffman-1969} introduced a Boolean network model
+to explain the stability of gene regulatory networks.
+Kauffman observed that the use of canalyzing Boolean
+functions in the model captured many observed properties of
+gene regulatory networks.
+The subclass of NCFs was
+introduced later by Kauffman et al. \cite{Kauffman-etal-2003} 
+to facilitate a rigorous analysis of the Boolean network model
+for gene regulatory networks.
 

### 2019-01-25 intro.tex
ssr

@@
 
-\subsection{Use of NCFs in Modeling Biological Systems}
+%\noindent
+
+\subsection{Summary of Results}
+\label{sse:contrib}
+
+Our focus is on the relationship between NCFs and symmetric
+Boolean functions.
+
+\subsection{Related Work}
+\label{sse:related}
 
@@
 for gene regulatory networks.
-
-%\noindent
-
-\subsection{Summary of Results}
-\label{sse:contrib}
-
-Our focus is on the relationship between NCFs and symmetric
-Boolean functions.
-
-\subsection{Related Work}
-
 Many researchers have studied mathematical properties of NCFs and
 pointed out the importance of NCFs in modeling biological
-phenomena (e.g., \cite{Kauffman-etal-2003,Kauffman-etal-2004,Layne-2011,
+phenomena (e.g., \cite{Layne-2011,
 Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}).
 
-\medskip
-
-\noindent
-\textbf{Equivalence and Implication Problems.}~
-We note that for general Boolean formulas, the implication and equivalence problems
-are \cnp-hard \cite{GJ-1979}.
-
-\bigskip
-
-\noindent
-\textbf{Sensitivity of Boolean Functions.}~
 There is extensive literature on the sensitivity of various classes
@@
 Layne-2011,Layne-etal-2012}.
-Observations regarding relationships between sensitivities of Boolean functions
-and their computational complexity are presented
-in \cite{Buhrman-etal-2002,Nisan-etal-1994}.
 Li and Adeyeye \cite{Li-etal-2012} present lower and upper bounds on the

### 2019-01-25 preliminaries.tex
ssr

@@
 
-\subsection{Closure Properties of NCFs}
+\subsection{Defintions: Symmetric Boolean Functions}
+\label{sse:symmetry}
 
-\begin{observation}\label{complement}
-The set of NCFs and the set of generalized NCFs
-are each closed under complement.
-Moreover, given an NCF or a generalized NCF representation $f$, 
-an NCF or generalized NCF $f'$ representing the complement of $f$ 
-can be constructed in linear time.
-\end{observation}
+A pair of variables of a Boolean function $f(x_1, x_2, \ldots, x_n)$ is
+said to be \textbf{symmetric} if their values can be interchanged without
+affecting the value of the function.
+This form of interchange symmetry partitions the set of variables into a set of
+ \textbf{symmetry groups}, where the members of each group are pairwise symmetric.
+Function $f$ is said to be
+$r$-\textbf{symmetric} if it has at most $r$ symmetry groups.
+In this case, the value of $f$
+depends only on how many of the variables in each symmetry group have the value 1.
+We say that $f$ is \textbf{properly} $r$-\textbf{symmetric} if
+it is $r$-symmetric, but not $(r-1)$-symmetric.
 
-\noindent
-\textbf{Proof:}~ We construct
-representation $f'$ by complementing 
-every canalyzed output in representation $f$.  \QED
-
-\medskip
-We also observe that the set of NCFs is not closed under OR,~ XOR~ or AND.
-To see this for OR and XOR,  note that the functions 
-$f_1 = x_1 \, \bar{x}_2$ and $f_2 = \bar{x}_1 \, x_2$
-are each NCFs, but their OR and XOR is the same Boolean function,
-and this Boolean function is not  a canalyzing function of either variable.
-To see this for AND, note that the functions 
-$f_1 = x_1 \vee \bar{x}_2$ and $f_2 = \bar{x}_1 \vee x_2$
-are each NCFs, but their AND is not  a canalyzing function of either variable.
-We now define a new operation (projection) on Boolean functions 
-and show through an explicit construction that
-NCFs are closed under this operation.
-This construction is used in Section~\ref{sse:gen_ncf_implication}
-where we present an efficient algorithm for the implication problem
-for generalized NCFs.
-
-\begin{definition}\label{def:projection}
-Given a Boolean function $f$, a variable $x$, and a Boolean value $a$,
-the {\bf projection} of $f$ on $x = a$,
-denoted by $f_{x = a}$, is the function on the remaining variables
-whose value on any assignment $\alpha$ to these variables
-is the value of $f$ when $\alpha$ is extended to a complete assignment for $f$
-by setting the variable $x$ to the value $a$.
-\end{definition}
-
-\noindent
-We now show that generalized NCFs are closed under projection.
-
-\begin{lemma}\label{lem:projection}
-Given a generalized NCF representation $f$, a variable $x$, and a
-Boolean value $a$, a generalized NCF representing the projection
-$f_{x = a}$ can be constructed in linear time.
-\end{lemma}
-
-\noindent
-\textbf{Proof:}~ We use a simple case analysis.
-We will assume that the layers of the representation of $f$ have
-been identified and that they are numbered 
-from top to bottom, starting with 1.
-
-\smallskip
-
-\noindent
-{\bf Case 1}:~ Variable $x$ does not occur in $f$.
-Then  $f_{x = a}$ is the same as $f$.
-
-\smallskip
-
-\noindent
-{\bf Case 2}:~ For some value $b$, $f$ contains the line~ 
-``$x:~ a \longrightarrow b$".~
-We have two subcases.
-\begin{description}
-\item {(i)} If this line occurs in the first layer of $f$, 
-then $f_{x = a}$ is the constant $b$.
-
-\item{(ii)} If this line occurs in a layer of $f$ other than the first layer, 
-$f_{x = a}$ is obtained from $f$ by deleting 
-the layer containing this line and all subsequent layers.
-(Note that in this case, the preceding layer has canalyzed output $\bar{b}$.)
-\end{description}
-
-%%\smallskip
-
-\noindent
-{\bf Case 3}:~ For some value $b$, $f$ contains the line~
-``$x:~ \bar{a} \longrightarrow b$".~
-We have three subcases here.
-\begin{description}
-\item{(i)}
-If this line occurs in a layer with more than one line, 
-or if this line occurs in a layer that is not the last layer,
-then $f_{x = a}$ is obtained from $f$ by deleting this line.
-\item{(ii)}
-If this line is the last line in $f$, 
-and there is no other line in $f$ whose canalyzed output is $b$,
-then $f_{x = a}$ is  the constant $\bar{b}$.
-\item{(iii)}
-If this line is the last line in $f$, is the only line is its layer, 
-and there is at least one other line in $f$ whose canalyzed output is $b$,
-then $f_{x = a}$ is obtained from $f$ by deleting the last two layers.
-\end{description}
-\QED
+In the literature, a 1-symmetric function $f$ is referred to simply
+as a \textbf{symmetric} function.
+Thus, if $f$ is a symmetric function, then for any
+input $(a_1, a_2, \ldots, a_n)$ to $f$, where $a_i \in \{0,1\}$ for
+$1 \leq i \leq n$, and any permutation $\pi$ of $\{1, 2, \ldots, n\}$,
+$f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.

### 2019-01-25 symmetry_properties.tex
ssr

@@
 
-A pair of variables of a Boolean function $f(x_1, x_2, \ldots, x_n)$ is
-said to be \textbf{symmetric} if their values can be interchanged without
-affecting the value of the function.
-This form of interchange symmetry partitions the set of variables into a set of 
- \textbf{symmetry groups}, where the members of each group are pairwise symmetric.
-Function $f$ is said to be
-$r$-\textbf{symmetric} if it has at most $r$ symmetry groups.
-In this case, the value of $f$
-depends only on how many of the variables in each symmetry group have the value 1.
-We say that $f$ is \textbf{properly} $r$-\textbf{symmetric} if 
-it is $r$-symmetric, but not $(r-1)$-symmetric.
+\subsection{Overview}
+\label{sse:res_overview}
 
-In the literature, a 1-symmetric function $f$ is referred to simply 
-as a \textbf{symmetric} function. 
-Thus, if $f$ is a symmetric function, then for any
-input $(a_1, a_2, \ldots, a_n)$ to $f$, where $a_i \in \{0,1\}$ for 
-$1 \leq i \leq n$, and any permutation $\pi$ of $\{1, 2, \ldots, n\}$,
-$f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
+\subsection{Symmetric Canalyzing and Nested Canalyzing Functions}
+\label{sse:sym_and_cf_ncf}
 
@@
 \textbf{Proof:}~
-Suppose symmetric function $f$ is a canalyzing function. Then there is a variable $x$,
-and values $a$ and $b$ such that whenever $x = a$, function $f$ has value $b$.
-Since $f$ is symmetric, $f$  has value $b$ whenever any of its variables have value $a$.
-Thus, if at least one of its variables has value $a$, then $f$ has value $b$.
-If $f$ has value $\bar{b}$ when none of its variables has value $a$,
-the four possible combinations of values for $a$ and $b$ correspond to the four functions
-OR, AND, NOR, and NAND.
-If $f$ has value $b$ when none of its variables has value $a$, then $f$ is the constant function $b$.
+Suppose symmetric function $f$ is a canalyzing function. Then there
+is a variable $x$, and values $a$ and $b$ such that whenever $x =
+a$, function $f$ has value $b$.  Since $f$ is symmetric, $f$  has
+value $b$ whenever any of its variables have value $a$.  Thus, if
+at least one of its variables has value $a$, then $f$ has value
+$b$.  If $f$ has value $\bar{b}$ when none of its variables has
+value $a$, the four possible combinations of values for $a$ and $b$
+correspond to the four functions OR, AND, NOR, and NAND.  If $f$
+has value $b$ when none of its variables has value $a$, then $f$
+is the constant function $b$.
 
-Suppose that $f$ is a NCF.
-Since constant functions are not NCFs, 
+Suppose that $f$ is a NCF.  Since constant functions are not NCFs,
 $f$ has value $\bar{b}$ when none of its variables has value $a$.
-The four possible combinations of values for $a$ and $b$ correspond to the four functions
-OR, AND, NOR, and NAND, each of which is a NCF.
-\QED
+The four possible combinations of values for $a$ and $b$ correspond
+to the four functions OR, AND, NOR, and NAND, each of which is a
+NCF.  \QED
+
+\subsection{Symmetric Pairs of Variables in an NCF: A Characterization
+and Its Applictions}
+\label{sse:ncf_strong_sym}
 
@@
 \textbf{Proof:}~
-Suppose that a pair of variables occur in the same layer of a canalyzing representation, 
-and the lines containing these variables have the same canalyzing value.
-Then in any input assignment, the values of these two variables can be interchanged,
-and the evaluation of the lines in that layer will have the same effect.
-Thus, the variables are symmetric.
+Suppose that a pair of variables occur in the same layer of a
+canalyzing representation, and the lines containing these variables
+have the same canalyzing value.  Then in any input assignment, the
+values of these two variables can be interchanged, and the evaluation
+of the lines in that layer will have the same effect.  Thus, the
+variables are symmetric.
 
 Now, let $f$ be the normalized NCF representation for the function.
-For purposes of notational simplicity, 
-wlog we assume that the canalyzing variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.
-Let  $x_j$ and $x_k$ be a pair of variables that occur in different layers of $f$,
-or occur in the same layer, but with different canalyzing values.
-Wlog, assume that $j < k$, so that the line for $x_j$ occurs above the line for $x_k$.
+For purposes of notational simplicity, wlog we assume that the
+canalyzing variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.
+Let  $x_j$ and $x_k$ be a pair of variables that occur in different
+layers of $f$, or occur in the same layer, but with different
+canalyzing values.  Wlog, assume that $j < k$, so that the line for
+$x_j$ occurs above the line for $x_k$.
 
@@
 
-Note that $f(\alpha) = b_j$,
-since the variables in all lines above line $j$ have the complement of their canalyzing value,
-and variable $x_j$ has its canalyzing value.
+Note that $f(\alpha) = b_j$, since the variables in all lines above
+line $j$ have the complement of their canalyzing value, and variable
+$x_j$ has its canalyzing value.
 
-Now let $\alpha'$ be the assignment that is obtained from $\alpha$ by interchanging the values
-of variables $x_j$ and $x_k$,
-so that after the interchange variable $x_j$ has value $\overline{a_j}$
-and variable $x_k$ has value $a_j$.
-We will show that 
-$f(\alpha') = \overline{b_j}$,
-so variables $x_j$ and $x_k$ are not symmetric.
+Now let $\alpha'$ be the assignment that is obtained from $\alpha$
+by interchanging the values of variables $x_j$ and $x_k$, so that
+after the interchange variable $x_j$ has value $\overline{a_j}$ and
+variable $x_k$ has value $a_j$.  We will show that $f(\alpha') =
+\overline{b_j}$, so variables $x_j$ and $x_k$ are not symmetric.
 
-In line $k$ of $f$, canalyzing value $a_k$ is either the same or the complement of canalyzing value $a_j$,
-and canalyzed output $b_k$ is either the same or the complement of canalyzed output $b_j$,
-Thus, there are four possible cases for the form of line $k$.
-We now consider each of the four cases.
+In line $k$ of $f$, canalyzing value $a_k$ is either the same or
+the complement of canalyzing value $a_j$, and canalyzed output $b_k$
+is either the same or the complement of canalyzed output $b_j$,
+Thus, there are four possible cases for the form of line $k$.  We
+now consider each of the four cases.
 
-{\bf Case 1:} $(x_k : a_j \rightarrow b_j)$. \\
-Since line $k$ has the same canalyzing value and canalyzed output as line $j$,
-line $k$ must occur in a lower layer than that of line $j$.
-Thus, there is at least one line between line $j$ and line $k$ 
-for which the canalyzed output is $\overline{b_j}$.
-Let line $q$ be the first such line.
-Note that for all $i$ such that $1 \leq i < q$, 
-the value of variable $x_i$ in assignment $\alpha'$ does not match the canalyzing value of line $i$,
-but the value of variable $x_q$ does match the canalyzing value of line $q$.
-Since canalyzed output $b_q$ equals $\overline{b_j}$,
+{\bf Case 1:} $(x_k : a_j \rightarrow b_j)$. \\ Since line $k$ has
+the same canalyzing value and canalyzed output as line $j$, line
+$k$ must occur in a lower layer than that of line $j$.  Thus, there
+is at least one line between line $j$ and line $k$ for which the
+canalyzed output is $\overline{b_j}$.  Let line $q$ be the first
+such line.  Note that for all $i$ such that $1 \leq i < q$, the
+value of variable $x_i$ in assignment $\alpha'$ does not match the
+canalyzing value of line $i$, but the value of variable $x_q$ does
+match the canalyzing value of line $q$.  Since canalyzed output
+$b_q$ equals $\overline{b_j}$, we have that $f(\alpha') =
+\overline{b_j}$.
+
+{\bf Case 2:} $(x_k : a_j \rightarrow \overline{b_j})$. \\ Let line
+$q$ be the first line such that $j < q \leq k$ and the canalyzed
+output of line $q$ is $\overline{b_j}$.  Note that $q$ might possibly
+equal $k$.  For all $i$ such that $1 \leq i < q$, the value of
+variable $x_i$ in assignment $\alpha'$ does not match the canalyzing
+value of line $i$, but the value of variable $x_q$ does match the
+canalyzing value of line $q$.  Since canalyzed output $b_q$ equals
+$\overline{b_j}$, we have that $f(\alpha') = \overline{b_j}$.
+
+{\bf Case 3:} $(x_k : \overline{a_j} \rightarrow b_j)$. \\ Suppose
+there is a line below line $j$ for which the canalyzed output is
+$\overline{b_j}$.  Let $q$ be the first such line.  Note that for
+all $i$ such that $1 \leq i < q$, the value of variable $x_i$ in
+assignment $\alpha'$ does not match the canalyzing value of line
+$i$, but the value of variable $x_q$ does match the canalyzing value
+of line $q$.  Since canalyzed output $b_q$ equals $\overline{b_j}$,
 we have that $f(\alpha') = \overline{b_j}$.
 
-{\bf Case 2:} $(x_k : a_j \rightarrow \overline{b_j})$. \\
-Let line $q$ be the first line such that $j < q \leq k$ and
-the canalyzed output of line $q$ is $\overline{b_j}$.
-Note that $q$ might possibly equal $k$.
-For all $i$ such that $1 \leq i < q$, 
-the value of variable $x_i$ in assignment $\alpha'$ does not match the canalyzing value of line $i$,
-but the value of variable $x_q$ does match the canalyzing value of line $q$.
-Since canalyzed output $b_q$ equals $\overline{b_j}$,
-we have that $f(\alpha') = \overline{b_j}$.
+Now suppose there is no line below line $j$ for which the canalyzed
+output is $\overline{b_j}$.  Then lines $j$ and $k$ both occur in
+the last layer of $f$.  Thus, for every line $i$, $1 \leq i \leq
+n$, the value of variable $x_i$ in assignment $\alpha'$ does not
+match the canalyzing value of line $i$.  Consequently, $f(\alpha')
+= \overline{b_j}$, the complement of the canalyzed output in the
+last layer of $f$.
 
-{\bf Case 3:} $(x_k : \overline{a_j} \rightarrow b_j)$. \\
-Suppose there is a line below line $j$ for which the canalyzed output is $\overline{b_j}$.
-Let $q$ be the first such line.
-Note that for all $i$ such that $1 \leq i < q$, 
-the value of variable $x_i$ in assignment $\alpha'$ does not match the canalyzing value of line $i$,
-but the value of variable $x_q$ does match the canalyzing value of line $q$.
-Since canalyzed output $b_q$ equals $\overline{b_j}$,
-we have that $f(\alpha') = \overline{b_j}$.
+{\bf Case 4:} $(x_k : \overline{a_j} \rightarrow \overline{b_j})$.
+\\ Suppose there is a line, other than line $k$, below line $j$ for
+which the canalyzed output is $\overline{b_j}$.  Let $q$ be the
+first such line.  Note that for all $i$ such that $1 \leq i < q$,
+the value of variable $x_i$ in assignment $\alpha'$ does not match
+the canalyzing value of line $i$, but the value of variable $x_q$
+does match the canalyzing value of line $q$.  Since canalyzed output
+$b_q$ equals $\overline{b_j}$, we have that $f(\alpha') =
+\overline{b_j}$.
 
-Now suppose there is no line below line $j$ for which the canalyzed output is $\overline{b_j}$.
-Then lines $j$ and $k$ both occur in the last layer of $f$.
-Thus, for every line $i$, $1 \leq i \leq n$, 
-the value of variable $x_i$ in assignment $\alpha'$ does not match the canalyzing value of line $i$.
-Consequently, $f(\alpha') = \overline{b_j}$, the complement of the canalyzed output in the last layer of $f$.
-
-{\bf Case 4:} $(x_k : \overline{a_j} \rightarrow \overline{b_j})$. \\
-Suppose there is a line, other than line $k$, below line $j$ for which the canalyzed output is $\overline{b_j}$.
-Let $q$ be the first such line.
-Note that for all $i$ such that $1 \leq i < q$, 
-the value of variable $x_i$ in assignment $\alpha'$ does not match the canalyzing value of line $i$,
-but the value of variable $x_q$ does match the canalyzing value of line $q$.
-Since canalyzed output $b_q$ equals $\overline{b_j}$,
-we have that $f(\alpha') = \overline{b_j}$.
-
-Now suppose that line $k$ is the only line below line $j$ for which the canalyzed output is $\overline{b_j}$.
-Since line $j$ has canalyzed output $b_j$, line $k$ is the only line in its layer.
-Since $f$ is normalized, the last layer of $f$ contains at least two lines.
-Thus, there is a last layer following the layer containing line $k$.
-The canalyzed output of all the lines in this last layer is $b_j$.
-Thus, for every line $i$, $1 \leq i \leq n$, 
-the value of variable $x_i$ in assignment $\alpha'$ does not match the canalyzing value of line $i$.
-Consequently, $f(\alpha') = \overline{b_j}$, the complement of the canalyzed output in the last layer of $f$.
-\QED
+Now suppose that line $k$ is the only line below line $j$ for which
+the canalyzed output is $\overline{b_j}$.  Since line $j$ has
+canalyzed output $b_j$, line $k$ is the only line in its layer.
+Since $f$ is normalized, the last layer of $f$ contains at least
+two lines.  Thus, there is a last layer following the layer containing
+line $k$.  The canalyzed output of all the lines in this last layer
+is $b_j$.  Thus, for every line $i$, $1 \leq i \leq n$, the value
+of variable $x_i$ in assignment $\alpha'$ does not match the
+canalyzing value of line $i$.  Consequently, $f(\alpha') =
+\overline{b_j}$, the complement of the canalyzed output in the last
+layer of $f$.  \QED
 
@@
 \textbf{Proof:}~
-From Theorem \ref{thm:ncf_symmetric_variables}, 
+From Theorem~\ref{thm:ncf_symmetric_variables}, 
 any pair of variables occurring in the same layer, with the same canalyzing value,
@@
 \begin{corollary}\label{cor:ncf_r_symmetric_layers}
-A NCF with a normalized nested canalyzing representation consisting of $q$ layers is $2q$-symmetric,
-and is not $(q-1)$-symmetric.
+A NCF with a normalized nested canalyzing representation consisting
+of $q$ layers is $2q$-symmetric, and is not $(q-1)$-symmetric.
 
-A $r$-symmetric NCF has a normalized nested canalyzing representation with at most $r-1$ layers,
-and has an unnormalized nested canalyzing representation with at most $r$ layers.
-\end{corollary}
+An $r$-symmetric NCF has a normalized nested canalyzing representation
+with at most $r-1$ layers, and has an unnormalized nested canalyzing
+representation with at most $r$ layers.  \end{corollary}
+
+\subsection{Strong Asymmetry and NCFs}
+\label{sse:strong_asym_ncf}
 
@@
 $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
-In general, there are functions with $n$ variables that are properly $n$-symmetric,
-but not strongly asymmetric \cite{KS-2000}.
+In general, there are functions with $n$ variables that are properly
+$n$-symmetric, but not strongly asymmetric \cite{KS-2000}.
 
@@
 \textbf{Proof:}~ 
-If a NCF is $(n-1)$-symmetric, then consider a permutation than interchanges two variables 
-from a symmetry group with at least two members.
-The value of the function is invariant under any such permutation,
-so the function is not strongly asymmetric.
+If a NCF is $(n-1)$-symmetric, then consider a permutation than
+interchanges two variables from a symmetry group with at least two
+members.  The value of the function is invariant under any such
+permutation, so the function is not strongly asymmetric.
 
 Now consider a $n$ variable NCF that is properly $n$-symmetric.
-Let $f$ be the normalized NCF representation for the function.
-For purposes of notational simplicity, 
-wlog we assume that the canalyzing variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.
-Let $\pi$ be any permutation of $\{1, 2, \ldots, n\}$
-\emph{except} the identity permutation.
-We will construct an assignment $(c_1, c_2, \ldots,  c_n)$ 
-to the variables of $f$ such that $f(c_1, c_2, \ldots,  c_n)$ $\neq$  
-$f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)})$.
+Let $f$ be the normalized NCF representation for the function.  For
+purposes of notational simplicity, wlog we assume that the canalyzing
+variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.  Let $\pi$
+be any permutation of $\{1, 2, \ldots, n\}$ \emph{except} the
+identity permutation.  We will construct an assignment $(c_1, c_2,
+\ldots,  c_n)$ to the variables of $f$ such that $f(c_1, c_2, \ldots,
+c_n)$ $\neq$ $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)})$.
 
-Let $k$ be the smallest index such that $k \neq \pi(k)$.
-Since $\pi$ is a permutation, $k < n$.
-Overall, for $1 \leq i < k$, $\pi(i) = i$;
- $\pi(k) > k$;
- and for $k < i \leq n$, $\pi(i) \geq k$.
+Let $k$ be the smallest index such that $k \neq \pi(k)$.  Since
+$\pi$ is a permutation, $k < n$.  Overall, for $1 \leq i < k$,
+$\pi(i) = i$;
+ $\pi(k) > k$; and for $k < i \leq n$, $\pi(i) \geq k$.
 
-Suppose that line $i$ of $f$, $1 \leq i \leq n$, is 
-\noindent
-$(x_i : a_i \rightarrow b_i)$. \\
-Assignment $(c_1, c_2, \ldots,  c_n)$ is constructed as follows.
+Suppose that line $i$ of $f$, $1 \leq i \leq n$, is \noindent $(x_i
+: a_i \rightarrow b_i)$. \\ Assignment $(c_1, c_2, \ldots,  c_n)$
+is constructed as follows.
 
 \noindent
-For $1 \leq i < k$, set $c_i = \overline{a_i}$. \\
-For $i = k$, set $c_k =a_k$. \\
-For $i > k$, let $i' = \pi^{-1}(i)$, so that $\pi(i') = i$.
-Thus, after the permutation of values, variable $x_{i'}$ will have value $c_i$.
-If $b_{i'} = b_k$, then set $c_i =  \overline{a_{i'}}$,
-else set $c_i =  a_{i'}$.
+For $1 \leq i < k$, set $c_i = \overline{a_i}$. \\ For $i = k$, set
+$c_k =a_k$. \\ For $i > k$, let $i' = \pi^{-1}(i)$, so that $\pi(i')
+= i$.  Thus, after the permutation of values, variable $x_{i'}$
+will have value $c_i$.  If $b_{i'} = b_k$, then set $c_i =
+\overline{a_{i'}}$, else set $c_i =  a_{i'}$.
 
-Since $c_k$ matches the canalyzing value of line $k$ of $f$,
-and $c_i$ does not match the canalyzing value of any other earlier line $i$, $1 \leq i < k$, 
-we have that $f(c_1, c_2, \ldots,  c_n) = b_k$.
+Since $c_k$ matches the canalyzing value of line $k$ of $f$, and
+$c_i$ does not match the canalyzing value of any other earlier line
+$i$, $1 \leq i < k$, we have that $f(c_1, c_2, \ldots,  c_n) = b_k$.
 
-Now consider $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)})$.
-We first note that for $1 \leq i < k$, 
-$c_{\pi(i)}$ does not match the canalyzing value of line $i$.
+Now consider $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)})$.  We
+first note that for $1 \leq i < k$, $c_{\pi(i)}$ does not match the
+canalyzing value of line $i$.
 
-Let $k' = \pi^{-1}(k)$, so that $\pi(k') = k$.
-Canalyzing value $a_{k'}$ is either the same or the complement of $a_k$,
-and canalyzed output $b_{k'}$ is either the same or the complement of $b_k$,
-Thus, there are four possible cases for the form of line $k'$ of $f$.
-We will show that in each of the four cases,
-$f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.
+Let $k' = \pi^{-1}(k)$, so that $\pi(k') = k$.  Canalyzing value
+$a_{k'}$ is either the same or the complement of $a_k$, and canalyzed
+output $b_{k'}$ is either the same or the complement of $b_k$, Thus,
+there are four possible cases for the form of line $k'$ of $f$.  We
+will show that in each of the four cases, $f(c_{\pi(1)}, c_{\pi(2)},
+\ldots, c_{\pi(n)}) = \overline{b_k}$.
 
+\noindent
 {\bf Case 1:} $(x_{k'} : a_k \rightarrow b_k)$. \\
-Since line $k'$ has the same canalyzing value and canalyzed output as line $k$,
-and $f$ is properly $n$-symmetric,
-line $k'$ must occur in a lower layer than that of line $k$.
-Thus, there is at least one line between line $k$ and line $k'$ for which the canalyzed output is $\overline{b_k}$.
-Let line $q$ be the first such line.
-Note that for all $i$ such that $1 \leq i < q$, 
-$c_{\pi(i)}$ does not match the canalyzing value of line $i$,
-but $c_{\pi(q)}$ does match the canalyzing value of line $q$.
-Since canalyzed output $b_q$ equals $\overline{b_k}$,
-we have that $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.
 
+Since line $k'$
+has the same canalyzing value and canalyzed output as line $k$, and
+$f$ is properly $n$-symmetric, line $k'$ must occur in a lower layer
+than that of line $k$.  Thus, there is at least one line between
+line $k$ and line $k'$ for which the canalyzed output is $\overline{b_k}$.
+Let line $q$ be the first such line.  Note that for all $i$ such
+that $1 \leq i < q$, $c_{\pi(i)}$ does not match the canalyzing
+value of line $i$, but $c_{\pi(q)}$ does match the canalyzing value
+of line $q$.  Since canalyzed output $b_q$ equals $\overline{b_k}$,
+we have that $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) =
+\overline{b_k}$.
+
+\noindent
 {\bf Case 2:} $(x_{k'} : a_k \rightarrow \overline{b_k})$. \\
+
 Let line $q$ be the first line such that $k < q \leq k'$ and
@@
 
+\noindent
 {\bf Case 3:} $(x_{k'} : \overline{a_k} \rightarrow b_k)$. \\
-Suppose there is a line below line $k$ for which the canalyzed output is $\overline{b_k}$.
-Let $q$ be the first such line.
-Note that for all $i$ such that $1 \leq i < q$, 
-$c_{\pi(i)}$ does not match the canalyzing value of line $i$,
-but $c_{\pi(q)}$ does match the canalyzing value of line $q$.
-Thus, $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.
 
-Now suppose there is no line below line $k$ for which the canalyzed output is $\overline{b_k}$.
-Then line $k$ occurs in the last layer of $f$.
-Since $f$ is both normalized and properly $n$-symmetric,
-this last layer contains exactly two lines.
-Thus $k = n-1$ and $k' = n$.
-Note that for all $i$, 
-$c_{\pi(i)}$ does not match the canalyzing value of line $i$.
-Thus, $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.
+Suppose there is a line below line $k$ for which the canalyzed
+output is $\overline{b_k}$.  Let $q$ be the first such line.  Note
+that for all $i$ such that $1 \leq i < q$, $c_{\pi(i)}$ does not
+match the canalyzing value of line $i$, but $c_{\pi(q)}$ does match
+the canalyzing value of line $q$.  Thus, $f(c_{\pi(1)}, c_{\pi(2)},
+\ldots, c_{\pi(n)}) = \overline{b_k}$.
 
+Now suppose there is no line below line $k$ for which the canalyzed
+output is $\overline{b_k}$.  Then line $k$ occurs in the last layer
+of $f$.  Since $f$ is both normalized and properly $n$-symmetric,
+this last layer contains exactly two lines.  Thus $k = n-1$ and $k'
+= n$.  Note that for all $i$, $c_{\pi(i)}$ does not match the
+canalyzing value of line $i$.  Thus, $f(c_{\pi(1)}, c_{\pi(2)},
+\ldots, c_{\pi(n)}) = \overline{b_k}$.
+
+\noindent
 {\bf Case 4:} $(x_{k'} : \overline{a_k} \rightarrow \overline{b_k})$. \\
-Suppose there is a line, other than line $k'$, below line $k$ for which the canalyzed output is $\overline{b_k}$.
-Let $q$ be the first such line.
-Note that for all $i$ such that $1 \leq i < q$, 
-$c_{\pi(i)}$ does not match the canalyzing value of line $i$,
-but $c_{\pi(q)}$ does match the canalyzing value of line $q$.
-Thus, $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.
 
-Now suppose that line $k'$ is the only line below line $k$ for which the canalyzed output is $\overline{b_k}$.
-Since line $k$ has canalyzed output $b_k$, line $k'$ is the only line in its layer.
-Since $f$ is normalized, there is a last layer following the layer containing line $k'$.
-Thus, for all $i$, 
-$c_{\pi(i)}$ does not match the canalyzing value of line $i$,
-so $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.
-\QED
+Suppose there is a line, other than line $k'$, below line $k$ for
+which the canalyzed output is $\overline{b_k}$.  Let $q$ be the
+first such line.  Note that for all $i$ such that $1 \leq i < q$,
+$c_{\pi(i)}$ does not match the canalyzing value of line $i$, but
+$c_{\pi(q)}$ does match the canalyzing value of line $q$.  Thus,
+$f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) = \overline{b_k}$.
 
-Examples of properly $n$-symmetric $n$-variable NCFs
-satisfying the conditions of Theorem \ref{thm:ncf_r_symmetric}
-are the following. For each $n > 1$, let $f_n$ be the function defined by the following formula.
+Now suppose that line $k'$ is the only line below line $k$ for which
+the canalyzed output is $\overline{b_k}$.  Since line $k$ has
+canalyzed output $b_k$, line $k'$ is the only line in its layer.
+Since $f$ is normalized, there is a last layer following the layer
+containing line $k'$.  Thus, for all $i$, $c_{\pi(i)}$ does not
+match the canalyzing value of line $i$, so $f(c_{\pi(1)}, c_{\pi(2)},
+\ldots, c_{\pi(n)}) = \overline{b_k}$.  \QED
+
+Examples of properly $n$-symmetric $n$-variable NCFs satisfying the
+conditions of Theorem \ref{thm:ncf_r_symmetric} are the following.
+For each $n > 1$, let $f_n$ be the function defined by the following
+formula.
 
@@
 
-Function $f_n(x_1, x_2, \ldots, x_n)$ is the NCF corresponding to the following NCF representation:
+Function $f_n(x_1, x_2, \ldots, x_n)$ is the NCF corresponding to
+the following NCF representation:
 
@@
 Note that if $x_i = 0$ for all $i$, $1 \leq i \leq n$, then the
-value of $f_n$ is 0 if $n$ is odd and 1 if $n$ is even.
-Also, note that the above representation is unnormalized. To normalize the representation,
-the last line would be changed to have canalyzing value 0, and the same canalyzed output as the preceding line.
+value of $f_n$ is 0 if $n$ is odd and 1 if $n$ is even.  Also, note
+that the above representation is unnormalized. To normalize the
+representation, the last line would be changed to have canalyzing
+value 0, and the same canalyzed output as the preceding line.
 
@@
 Equation~(\ref{eqn:f_asymmetry}), and the result follows. \QED
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \fi

### 2019-01-25 abstract.tex
ssr

@@
 
-Several researchers have studied 
-a class of Boolean functions, 
+A class of Boolean functions, 
 called \textbf{nested canalyzing functions} (NCFs),
-which are used to model certain biological 
-phenomena \cite{Kauffman-1969,Kauffman-etal-2003,Kauffman-etal-2004}.
-We study the relationships between NCFs and symmetric 
-Boolean functions. Our results ...
+has been used to model certain biological phenomena.
+Rearchers have studied some properties of these functions
+such as sensitivity and stability.
+In this note, we examine the relationships between NCFs and symmetric 
+Boolean functions. 
+We identify all NCFs that are also symmetric. 
+We also consider a more general form of symmetric Boolean functions
+called $r$-symmetric functions, where $r$ is the symmetry level.
+We show how one can get lower and upper bounds on the symmetry
+level of an NCF $f$ given a suitable respresentation of $f$.

### 2019-01-25 symmetry_properties.tex
ssr

@@
 
-{\bf Case 1:} $(x_k : a_j \rightarrow b_j)$. \\ Since line $k$ has
+\noindent
+{\bf Case 1:} $(x_k : a_j \rightarrow b_j)$. \\
+
+Since line $k$ has
 the same canalyzing value and canalyzed output as line $j$, line
@@
 
-{\bf Case 2:} $(x_k : a_j \rightarrow \overline{b_j})$. \\ Let line
+\noindent
+{\bf Case 2:} $(x_k : a_j \rightarrow \overline{b_j})$. \\
+
+Let line
 $q$ be the first line such that $j < q \leq k$ and the canalyzed
@@
 
-{\bf Case 3:} $(x_k : \overline{a_j} \rightarrow b_j)$. \\ Suppose
-there is a line below line $j$ for which the canalyzed output is
+\noindent
+{\bf Case 3:} $(x_k : \overline{a_j} \rightarrow b_j)$. \\
+
+Suppose there is a line below line $j$ for which the canalyzed output is
 $\overline{b_j}$.  Let $q$ be the first such line.  Note that for
@@
 
-{\bf Case 4:} $(x_k : \overline{a_j} \rightarrow \overline{b_j})$.
-\\ Suppose there is a line, other than line $k$, below line $j$ for
+\noindent
+{\bf Case 4:} $(x_k : \overline{a_j} \rightarrow \overline{b_j})$.  \\ 
+Suppose there is a line, other than line $k$, below line $j$ for
 which the canalyzed output is $\overline{b_j}$.  Let $q$ be the

### 2019-01-28 abstract.tex
ssr

@@
 We identify all NCFs that are also symmetric. 
-We also consider a more general form of symmetric Boolean functions
-called $r$-symmetric functions, where $r$ is the symmetry level.
-We show how one can get lower and upper bounds on the symmetry
-level of an NCF $f$ given a suitable respresentation of $f$.
+A more general form of symmetric Boolean functions,
+called $r$-symmetric functions (where $r$ is the symmetry level),
+have also been studies in the literature.
+We establish lower and upper bounds on the symmetry
+level of an NCF $f$ as functions of the number of layers
+in a standard respresentation of $f$.

### 2019-01-28 intro.tex
ssr

@@
 \label{sse:ncf_def}
+
+The following definition of \textbf{canalyzing} Boolean functions was introduced
+in \cite{Kauffman-1969}.
 
@@
 \textbf{Example 1:}~ Consider the function 
-$f(x_1, x_2, x_3) ~=~ x_1 \vee (x_2 \wedge x_3)$.
+$f(x_1, x_2, x_3) ~=~ \overline{x_1} \wedge (x_2 \vee \overline{x_3})$.
 This function is canalyzing since if we set to $x_1 = 1$,
-$f(1, x_2, x_3) ~=~ 1$, regardless of the values assigned to $x_2$ and $x_3$. \qed
+$f(1, x_2, x_3) ~=~ 0$, regardless of the values assigned to $x_2$ and $x_3$. \qed
 
@@
 
-A class of Boolean functions, called \textbf{nested canalyzing functions} (NCFs),
-was introduced in \cite{Kauffman-etal-2003} to model the behavior of
+Another class of Boolean functions, called \textbf{nested canalyzing functions} (NCFs),
+was introduced later in \cite{Kauffman-etal-2003} to model the behavior of
 certain biological systems.
@@
 %% \clearpage
-For convenience, we will use a computational notation to represent NCFs.
-For $1 \leq i \leq n$, line $i$ of our representation has the following form:
+For convenience, we will use a computational notation to represent NCFs
+introduced in \cite{Stearns-etal-2018}.
+For $1 \leq i \leq n$, line $i$ of this representation has the following form:
 
@@
 \noindent 
-We say that $x_{\pi(i)}$ is the \textbf{canalyzing variable} that is
+Following the terminology in \cite{Stearns-etal-2018},
+ $x_{\pi(i)}$ is the \textbf{canalyzing variable} that is
 \textbf{tested} in line $i$, 
@@
 canalyzed values in line $i$ as before,~ $1 \leq i \leq n$.
-We refer to each such line as a \textbf{rule}.
+each such line is called a \textbf{rule}.
 When none of the conditions ``$x_{\pi(i)} ~=~ a_i$" 
@@
 \noindent
-In the remainder of this paper, we will refer to the above specification
+As in \cite{Stearns-etal-2018},, we will refer to the above specification
 of an NCF as the \textbf{simplified representation} and assume
 (without loss of generality) that each NCF is specified in this manner.
-The simplified representation provides the following 
-convenient computational view of an NCF.~
-Lines defining an NCF are 
-considered sequentially in a top-down manner.
-The computation stops at the first line where the 
-specified condition is satisfied, and the value of the function
-is the canalyzed value on that line. 
+%% The simplified representation provides the following 
+%% convenient computational view of an NCF.~
+%% Lines defining an NCF are 
+%% considered sequentially in a top-down manner.
+%% The computation stops at the first line where the 
+%% specified condition is satisfied, and the value of the function
+%% is the canalyzed value on that line. 
 We now present an example of an NCF using the two representations
@@
 \textbf{Example 2:}~ Consider the function 
-$f(x_1, x_2, x_3) ~=~ x_1 \vee (x_2 \wedge x_3)$.
+$f(x_1, x_2, x_3) ~=~ \overline{x_1} \wedge (x_2 \vee \overline{x_3})$.
 This function is nested canalyzing using the identity permutation $\pi$ on $\{1,2,3\}$
-with canalyzing values $1,0,0$ and canalyzed values $1, 0, 0$.
+with canalyzing values $1,1,0$ and canalyzed values $0, 1, 1$.
 We first show how this function can be expressed using the syntax of
@@
    \begin{cases}
-       \:1 & \mathrm{if~~} x_{1} ~=~ 1 \\
+       \:0 & \mathrm{if~~} x_{1} ~=~ 1 \\
+       \:1 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~}
+            x_{2} ~=~ 1 \\
+       \:1 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~}
+            x_{2} ~\neq~ 1 \mathrm{~~and~~} x_{3} = 0 \\
        \:0 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~}
-            x_{2} ~=~ 0 \\
-       \:0 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~}
-            x_{2} ~\neq~ 0 \mathrm{~~and~~} x_{3} = 0 \\
-       \:1 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~}
-            x_{2} ~\neq~ 0 \mathrm{~~and~~} x_{3} ~\neq~ 0 \\
+            x_{2} ~\neq~ 1 \mathrm{~~and~~} x_{3} ~\neq~ 0 \\
    \end{cases}
@@
 \begin{tabular}{ll}
-\hspace*{1.1in} & $x_1:~$  $1 ~\longrightarrow~ 1$ \\ [1ex]
-\hspace*{1.1in} & $x_2:~$  $0 ~\longrightarrow~ 0$ \\ [1ex]
-\hspace*{1.1in} & $x_3:~$  $0 ~\longrightarrow~ 0$ \\ [1ex]
-\hspace*{1.1in} & Default:~ $1$ \\
+\hspace*{1.1in} & $x_1:~$  $1 ~\longrightarrow~ 0$ \\ [1ex]
+\hspace*{1.1in} & $x_2:~$  $1 ~\longrightarrow~ 1$ \\ [1ex]
+\hspace*{1.1in} & $x_3:~$  $0 ~\longrightarrow~ 1$ \\ [1ex]
+\hspace*{1.1in} & Default:~ $0$ \\
 \end{tabular}
@@
 
+\subsection{Symmetric Boolean Functions}
+\label{sse:symmetry}
+
+A pair of variables of a Boolean function $f(x_1, x_2, \ldots, x_n)$ is
+said to be \textbf{symmetric} if their values can be interchanged without
+affecting the value of the function.
+This form of interchange symmetry partitions the set of variables into a set of
+\textbf{symmetry groups}, where the members of each group are pairwise symmetric.
+Function $f$ is said to be
+$r$-\textbf{symmetric} if it has at most $r$ symmetry groups.
+In this case, the value of $f$
+depends only on how many of the variables in each symmetry group have the value 1.
+We say that $f$ is \textbf{properly} $r$-\textbf{symmetric} if
+it is $r$-symmetric, but not $(r-1)$-symmetric.
+
+In the literature (see, e.g., \cite{Crama-Hammer-2011,HT-2016,Toth-etal-1977}),
+a 1-symmetric function $f$ is referred to simply
+as a \textbf{symmetric} function.
+Thus, if $f$ is a symmetric function, then for any
+input $(a_1, a_2, \ldots, a_n)$ to $f$, where $a_i \in \{0,1\}$ for
+$1 \leq i \leq n$, and any permutation $\pi$ of $\{1, 2, \ldots, n\}$,
+$f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
+
 %\noindent

### 2019-01-29 intro.tex
ssr

@@
 This function is canalyzing since if we set to $x_1 = 1$,
-$f(1, x_2, x_3) ~=~ 0$, regardless of the values assigned to $x_2$ and $x_3$. \qed
+$f(1, x_2, x_3) ~=~ 0$, no matter which combination of values is assigned 
+to $x_2$ and $x_3$. \qed
 
@@
 \noindent
-As in \cite{Stearns-etal-2018},, we will refer to the above specification
+As in \cite{Stearns-etal-2018}, we will refer to the above specification
 of an NCF as the \textbf{simplified representation} and assume
@@
 it is $r$-symmetric, but not $(r-1)$-symmetric.
+For a Boolean function $f$, the integer $r$ such that $f$ is properly $r$-symmetric
+will be referred to as the \textbf{symmetry level} of $f$.
+Thus the symmetry level of $f$ is the smallest integer $r$ such that
+$f$ is $r$-symmetric.
 
@@
 $f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
+
+The following result points out that given a Boolean function $f$
+in the form of a Boolean expression, it is \np-hard to approximate
+the symmetry level to within any factor $\rho \geq 1$.
+
+\begin{theorem}\label{thm:approx_sym_level_hard}
+For any $\rho \geq 1$, it is \cnp-hard to approx
+\end{theorem}
 

### 2019-01-30 intro.tex
ssr

@@
 
+\iffalse
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 The following result points out that given a Boolean function $f$
-in the form of a Boolean expression, it is \np-hard to approximate
+in the form of a Boolean expression, it is \cnp-hard to approximate
 the symmetry level to within any factor $\rho \geq 1$.
@@
 \end{theorem}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\fi
 

### 2019-01-30 preliminaries.tex
ssr

@@
 
-\subsection{Defintions: Symmetric Boolean Functions}
-\label{sse:symmetry}
+\subsection{Complexity of Estimating the Symmetry Level of a i
+Boolean Function}
+\label{sse:symmetry_level_hardness}
 
-A pair of variables of a Boolean function $f(x_1, x_2, \ldots, x_n)$ is
-said to be \textbf{symmetric} if their values can be interchanged without
-affecting the value of the function.
-This form of interchange symmetry partitions the set of variables into a set of
- \textbf{symmetry groups}, where the members of each group are pairwise symmetric.
-Function $f$ is said to be
-$r$-\textbf{symmetric} if it has at most $r$ symmetry groups.
-In this case, the value of $f$
-depends only on how many of the variables in each symmetry group have the value 1.
-We say that $f$ is \textbf{properly} $r$-\textbf{symmetric} if
-it is $r$-symmetric, but not $(r-1)$-symmetric.
+Recall that the symmetry level of a Boolean function $f$ is the smallest
+integer $r$ such that the following condition holds:
+the  inputs to $f$ can be partitioned into $r$ subsets 
+(symmetry groups) so that the value of $f$ depends only on 
+how many of the inputs in each group have the value 1.
+The following result shows that given a Boolean function $f$
+in the form of a Boolean expression, it is \cnp-hard to 
+approximate\footnote{An algorithm approximates the symmetry
+level of a Boolean function $f$ within the factor $\rho$ if it finds a
+partition of the inputs to $f$ into at most $\rho \timex r^*$ symmetry
+groups, where $r^*$ is the symmetry level of $f$.}
+the symmetry level of $f$ to within any factor $\rho \geq 1$.
 
-In the literature, a 1-symmetric function $f$ is referred to simply
-as a \textbf{symmetric} function.
-Thus, if $f$ is a symmetric function, then for any
-input $(a_1, a_2, \ldots, a_n)$ to $f$, where $a_i \in \{0,1\}$ for
-$1 \leq i \leq n$, and any permutation $\pi$ of $\{1, 2, \ldots, n\}$,
-$f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
+\newcommand{\cala}{\mbox{$\mathcal{A}$}}
+
+\begin{proposition}\label{pro:approx_sym_level_hard}
+For any $\rho \geq 1$, it is \cnp-hard to approximate the
+symmetry level of $f$ to within the factor $\rho$.
+\end{proposition}
+
+\noindent
+\textbf{Proof:}~ Suppose there is an efficient algorithm \cala{} 
+that provides a $\rho$ approximation for the symmetry level
+of a Boolean function for some $\rho \geq 1$.
+We will show that \cala{} can be used to efficiently solve
+the CNF Satisfiability problem (SAT) which is known 
+to be \cnp-hard \cite{GJ-1979}.
+
+Let $f$ be a CNF formula representing an instance of SAT.
+Let $X$ = $\{x_1, x_2, \ldots, x_n\}$ denote the set of
+Boolean variables used in $f$.
+
+\QED
+
+In contrast, we will show that the symmetry level of an NCF (or that of a
+generalized NCF) can be computed efficiently.

### 2019-01-30 preliminaries.tex
ssr

@@
 level of a Boolean function $f$ within the factor $\rho$ if it finds a
-partition of the inputs to $f$ into at most $\rho \timex r^*$ symmetry
+partition of the inputs to $f$ into at most $\rho \times r^*$ symmetry
 groups, where $r^*$ is the symmetry level of $f$.}
@@
 
-\begin{proposition}\label{pro:approx_sym_level_hard}
-For any $\rho \geq 1$, it is \cnp-hard to approximate the
-symmetry level of $f$ to within the factor $\rho$.
-\end{proposition}
+\begin{theorem}\label{thm:approx_sym_level_hard}
+Unless \textbf{P} = \cnp,
+for any $\rho \geq 1$, there is no polynomial time 
+algorithm for approximating the 
+symmetry level of a Boolean function $f$ to within the factor $\rho$.
+This result holds even when $f$ is a CNF expression.
+\end{theorem}
 
@@
 Boolean variables used in $f$.
+We create another CNF formula $f'$ as follows.
+Let $Y = \{y_1, y_2, \ldots, y_{\rho+1}\}$ and
+$Z = \{z_1, z_2, \ldots, z_{\rho+1}\}$ be two new sets of variables.
+The expression for $f'$ is given by
+\[
+f'(x_1, \ldots, x_n, y_1, \ldots, y_{\rho+1},
+z_1, \ldots, z_{\rho+1}) ~=~ 
+    f(x_1, \ldots x_n) \wedge (y_1 \vee \overline{z_1}) \wedge \ldots \wedge
+                              (y_{\rho+1} \vee \overline{z_{\rho+1}}).
+\]
+We note that $f'$ is also a CNF formula.
+We have the following claims.
 
+\medskip
+
+\noindent
+\textbf{Claim 1:}~ If $f$ is \emph{not} satisfiable, 
+the symmetry level of $f'$ is 1.
+
+\smallskip
+
+\noindent
+\textbf{Proof of Claim 1:}~ If $f$ is not satisfiable, $f'$ is also
+not satisfiable; that is, for all inputs, the value of $f'$ is 0.
+Thus, all the variables in $X \cup Y \cup Z$ can be put in one
+symmetry group; regardless of the number of variables 
+which are 1 in the group, the value of $f'$ is always 0. 
+\hfill$\Box$
+
+\medskip
+
+\noindent
+\textbf{Claim 2:}~ If $f$ \emph{is} satisfiable, 
+the symmetry level of $f'$ is at least $\rho+1$.
+
+\smallskip
+
+\noindent
+\textbf{Proof of Claim 2:}~ Suppose $f$ is satisfiable.
+We argue that for $i \neq j$, variables $y_i$ and $y_j$
+cannot be in the same symmetry group.
+To see this, consider the following assignment $\alpha$ to
+the variables of $f'$.
+\begin{description}
+\item{(i)} For the variables in $X$, choose an assignment that
+satisfies $f$.
+\item{(ii)} Let $y_i = 0$, $z_i = 1$, $y_j = 1$ and $z_j = 0$.
+\item{(iii)} For $1 \leq p \leq \rho+1$, $p \neq i$ and $p \neq j$,
+let $y_p = 1$ and $z_p$ = 0.
+\end{description}
+Since this assignment $\alpha$ sets the clause $(y_i \vee \overline{z_i})$ to 0,
+the value of $f'$ under the assignment $\alpha$ is 0.
+Now, consider the assignment $\alpha'$ that is obtained by exchanging
+the values of $y_i$ and $y_j$ in $\alpha$ without changing the
+values assigned to the other variables. 
+It can be seen that under assignment $\alpha'$, the value of $f'$ is 1.
+In both $\alpha$ and $\alpha'$, the number of 1-valued variables in 
+the set $\{y_i, y_j\}$ is 1; however, the two assignments lead to 
+different values for $f'$.
+Therefore, if $f$ is not satisfiable, for $i \neq j$, $y_i$ and $y_j$ cannot
+be in the same symmetry group.
+Since there are $\rho+1$ variables in $Y$, the number of symmetry groups 
+for $f'$ must be at least $\rho+1$. \hfill$\Box$
+
+\medskip
+
+Now, suppose we give $f'$ as the input to Algorithm \cala.
+If $f$ is \emph{not} satisfiable, then from Claim 1, 
+the symmetry level of $f'$ is 1. 
+Since \cala{} is a $\rho$-approximation algorithm,
+it should produce at most $\rho$ symmetry groups.
+On the other hand, if $f$ is satisfiable, by Claim 2, the symmetry
+level of $f'$ is at least $\rho+1$ and \cala{} should produce at least
+$\rho+1$ groups.
+Thus, by examining the output of \cala, we can determine whether or not $f$
+is satisfiable.
+Since \cala{} runs in polynomial time, we have an efficient 
+algorithm for SAT, contradicting the assumption that \textbf{P} $\neq$ \cnp.
 \QED
 
-In contrast, we will show that the symmetry level of an NCF (or that of a
+\medskip
+
+In contrast to the above result,
+we will show that the symmetry level of an NCF (or that of a
 generalized NCF) can be computed efficiently.

### 2019-01-30 abstract.tex
ssr

--- 2019-01-25 abstract.tex
@@
 \begin{center}
-\textbf{Abstract}~~ \textcolor{red}{\textbf{(To be revised)}}
+\textbf{Abstract}  %% ~~ \textcolor{red}{\textbf{(To be revised)}}
 \end{center}
@@
 has been used to model certain biological phenomena.
-Rearchers have studied some properties of these functions
+Researchers have studied some properties of these functions
 such as sensitivity and stability.
@@
 Boolean functions. 
-We identify all NCFs that are also symmetric. 
-We also consider a more general form of symmetric Boolean functions
-called $r$-symmetric functions, where $r$ is the symmetry level.
-We show how one can get lower and upper bounds on the symmetry
-level of an NCF $f$ given a suitable respresentation of $f$.
+We identify all the Boolean functions that both  NCFs and symmetric. 
+A more general form of symmetric Boolean functions,
+called $r$-symmetric functions (where $r$ is the symmetry level),
+have also been studied in the literature.
+%% We establish lower and upper bounds on the symmetry
+%% level of an NCF $f$ as functions of the number of layers
+%% in a standard respresentation of $f$.  
+We show that, in general,
+one cannot efficiently approximate the symmetry level of
+a Boolean function to within any factor, unless \textbf{P} = \cnp. 
+In contrast, we show that the symmetry level of an NCF $f$
+can be easily computed from a standard (and efficiently
+computable) representation of $f$.

### 2019-01-30 intro.tex
ssr

@@
 
-\iffalse
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \noindent
 \underline{\textsf{Generalized NCFs:}}~
-For some problems, we will also consider a more general form
-of NCFs defined below.
+We will also consider a more general form of NCFs defined below.
 
@@
 
-\bigskip
+%% \bigskip
 
-\noindent
-\textbf{Note:} If a generalized NCF is not a constant, 
-there is an assignment to its variables
-such that the value of the function is 0,
-and there is an assignment to its variables
-such that the value of the function is 1.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\fi
+%% \noindent
+%% \textbf{Note:} If a generalized NCF is not a constant, 
+%% there is an assignment to its variables
+%% such that the value of the function is 0,
+%% and there is an assignment to its variables
+%% such that the value of the function is 1.
 

### 2019-01-30 intro.tex
ssr

@@
 certain biological systems.
-We follow the presentation in \cite{Layne-2011} in defining 
-such a Boolean function. 
+We follow the presentation in \cite{Layne-2011} in defining NCFs.
 (For a Boolean value $b$,~ the complement is denoted by $\overline{b}$.)
@@
 \end{definition}
-
-\medskip
 %% \clearpage
-For convenience, we will use a computational notation to represent NCFs
-introduced in \cite{Stearns-etal-2018}.
+For convenience, we will use a notation introduced in \cite{Stearns-etal-2018}
+to represent NCFs.
 For $1 \leq i \leq n$, line $i$ of this representation has the following form:
@@
 canalyzed values in line $i$ as before,~ $1 \leq i \leq n$.
-each such line is called a \textbf{rule}.
+Each such line is called a \textbf{rule}.
 When none of the conditions ``$x_{\pi(i)} ~=~ a_i$" 
@@
 \textbf{Example 2:}~ Consider the function 
-$f(x_1, x_2, x_3) ~=~ \overline{x_1} \wedge (x_2 \vee \overline{x_3})$.
+$f(x_1, x_2, x_3) ~=~ \overline{x_1} \wedge (x_2 \vee \overline{x_3})$
+used in Example~1.
 This function is nested canalyzing using the identity permutation $\pi$ on $\{1,2,3\}$
@@
 \noindent
-\underline{\textsf{Generalized NCFs:}}~
 We will also consider a more general form of NCFs defined below.
@@
 affecting the value of the function.
+As simple examples, the functions 
+$f_1(x_1, x_2, x_3)$ = $x_1 \vee x_2 \vee x_3$ and
+$f_2(x_1, x_2, x_3)$ = $x_1 \oplus x_2 \oplus x_3$ are both symmetric.
 This form of interchange symmetry partitions the set of variables into a set of
@@
 it is $r$-symmetric, but not $(r-1)$-symmetric.
+For example, the function $f_3(x_1, x_2) = x_1 \vee \overline{x_2}$
+is not symmetric since $f(1, 0) \neq f(0, 1)$; however, 
+it is trivially 2-symmetric.
 For a Boolean function $f$, the integer $r$ such that $f$ is properly $r$-symmetric
 will be referred to as the \textbf{symmetry level} of $f$.
-Thus the symmetry level of $f$ is the smallest integer $r$ such that
-$f$ is $r$-symmetric.
+(Thus, the symmetry level of $f$ is the smallest integer $r$ such that
+$f$ is $r$-symmetric.)
 
@@
 $f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
-
-\iffalse
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The following result points out that given a Boolean function $f$
-in the form of a Boolean expression, it is \cnp-hard to approximate
-the symmetry level to within any factor $\rho \geq 1$.
-
-\begin{theorem}\label{thm:approx_sym_level_hard}
-For any $\rho \geq 1$, it is \cnp-hard to approx
-\end{theorem}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\fi
+Symmetric and $r$-symmetric functions for $r \geq 2$ have also been
+studied in the literature in the context of discrete
+dynamical systems (see, e.g., 
+\cite{Barrett-etal-2007,Rosenkrantz-etal-2015,MR-2007}).
 

### 2019-01-30 intro.tex
ssr

@@
 affecting the value of the function.
-As simple examples, the functions 
-$f_1(x_1, x_2, x_3)$ = $x_1 \vee x_2 \vee x_3$ and
-$f_2(x_1, x_2, x_3)$ = $x_1 \oplus x_2 \oplus x_3$ are both symmetric.
+As a simple example, each pair of variables in the function 
+$f_2(x_1, x_2, x_3)$ = $x_1 \oplus x_2 \oplus x_3$ is symmetric.
 This form of interchange symmetry partitions the set of variables into a set of
@@
 as a \textbf{symmetric} function.
-Thus, if $f$ is a symmetric function, then for any
+Thus, in any symmetric function, each pair of variables is symmetric.
+As a simple example, the function $f_2(x_1, x_2, x_3)$ = 
+$x_1 \oplus x_2 \oplus x_3$ (defined above) is symmetric.
+If $f$ is a symmetric function, then for any
 input $(a_1, a_2, \ldots, a_n)$ to $f$, where $a_i \in \{0,1\}$ for
@@
 for gene regulatory networks.
+It is known that the class of NCFs coincides with that
+of unate cascade functions \cite{}
 Many researchers have studied mathematical properties of NCFs and
@@
 Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}).
-
-There is extensive literature on the sensitivity of various classes
-of Boolean functions
-(e.g., \cite{Buhrman-etal-2002,Nisan-etal-1994,Zhang-2011}).
-For a discussion on how the sensitivity of a Boolean function provides
-an indication of its stability, the reader is referred to \cite{Kauffman-etal-2004,
-Layne-2011,Layne-etal-2012}.
-Li and Adeyeye \cite{Li-etal-2012} present lower and upper bounds on the
-sensitivity of any NCF.
-Li et al. \cite{Li-etal-2011,Li-etal-2013} conjectured that the average
-sensitivity of any NCF is strictly less than  $4/3$.
-This conjecture was proved by Klotz et al. \cite{Klotz-etal-2013} by establishing
-a tight upper bound on the average sensitivity of any NCF.
-Their methods rely on Fourier analysis of Boolean functions
-\cite{Odonnell-2014}.
-In \cite{Stearns-etal-2018}, 
-we presented a characterization of NCFs with maximum average sensitivity.
-This characterization allowed us to present an alternative proof 
-of the upper bound in \cite{Klotz-etal-2013} and derive a closed
-form expression for the number of such NCFs.
+Additionally, properties of NCFs such as sensitivity and 
+stability have also been studied
+by many researchers 
+\cite{Kauffman-etal-2004, Layne-2011,Layne-etal-2012,
+Li-etal-2011,Li-etal-2013,Klotz-etal-2013,
+Stearns-etal-2018}. 
+To our knowledge, the relationship between NCFs and symmetric
+Boolean functions has not been studied in the literature.

### 2019-01-31 abstract.tex
ssr

@@
 Boolean functions. 
-We identify all the Boolean functions that both  NCFs and symmetric. 
+We identify all the Boolean functions that NCFs as well as symmetric. 
 A more general form of symmetric Boolean functions,
 called $r$-symmetric functions (where $r$ is the symmetry level),
-have also been studied in the literature.
+has also been studied in the literature.
 %% We establish lower and upper bounds on the symmetry
@@
 %% in a standard respresentation of $f$.  
-We show that, in general,
+We observe that, in general,
 one cannot efficiently approximate the symmetry level of

### 2019-01-31 intro.tex
ssr

@@
 
-The following definition of \textbf{canalyzing} Boolean functions was introduced
-in \cite{Kauffman-1969}.
+The class of \textbf{canalyzing} Boolean functions, introduced 
+in \cite{Kauffman-1969}, is defined as follows.
 
 \begin{definition}\label{def:canalyzing}
-Let $X = $ $\{x_1, x_2, \ldots, x_n\}$ be a set of $n$  Boolean variables.
-A Boolean function $f(x_1, x_2, \ldots, x_n)$ over $X$ is \textbf{canalyzing}
+Given a set $X = $ $\{x_1, x_2, \ldots, x_n\}$ of $n$  Boolean variables,
+a Boolean function $f(x_1, x_2, \ldots, x_n)$ over $X$ is \textbf{canalyzing}
 if there is a variable $x_i \in X$ and values $a$, $b$ $\in \{0,1\}$ such that
-$f(x_1, \ldots, x_{i-1}, a, x_{i+1}, \ldots, x_n)$ $~=~ b$, regardless of the values
-assigned to the variables in $X - \{x_i\}$.
+\[
+f(x_1, \ldots, x_{i-1}, a, x_{i+1}, \ldots, x_n) ~=~ b, 
+\]
+regardless of the values assigned to the variables in $X - \{x_i\}$.
 \end{definition}
@@
 Another class of Boolean functions, called \textbf{nested canalyzing functions} (NCFs),
-was introduced later in \cite{Kauffman-etal-2003} to model the behavior of
-certain biological systems.
+was introduced later in \cite{Kauffman-etal-2003} to carry out a detailed
+analysis of the behavior of certain biological systems.
 We follow the presentation in \cite{Layne-2011} in defining NCFs.
@@
 to represent NCFs.
-For $1 \leq i \leq n$, line $i$ of this representation has the following form:
+For $1 \leq i \leq n$, line $i$ of this representation has the form
 
@@
 \noindent 
-Following the terminology in \cite{Stearns-etal-2018},
- $x_{\pi(i)}$ is the \textbf{canalyzing variable} that is
+with $x_{\pi(i)}$ being the \textbf{canalyzing variable} that is
 \textbf{tested} in line $i$, 
-with $a_i$ and $b_i$ denoting respectively the canalyzing and 
-canalyzed values in line $i$ as before,~ $1 \leq i \leq n$.
+and $a_i$ and $b_i$ being respectively the canalyzing and 
+canalyzed values in line $i$,~ $1 \leq i \leq n$.
 Each such line is called a \textbf{rule}.
@@
 it is $r$-symmetric, but not $(r-1)$-symmetric.
-For example, the function $f_3(x_1, x_2) = x_1 \vee \overline{x_2}$
-is not symmetric since $f(1, 0) \neq f(0, 1)$; however, 
+For example, the function $f_3(x_1, x_2) = x_1 \vee\, \overline{x_2}$
+is not 1-symmetric since $f_3(1, 0) \neq f_3(0, 1)$; however, 
 it is trivially 2-symmetric.
@@
 will be referred to as the \textbf{symmetry level} of $f$.
-(Thus, the symmetry level of $f$ is the smallest integer $r$ such that
-$f$ is $r$-symmetric.)
+Thus, the symmetry level of $f$ is the smallest integer $r$ such that
+$f$ is $r$-symmetric.
 
@@
 
-The term \textbf{canalization} was first used by
-Conrad Waddington (a British geneticist)
-in 1942 \cite{Waddington-1942}
-to express the ability of a species
-to produce members with very similar observable features over a period
-of time even when there are significant changes in the environment
-during that period.
-Since that time, the term ``canalization'' has been used to describe
+The term \textbf{canalization}, first used by
+Waddington \cite{Waddington-1942}, is generally used to describe
 the stability of a biological system with changes
 in external conditions.
-In 1969, Kauffman \cite{Kauffman-1969} introduced a Boolean network model
+In 1969, Kauffman \cite{Kauffman-1969} introduced canalyzing Boolean functions
 to explain the stability of gene regulatory networks.
-Kauffman observed that the use of canalyzing Boolean
-functions in the model captured many observed properties of
-gene regulatory networks.
 The subclass of NCFs was
@@
 It is known that the class of NCFs coincides with that
-of unate cascade functions \cite{}
-Many researchers have studied mathematical properties of NCFs and
-pointed out the importance of NCFs in modeling biological
-phenomena (e.g., \cite{Layne-2011,
+of unate cascade Boolean functions \cite{Jarrah-etal-2007}.
+In the literature on computational learning theory,
+generalized NCFs are referred to as $1$-\textbf{decision lists} \cite{KV-1994}.
+Many researchers have pointed out the importance of NCFs 
+in modeling biological phenomena 
+(e.g., \cite{Layne-2011,
 Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}).

### 2019-01-31 preliminaries.tex
ssr

@@
 
-We begin with a two-part observation from \cite{Stearns-etal-2018}.
+We begin with an observation from \cite{Stearns-etal-2018}.
 
@@
 \begin{tabular}{ll}
-\hspace*{1.1in} & $x_1:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
-\hspace*{1.1in} & $x_2:~$  $0 ~\longrightarrow~ 1$ \\ [0.5ex]
+\hspace*{1.1in} & $x_1:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
+\hspace*{1.1in} & $x_2:~$  $0 ~\longrightarrow~ 0$ \\ [0.5ex]
 \hspace*{1.1in} & $x_3:~$  $0 ~\longrightarrow~ 0$ \\ [0.5ex]
-\hspace*{1.1in} & $x_4:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
+\hspace*{1.1in} & $x_4:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
 \hspace*{1.1in} & $x_5:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
@@
 \noindent
-This function has 4 layers, with 
-Layer~1 consisting of the lines with $x_1$ and $x_2$, 
-Layer~2 consisting of the lines with $x_3$ and $x_4$, 
-Layer~3 consisting of the line with $x_5$ and 
-Layer~4 consisting of the line with $x_6$. \qed
+This function has 3 layers, with 
+Layer~1 consisting of the lines with $x_1$, $x_2$ and $x_3$, 
+Layer~2 consisting of the lines with $x_4$ and $x_5$, and
+Layer~3 consisting of the line with $x_6$.  \qed
 
@@
 lexicographic ordering given by  
-$\langle x_{\pi(1)}, x_{\pi(2)}, x_{\pi(n)}\rangle$.
+$\langle x_{\pi(1)}, x_{\pi(2)}, \ldots, x_{\pi(n)}\rangle$.
 We say that the representation is {\bf normalized} if
@@
 
-\subsection{Complexity of Estimating the Symmetry Level of a i
+\subsection{Complexity of Estimating the Symmetry Level of a 
 Boolean Function}
@@
 how many of the inputs in each group have the value 1.
-The following result shows that given a Boolean function $f$
+We now show that given a Boolean function $f$
 in the form of a Boolean expression, it is \cnp-hard to 
@@
 algorithm for approximating the 
-symmetry level of a Boolean function $f$ to within the factor $\rho$.
+symmetry level of a Boolean function $f$ 
+specified as a Boolean expression to within the factor $\rho$.
 This result holds even when $f$ is a CNF expression.
@@
 \noindent
-\textbf{Proof:}~ Suppose there is an efficient algorithm \cala{} 
+\textbf{Proof:}~ For some $\rho \geq 1$, suppose there is an efficient algorithm \cala{} 
 that provides a $\rho$ approximation for the symmetry level
-of a Boolean function for some $\rho \geq 1$.
+of any Boolean function specified as a Boolean expression.
 We will show that \cala{} can be used to efficiently solve
@@
 not satisfiable; that is, for all inputs, the value of $f'$ is 0.
-Thus, all the variables in $X \cup Y \cup Z$ can be put in one
-symmetry group; regardless of the number of variables 
-which are 1 in the group, the value of $f'$ is always 0. 
+Thus, all the variables in $X \cup Y \cup Z$ can be included in one
+symmetry group.
+The value of $f'$ is 0 regardless of the number of variables 
+which are 1 in the group. 
 \hfill$\Box$
@@
 \textbf{Proof of Claim 2:}~ Suppose $f$ is satisfiable.
-We argue that for $i \neq j$, variables $y_i$ and $y_j$
-cannot be in the same symmetry group.
-To see this, consider the following assignment $\alpha$ to
-the variables of $f'$.
+We argue that if $i \neq j$, variables $y_i$ and $y_j$
+cannot be in the same symmetry group for the function $f'$.
+To see this, consider any two variables $y_i$ and $y_j$ with $i \neq j$,
+and construct an assignment $\alpha$ to
+the variables of $f'$ in the following manner.
 \begin{description}
@@
 the value of $f'$ under the assignment $\alpha$ is 0.
-Now, consider the assignment $\alpha'$ that is obtained by exchanging
+Now, consider the assignment $\alpha'$ that is obtained by interchanging
 the values of $y_i$ and $y_j$ in $\alpha$ without changing the
@@
 
-Now, suppose we give $f'$ as the input to Algorithm \cala.
+Now, suppose we execute Algorithm \cala{} on the function $f'$.
 If $f$ is \emph{not} satisfiable, then from Claim 1, 
@@
 On the other hand, if $f$ is satisfiable, by Claim 2, the symmetry
-level of $f'$ is at least $\rho+1$ and \cala{} should produce at least
+level of $f'$ is at least $\rho+1$; so, \cala{} will produce at least
 $\rho+1$ groups.
@@
 In contrast to the above result,
-we will show that the symmetry level of an NCF (or that of a
+we will show in the next section that the symmetry level 
+of an NCF (or that of a
 generalized NCF) can be computed efficiently.

### 2019-01-31 abstract.tex
ssr

@@
 has also been studied in the literature.
-%% We establish lower and upper bounds on the symmetry
-%% level of an NCF $f$ as functions of the number of layers
-%% in a standard respresentation of $f$.  
 We observe that, in general,
@@
 In contrast, we show that the symmetry level of an NCF $f$
-can be easily computed from a standard (and efficiently
-computable) representation of $f$.
+can be easily computed given a standard representation of $f$.
+We also that for any NCF with $n$ variables, the notion of
+strong asymmetry considered in the literature is equivalent to
+the property that the symmetry level of the function is $n$.

### 2019-01-31 intro.tex
ssr

@@
 
+Other forms of symmetry which can be more general than the permutations 
+corresponding to symmetry groups have been considered \cite{KS-2000}.
+We say that a Boolean function $f(x_1, x_2, \ldots, x_n)$
+is \textbf{strongly asymmetric}
+if for any permutation $\pi$ of $\{1, 2, \ldots, n\}$
+\emph{except} the identity permutation,
+there exists an input $(a_1, a_2, \ldots,  a_n)$
+to $f$ such that $f(a_1, a_2, \ldots, a_n)$ $\neq$
+$f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
+In general, there are functions with $n$ variables that are properly
+$n$-symmetric, but not strongly asymmetric \cite{KS-2000}.
+However, in the case of NCFs with $n$ variables, we show that the notion of 
+strong asymmetry coincides with that of being properly $n$-symmetric.
+
 %\noindent

### 2019-01-31 symmetry_properties.tex
ssr

@@
 \label{sse:res_overview}
+Our first result (Proposition~\ref{pro:ncf_symmetric})
+identifies all the functions that are symmetric and
+canalyzing as well as those that are symmetric and NCFs.
+Our next result (Theorem~\ref{thm:ncf_symmetric_variables})
+provides a characterization of pairs of variables of an NCF
+that are part of the same symmetry group. 
+This characterization allows us to give a simple closed form
+expression (Theorem~\ref{thm:ncf_r_symmetric}) 
+for the symmetry level of an NCF specified using
+its normalized representation.
 
@@
 
-\begin{theorem}\label{thm:ncf_symmetric}
+\begin{proposition}\label{pro:ncf_symmetric}
 The only symmetric canalyzing functions are OR, AND, NOR, NAND, 
@@
 The only symmetric NCFs are OR, AND, NOR, and NAND.
-\end{theorem}
+\end{proposition}
 \noindent
@@
 \subsection{Symmetric Pairs of Variables in an NCF: A Characterization
-and Its Applictions}
+and Its Applications}
 \label{sse:ncf_strong_sym}
@@
 
+\iffalse
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Generalized forms of symmetry have been considered \cite{KS-2000}
@@
 $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\fi
+
+The notion of strong asymmetry of Boolean functions was defined 
+in Section~\ref{sec:prelim}.
 In general, there are functions with $n$ variables that are properly
 $n$-symmetric, but not strongly asymmetric \cite{KS-2000}.
-
 We now show that any $n$ variable NCF that is properly $n$-symmetric
 is also strongly asymmetric.
-As a consequence, there are NCFs with $n$ variables
+As a consequence, we point out there are NCFs with $n$ variables
 which are properly $n$-symmetric.

### 2019-02-01 abstract.tex
ssr

@@
 Boolean functions. 
-We identify all the Boolean functions that NCFs as well as symmetric. 
 A more general form of symmetric Boolean functions,
@@
 can be easily computed given a standard representation of $f$.
-We also that for any NCF with $n$ variables, the notion of
+We also show that for any NCF $f$ with $n$ variables, the notion of
 strong asymmetry considered in the literature is equivalent to
-the property that the symmetry level of the function is $n$.
+the property that the symmetry level of $f$ is $n$.
+We also identify all the Boolean functions that NCFs 
+as well as symmetric.

### 2019-02-01 intro.tex
ssr

@@
 This function is canalyzing since if we set to $x_1 = 1$,
-$f(1, x_2, x_3) ~=~ 0$, no matter which combination of values is assigned 
-to $x_2$ and $x_3$. \qed
+$f(1, x_2, x_3) ~=~ 0$, for all combinations of values for 
+$x_2$ and $x_3$. \qed
 
@@
 \textbf{symmetry groups}, where the members of each group are pairwise symmetric.
-Function $f$ is said to be
+A Boolean function $f$ is 
 $r$-\textbf{symmetric} if it has at most $r$ symmetry groups.
@@
 it is $r$-symmetric, but not $(r-1)$-symmetric.
-For example, the function $f_3(x_1, x_2) = x_1 \vee\, \overline{x_2}$
-is not 1-symmetric since $f_3(1, 0) \neq f_3(0, 1)$; however, 
-it is trivially 2-symmetric.
+For example, the function $f_3(x_1, x_2, x_3) = (x_1 \wedge x_2) \vee\, \overline{x_3}$
+is not 1-symmetric since $f_3(1, 0, 1) \neq f_3(1, 1, 0)$; however, 
+it is 2-symmetric with the symmetric groups being $\{x_1, x_2\}$ and $\{x_3\}$.
 For a Boolean function $f$, the integer $r$ such that $f$ is properly $r$-symmetric
@@
 $f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
-Symmetric and $r$-symmetric functions for $r \geq 2$ have also been
-studied in the literature in the context of discrete
-dynamical systems (see, e.g., 
+The class of $r$-symmetric functions have also been
+studied in the literature on discrete dynamical systems (see, e.g., 
 \cite{Barrett-etal-2007,Rosenkrantz-etal-2015,MR-2007}).
@@
 corresponding to symmetry groups have been considered \cite{KS-2000}.
-We say that a Boolean function $f(x_1, x_2, \ldots, x_n)$
+A Boolean function $f(x_1, x_2, \ldots, x_n)$
 is \textbf{strongly asymmetric}
@@
 $n$-symmetric, but not strongly asymmetric \cite{KS-2000}.
-However, in the case of NCFs with $n$ variables, we show that the notion of 
-strong asymmetry coincides with that of being properly $n$-symmetric.
+However, in the case of NCFs with $n$ variables, we will show 
+(see Section~\ref{sec:ncf_and_symmetry}) that the notion of 
+strong asymmetry coincides with that 
+of being properly $n$-symmetric.
 
@@
 
-Our focus is on the relationship between NCFs and symmetric
-Boolean functions.
+In this note, our focus is on the relationship between 
+NCFs and symmetric Boolean functions.
+We observe that, in general,
+one cannot efficiently approximate the symmetry level of
+a Boolean function to within any factor, unless \textbf{P} = \cnp.
+In contrast, we show that the symmetry level of an NCF $f$
+can be easily computed given a standard representation of $f$.
+We also show that for any NCF $f$ with $n$ variables, the notion of
+strong asymmetry considered in the literature is equivalent to
+the property that the symmetry level of $f$ is $n$.
+Finally, we identify all the Boolean functions that NCFs 
+as well as symmetric.
 
@@
 
-The term \textbf{canalization}, first used by
+The term \textbf{canalization}, coined by
 Waddington \cite{Waddington-1942}, is generally used to describe
@@
 generalized NCFs are referred to as $1$-\textbf{decision lists} \cite{KV-1994}.
-Many researchers have pointed out the importance of NCFs 
+Many researchers have pointed out the usefulness of NCFs 
 in modeling biological phenomena 
@@
 Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}).
-Additionally, properties of NCFs such as sensitivity and 
-stability have also been studied
-by many researchers 
-\cite{Kauffman-etal-2004, Layne-2011,Layne-etal-2012,
-Li-etal-2011,Li-etal-2013,Klotz-etal-2013,
-Stearns-etal-2018}. 
+Properties of NCFs such as sensitivity and 
+stability have also been studied in the 
+literature \cite{Kauffman-etal-2004, Layne-2011,Layne-etal-2012,
+Li-etal-2011,Li-etal-2013,Klotz-etal-2013, Stearns-etal-2018}. 
 To our knowledge, the relationship between NCFs and symmetric
-Boolean functions has not been studied in the literature.
+Boolean functions has not been addressed in the literature.

### 2019-02-01 preliminaries.tex
ssr

@@
 
-Here, we present some definitions and preliminary results which enable
-us to define a normalized representation for NCFs.
+We present some definitions and recall
+a known result regarding NCFs which allow us
+to define a normalized representation for NCFs.
 This representation plays an important role in several
 problems considered in later sections.
+We also show that, in general, the symmetry level
+of a Boolean function cannot be approximated to within any
+factor $\rho \geq 1$, unless \textbf{P} = \cnp.
 
 \medskip
+
+\subsection{Layers and Normalized Representation}
+\label{sse:ncf_layer}
 
@@
 
-Throughout this section, we assume that an NCF $f$ with $n$
+Throughout this paper, we assume that an NCF $f$ with $n$
 variables is specified using the simplified representation with $n$ lines. 
@@
 
-\subsection{Layers and Normalized Representation}
-\label{sse:ncf_layer}
-
-We begin with an observation from \cite{Stearns-etal-2018}.
+The following observation is from \cite{Stearns-etal-2018}.
 
@@
 For our purposes, we use the following definition based on the
-simplified representation of NCFs given in Section~\ref{sec:intro}.
+simplified representation of NCFs.
 
@@
 \textbf{Proof:}~ For some $\rho \geq 1$, suppose there is an efficient algorithm \cala{} 
-that provides a $\rho$ approximation for the symmetry level
-of any Boolean function specified as a Boolean expression.
+that approximates the symmetry level of a Boolean function
+specified as Boolean expression to within the factor $\rho$.
+Without loss of generality, we can assume that $\rho$ is an integer.
 We will show that \cala{} can be used to efficiently solve
@@
 Let $Y = \{y_1, y_2, \ldots, y_{\rho+1}\}$ and
-$Z = \{z_1, z_2, \ldots, z_{\rho+1}\}$ be two new sets of variables.
+$Z = \{z_1, z_2, \ldots, z_{\rho+1}\}$ be two 
+new sets of variables, with each set containing $\rho+1$ variables.
 The expression for $f'$ is given by
@@
 symmetry group.
-The value of $f'$ is 0 regardless of the number of variables 
-which are 1 in the group. 
-\hfill$\Box$
+The value of $f'$ is 0 regardless of how many variables in the
+group have the value 1.  \hfill$\Box$
 
@@
 different values for $f'$.
-Therefore, if $f$ is not satisfiable, for $i \neq j$, $y_i$ and $y_j$ cannot
+Therefore, if $f$ is satisfiable, for $i \neq j$, $y_i$ and $y_j$ cannot
 be in the same symmetry group.

### 2019-02-01 symmetry_properties.tex
ssr

@@
 \section{Symmetry of Nested Canalyzing Functions}
-\label{sse:ncf_and_symmetry}
+\label{sec:ncf_and_symmetry}
 
@@
 \label{sse:res_overview}
-Our first result (Proposition~\ref{pro:ncf_symmetric})
-identifies all the functions that are symmetric and
-canalyzing as well as those that are symmetric and NCFs.
-Our next result (Theorem~\ref{thm:ncf_symmetric_variables})
+Our first result (Theorem~\ref{thm:ncf_symmetric_variables})
 provides a characterization of pairs of variables of an NCF
@@
 its normalized representation.
-
-\subsection{Symmetric Canalyzing and Nested Canalyzing Functions}
-\label{sse:sym_and_cf_ncf}
-
-\begin{proposition}\label{pro:ncf_symmetric}
-The only symmetric canalyzing functions are OR, AND, NOR, NAND, 
-the constant function 0, and the constant function 1.
-
-The only symmetric NCFs are OR, AND, NOR, and NAND.
-\end{proposition}
-\noindent
-\textbf{Proof:}~
-Suppose symmetric function $f$ is a canalyzing function. Then there
-is a variable $x$, and values $a$ and $b$ such that whenever $x =
-a$, function $f$ has value $b$.  Since $f$ is symmetric, $f$  has
-value $b$ whenever any of its variables have value $a$.  Thus, if
-at least one of its variables has value $a$, then $f$ has value
-$b$.  If $f$ has value $\bar{b}$ when none of its variables has
-value $a$, the four possible combinations of values for $a$ and $b$
-correspond to the four functions OR, AND, NOR, and NAND.  If $f$
-has value $b$ when none of its variables has value $a$, then $f$
-is the constant function $b$.
-
-Suppose that $f$ is a NCF.  Since constant functions are not NCFs,
-$f$ has value $\bar{b}$ when none of its variables has value $a$.
-The four possible combinations of values for $a$ and $b$ correspond
-to the four functions OR, AND, NOR, and NAND, each of which is a
-NCF.  \QED
+Our next result (Theorem~\ref{thm:ncf_strong_asymmetry})
+shows that for any NCF $f$ with $n$ variables,
+the two statements ``$f$ is strongly asymmetric"
+``the symmetry level of $f$ is $n$" are equivalent. 
+Our final result (Proposition~\ref{pro:ncf_symmetric})
+identifies all the functions that are symmetric and
+canalyzing as well as those that are symmetric and NCFs.
 
@@
 \begin{theorem}\label{thm:ncf_symmetric_variables}
-A pair of variables of a NCF are symmetric iff
-in the normalized representation of the NCF
+Two variables of an NCF $f$ are symmetric iff
+in the normalized representation of $f$,
 they occur in the same layer and have the same canalyzing value.
@@
 \textbf{Proof:}~
-Suppose that a pair of variables occur in the same layer of a
-canalyzing representation, and the lines containing these variables
+Suppose that two variables of an NCF $f$ occur in the same layer of a
+normalized representation, and the lines containing these variables
 have the same canalyzing value.  Then in any input assignment, the
@@
 
-Now, let $f$ be the normalized NCF representation for the function.
-For purposes of notational simplicity, wlog we assume that the
+For the converse, consider the normalized NCF representation of $f$. 
+For purposes of notational simplicity, we assume without loss of generality that the
 canalyzing variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.
-Let  $x_j$ and $x_k$ be a pair of variables that occur in different
+Let  $x_j$ and $x_k$ be two variables that occur in different
 layers of $f$, or occur in the same layer, but with different
-canalyzing values.  Wlog, assume that $j < k$, so that the line for
+canalyzing values.  Without loss of generality, 
+assume that $j < k$, so that the line for
 $x_j$ occurs above the line for $x_k$.
-
 Suppose that line $i$ of $f$, $1 \leq i \leq n$, is
 
+\smallskip
+
+\hspace*{1in}
+$x_i : a_i ~\longrightarrow~ b_i$.
+
+\smallskip
+
 \noindent
-$(x_i : a_i \rightarrow b_i)$. \\
 Consider the following input assignment $\alpha = (c_1, c_2, \ldots,  c_n)$ 
 to the variables of $f$.
-
-\noindent
-For $1 \leq i < j$, set $c_i = \overline{a_i}$. \\
-For $i = j$, set $c_j =a_j$. \\
-For $i = k$, set $c_k =\overline{a_j}$. \\
-For $j < i < k$, and for  $k < i \leq n$, 
+\begin{description}
+\item{(i)} For $1 \leq i < j$, set $c_i = \overline{a_i}$. 
+\item{(ii)} For $i = j$, set $c_j =a_j$. 
+\item{(iii)} For $i = k$, set $c_k =\overline{a_j}$. 
+\item{(iv)} For $j < i < k$, and for  $k < i \leq n$, 
 if $b_i = b_j$, then set $c_i =  \overline{a_i}$,
 else set $c_i =  a_i$.
-
+\end{description}
 Note that $f(\alpha) = b_j$, since the variables in all lines above
@@
 
-Now let $\alpha'$ be the assignment that is obtained from $\alpha$
+\medskip
+
+Now, let $\alpha'$ be the assignment that is obtained from $\alpha$
 by interchanging the values of variables $x_j$ and $x_k$, so that
@@
 
+\medskip
+
 \noindent
-{\bf Case 1:} $(x_k : a_j \rightarrow b_j)$. \\
+{\bf Case 1:} Line $k$ has the form $x_k : a_j ~\longrightarrow~ b_j$.
+
+\smallskip
 
@@
 
+\medskip
+
 \noindent
-{\bf Case 2:} $(x_k : a_j \rightarrow \overline{b_j})$. \\
+{\bf Case 2:} Line $k$ has the form  $x_k : a_j ~\longrightarrow~ \overline{b_j}$.
 
-Let line
-$q$ be the first line such that $j < q \leq k$ and the canalyzed
+\smallskip
+
+Let line $q$ be the first line such that $j < q \leq k$ and the canalyzed
 output of line $q$ is $\overline{b_j}$.  Note that $q$ might possibly
@@
 
+\medskip
+
 \noindent
-{\bf Case 3:} $(x_k : \overline{a_j} \rightarrow b_j)$. \\
+{\bf Case 3:} Line $k$ has the form $x_k : \overline{a_j} ~\longrightarrow~ b_j$.
+
+\smallskip
 
@@
 
+\smallskip
+
 Now suppose there is no line below line $j$ for which the canalyzed
@@
 
+\medskip
+
 \noindent
-{\bf Case 4:} $(x_k : \overline{a_j} \rightarrow \overline{b_j})$.  \\ 
+{\bf Case 4:} Line $k$ has the form  $x_k : \overline{a_j} ~\longrightarrow~ \overline{b_j}$.
+
+\smallskip
+
 Suppose there is a line, other than line $k$, below line $j$ for
@@
 \end{theorem}
+
 \noindent
@@
 From Theorem~\ref{thm:ncf_symmetric_variables}, 
-any pair of variables occurring in the same layer, with the same canalyzing value,
-are symmetric.
+any pair of variables occurring in the same layer, 
+with the same canalyzing value, are symmetric.
 Thus the function is at most $(r_1 + 2 r_2)$-symmetric.
@@
 \begin{corollary}\label{cor:ncf_r_symmetric_layers}
-A NCF with a normalized nested canalyzing representation consisting
+(i) An NCF with a normalized nested canalyzing representation consisting
 of $q$ layers is $2q$-symmetric, and is not $(q-1)$-symmetric.
-
-An $r$-symmetric NCF has a normalized nested canalyzing representation
+(ii) An $r$-symmetric NCF has a normalized nested canalyzing representation
 with at most $r-1$ layers, and has an unnormalized nested canalyzing
-representation with at most $r$ layers.  \end{corollary}
+representation with at most $r$ layers.  
+\end{corollary}
 
@@
 \label{sse:strong_asym_ncf}
-
-\iffalse
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Generalized forms of symmetry have been considered \cite{KS-2000}
-in which a Boolean function is symmetric with respect to some permutation,
-which can be more general than the permutations corresponding to symmetry groups.
-We say that a Boolean function $f(x_1, x_2, \ldots, x_n)$ 
-is \textbf{strongly asymmetric}
-if for any permutation $\pi$ of $\{1, 2, \ldots, n\}$
-\emph{except} the identity permutation,
-there exists an input $(a_1, a_2, \ldots,  a_n)$ 
-to $f$ such that $f(a_1, a_2, \ldots, a_n)$ $\neq$  
-$f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\fi
 
@@
 is also strongly asymmetric.
-As a consequence, we point out there are NCFs with $n$ variables
-which are properly $n$-symmetric.
+As a consequence, we observe that there are NCFs with $n$ variables
+that are properly $n$-symmetric.
 
 \begin{theorem}\label{thm:ncf_strong_asymmetry}
-A NCF with $n$ variables is strongly asymmetric iff
+An NCF with $n$ variables is strongly asymmetric iff
 it is properly $n$-symmetric.
@@
 \textbf{Proof:}~ 
-If a NCF is $(n-1)$-symmetric, then consider a permutation than
+If an NCF is $(n-1)$-symmetric, then consider a permutation than
 interchanges two variables from a symmetry group with at least two
@@
 
-Now consider a $n$ variable NCF that is properly $n$-symmetric.
+For the converse, consider an $n$ variable NCF that is properly $n$-symmetric.
 Let $f$ be the normalized NCF representation for the function.  For
-purposes of notational simplicity, wlog we assume that the canalyzing
+purposes of notational simplicity, without loss of generality,
+we assume that the canalyzing
 variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.  Let $\pi$
@@
 $\pi$ is a permutation, $k < n$.  Overall, for $1 \leq i < k$,
-$\pi(i) = i$;
- $\pi(k) > k$; and for $k < i \leq n$, $\pi(i) \geq k$.
+$\pi(i) = i$,~ $\pi(k) > k$,~ and for $k < i \leq n$,~ $\pi(i) \geq k$.
 
-Suppose that line $i$ of $f$, $1 \leq i \leq n$, is \noindent $(x_i
-: a_i \rightarrow b_i)$. \\ Assignment $(c_1, c_2, \ldots,  c_n)$
-is constructed as follows.
+Suppose that line $i$ of $f$, $1 \leq i \leq n$, is
+
+\smallskip
+
+\hspace*{1in} $x_i : a_i ~\longrightarrow~ b_i$.
+
+\smallskip
 
 \noindent
-For $1 \leq i < k$, set $c_i = \overline{a_i}$. \\ For $i = k$, set
-$c_k =a_k$. \\ For $i > k$, let $i' = \pi^{-1}(i)$, so that $\pi(i')
+Assignment $(c_1, c_2, \ldots,  c_n)$ is constructed as follows.
+\begin{description}
+\item{(i)} For $1 \leq i < k$, set $c_i = \overline{a_i}$.  
+\item{(ii)} For $i = k$, set $c_k =a_k$. 
+\item{(iii)} For $i > k$, let $i' = \pi^{-1}(i)$, so that $\pi(i')
 = i$.  Thus, after the permutation of values, variable $x_{i'}$
@@
 \overline{a_{i'}}$, else set $c_i =  a_{i'}$.
-
+\end{description}
 Since $c_k$ matches the canalyzing value of line $k$ of $f$, and
@@
 
+\medskip
+
 \noindent
-{\bf Case 1:} $(x_{k'} : a_k \rightarrow b_k)$. \\
+{\bf Case 1:} Line $k'$ has the form  $x_{k'} : a_k ~\longrightarrow~ b_k$.
+
+\smallskip
 
@@
 
+\medskip
+
 \noindent
-{\bf Case 2:} $(x_{k'} : a_k \rightarrow \overline{b_k})$. \\
+{\bf Case 2:} Line $k'$ has the form $x_{k'} : a_k ~\longrightarrow~ \overline{b_k}$.
+
+\smallskip
 
@@
 
+\medskip
+
 \noindent
-{\bf Case 3:} $(x_{k'} : \overline{a_k} \rightarrow b_k)$. \\
+{\bf Case 3:} Line $k'$ has the form $x_{k'} : \overline{a_k} ~\longrightarrow~ b_k$.
+
+\smallskip
 
@@
 
+\medskip
+
 \noindent
-{\bf Case 4:} $(x_{k'} : \overline{a_k} \rightarrow \overline{b_k})$. \\
+{\bf Case 4:} Line $k'$ has the form $x_{k'} : \overline{a_k} ~\longrightarrow~ 
+              \overline{b_k}$.
+
+\smallskip
 
@@
 
+\medskip
+
 Examples of properly $n$-symmetric $n$-variable NCFs satisfying the
@@
 For each $n > 1$, let $f_n$ be the function defined by the following
-formula.
+formula:
 
-$$x_1 \vee \bar{x}_2 ( x_3 \vee \bar{x}_4 (\cdots  ) )$$
+\smallskip
 
+\hspace*{1in} $x_1 \vee (\overline{x}_2 \wedge ( x_3 \vee (\overline{x}_4 \wedge (\cdots ))$
+
+\smallskip
+
+\noindent
 For instance,
-$$f_6 = x_1 \vee \bar{x}_2 ( x_3 \vee \bar{x}_4 ( x_5 \vee \bar{x}_6  ) ),$$
-and
-$$f_7 = x_1 \vee \bar{x}_2 ( x_3 \vee \bar{x}_4 ( x_5 \vee \bar{x}_6   x_7) ).$$
-
+\[
+f_6 = x_1 \vee (\overline{x}_2 \wedge (x_3 \vee (\overline{x}_4 \wedge (x_5 \vee \overline{x}_6))
+~~~~\mathrm{and}~~~~
+f_7 = x_1 \vee (\overline{x}_2 \wedge (x_3 \vee (\overline{x}_4 \wedge (x_5 
+          \vee (\overline{x}_6 \wedge  x_7)).
+\]
 Function $f_n(x_1, x_2, \ldots, x_n)$ is the NCF corresponding to
@@
 
-\medskip
+\smallskip
+
 \noindent
 \hspace*{0.5in}
-$x_1:~ (1 \rightarrow 1)$ \\
+$x_1:~ 1 ~\longrightarrow~ 1$ \\
 \hspace*{0.5in}
-$x_2:~ (1 \rightarrow 0)$ \\
+$x_2:~ 1 ~\longrightarrow~ 0$ \\
 \hspace*{0.5in}
-$x_3:~ (1 \rightarrow 1)$ \\
+$x_3:~ 1 ~\longrightarrow~ 1$ \\
 \hspace*{0.75in}
@@
 \hspace*{0.5in}
-$x_n:~ (1 \rightarrow 1)$
+$x_n:~ 1 ~\longrightarrow~ 1$
 
@@
 \hspace*{0.5in}
-$x_n:~ (1 \rightarrow 0)$. \\
+$x_n:~ 1 ~\longrightarrow~ 0$.
 
+\smallskip
+
+\noindent
 Note that if $x_i = 0$ for all $i$, $1 \leq i \leq n$, then the
@@
 
-\iffalse
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%\begin{theorem}\label{thm:ncf_asymmetry}
-\begin{theorem}\label{thm:ncf_asymmetry}.
-Consider the set of variables $x_i$ for $1 \leq i \leq n$ and 
-let $f(x_1, x_2, \ldots, x_n)$ be the NCF defined by the following sequence
-of lines:
+\subsection{Symmetric Canalyzing and Nested Canalyzing Functions}
+\label{sse:sym_and_cf_ncf}
 
-\medskip
+\begin{proposition}\label{pro:ncf_symmetric}
+(i) The only symmetric canalyzing functions are OR, AND, NOR, NAND, 
+the constant function 0 and the constant function 1.
+(ii) The only symmetric NCFs are OR, AND, NOR and NAND.
+\end{proposition}
+%%\medskip
 \noindent
-\hspace*{0.5in}
-$x_1:~ (1 \rightarrow 1)$ \\
-\hspace*{0.5in}
-$x_2:~ (1 \rightarrow 0)$ \\
-\hspace*{0.5in}
-$x_3:~ (1 \rightarrow 1)$ \\
-\hspace*{0.75in}
-$\vdots$
+\textbf{Proof:}~
+
+\smallskip
 
 \noindent
-If $n$ is odd, the last line is
+\textbf{Part (i):}~
+Suppose symmetric function $f$ is also a canalyzing function. 
+Then there is a variable $x$, and values $a$ and $b$ 
+such that whenever $x = a$, function $f$ has value $b$.  
+Since $f$ is symmetric, $f$  has
+value $b$ whenever any of its variables has value $a$.  
+Thus, if at least one of its variables has value $a$, then $f$ has value $b$.  
+If $f$ has value $\overline{b}$ when none of its variables has
+value $a$, the four possible combinations of values for $a$ and $b$
+correspond to the four functions OR, AND, NOR, and NAND.  If $f$
+has value $b$ when none of its variables has value $a$, then $f$
+is the constant function $b$.
+
+\smallskip
 
 \noindent
-\hspace*{0.5in}
-$x_n:~ (1 \rightarrow 1)$
-
-\noindent
-and if $n$ is even, the last line is
-
-\noindent
-\hspace*{0.5in}
-$x_n:~ (1 \rightarrow 0)$. \\
-
-\noindent
-Let $\pi$ be any permutation of $\{1, 2, \ldots, n\}$
-\emph{except} the identity permutation.
-Then there exists an input $(a_1, a_2, \ldots,  a_n)$ 
-to $f$ such that $f(a_1, a_2, \ldots, ,a_n)$ $\neq$  
-$f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
-\end{theorem}
-
-\noindent
-\textbf{Proof:}~ 
-Note that if $x_i = 0$ for all $i$, $1 \leq i \leq n$, then the
-value of the function $f$ is 0 if $n$ is odd and 1 if $n$ is even.
-
-First assume there is an odd $i$ such that $\pi(i)$ is even. 
-Consider the input where $a_i$ is 1 and all others are 0. 
-It can be verified that $f(a_1, a_2, \ldots, a_n)$ = 1 while
-$f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$ = 0.
-
-\smallskip
-Now assume the permutation $\pi$ is such that for each odd $i$, $\pi(i)$ is odd 
-and for each even $i$, $\pi(i)$ is even.
-Let $k$ be the smallest index such that $k \neq \pi(k)$.
-Assume that $k$  is odd. 
-Construct the input $(a_1, a_2, \ldots,  a_n)$ to $f$ as follows: 
-\begin{enumerate}
-\item Set $a_k$ to 1 and $a_i$ to 0 for all other odd $i$. 
-\item For even $i$, set $a_i$ to 0 if $i < k$ and to 1 otherwise. 
-\end{enumerate}
-We have the following claim.
-
-\smallskip
-\noindent
-\textbf{Claim:}~ For the input $(a_1, a_2, \ldots, a_n)$, 
-$f(a_1, a_2, \ldots, a_n)$ = 1 and
-$f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$ = 0.
-
-\smallskip
-\noindent
-\textbf{Proof of claim:}~ Note that the permutation
-on the even indices maps 1's to 1's and leaves 0's in place.  
-Before the permutation,  
-$x_j = 0$ for $1 \leq j < k$ and $x_k = 1$,
-so the canalyzing variable is $x_k$ and the function value is 1.
-After the permutation, 
-$x_j = a_j = 0$ for $1 \leq j < k$,
-$x_k = 0$, and $x_{k+1} = 1$;
-so the canalyzing variable is $x_{k+1}$ and the function value is 0.
-This establishes the claim.
-
-A similar proof can be given when $k$ is even and this completes
-the proof of Theorem~\ref{thm:ncf_asymmetry}.
-\QED
-
-\noindent
-The following result is a consequence of the above theorem.
-
-\begin{corollary}\label{cor:ncf_not_rsymm}
-For every $n \geq 2$, there is an $n$-variable NCF that is not $n-1$ symmetric.
-\end{corollary}
-
-\noindent
-\textbf{Proof:}~ 
-Let $f$ be the Boolean function defined in Theorem~\ref{thm:ncf_asymmetry}.
-We will show by contradiction that $f$ is not $r$-symmetric for any $r < n$.
-
-Suppose $f$ is $r$-symmetric for some $r < n$.
-Then, by definition, there is a partition of the set $X$ =
-$\{x_1, x_2, \ldots, x_n\}$ of variables into at most $r$ subsets such that
-the value of $f$ depends only on the number of variables with value 1
-in each subset.
-Since $r < n$, this partition has a subset, say $X_j$, 
-such that $|X_j| \geq 2$.
-Let $X_j = \{x_{j_1}, x_{j_2}, \ldots, x_{j_k}\}$, where $k \geq 2$.
-Assume without loss of generality that $j_1 < j_2 < \ldots < j_k$.
-Define a permutation $\pi$ of $\{1, 2, \ldots, n\}$ as follows. 
-\begin{enumerate}
-\item For all $i$ such that $x_i \not\in X_j$, $\pi(i) = i$.
-\item $\pi(j_{\ell})$ = $j_{\ell+1}$ for $1 \leq \ell_t \leq k-1$
-      and $\pi(j_{k})$ = $j_1$. 
-\end{enumerate}
-Note that $\pi$ is not the identity permutation.
-Thus, by Theorem~\ref{thm:ncf_asymmetry},
-there is an input $(a_1, a_2, \ldots, a_n)$ to $f$ such that
-such that
-\begin{equation}\label{eqn:f_asymmetry}
-f(a_1, a_2, \ldots, a_n) ~\neq~  f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})
-\end{equation}
-However, from the definition of permutation $\pi$, it follows
-that the number of 1's in any subset of the partition 
-is the same in both the inputs
-$(a_1, a_2, \ldots, a_n)$ and
-$(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
-Therefore, since $f$ is $r$-symmetric, we must have
-\begin{equation}\label{eqn:f_contra}
-f(a_1, a_2, \ldots, a_n) ~=~ f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)}).
-\end{equation}
-Equation~(\ref{eqn:f_contra}) contradicts 
-Equation~(\ref{eqn:f_asymmetry}), and the result follows. \QED
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\fi
-
-\iffalse
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-$$x_1 \vee \bar{x}_2 ( x_3 \vee \bar{x}_4 (\cdots  ) )$$
-
-For instance,
-$$f_6 = x_1 \vee \bar{x}_2 ( x_3 \vee \bar{x}_4 ( x_5 \vee \bar{x}_6  ) ),$$
-and
-$$f_7 = x_1 \vee \bar{x}_2 ( x_3 \vee \bar{x}_4 ( x_5 \vee \bar{x}_6   x_7) ).$$
-
-Function $f_n$ is a NCF, 
-and can be representing as a sequence where the variables appear in lexicographic order,
-each $a_i$ is 1, and the $b_i$'s alternate between 1 and 0.
-For instance, $f_6$ can be represented as follows:
-
-\noindent
-$(x_1: 1 \rightarrow 1 )$ \\
-$(x_2: 1 \rightarrow 0 )$ \\
-$(x_3: 1 \rightarrow 1 )$ \\
-$(x_4: 1 \rightarrow 0 )$ \\
-$(x_5: 1 \rightarrow 1 )$ \\
-$(x_6: 1 \rightarrow 0 )$ \\
-
-We now argue that $f_n$ is not $n-1$ symmetric.
-Consider the classes of $f_n$, represented as an $r$-symmetric function for some $r$.
-If we compare the assignment where a given odd-numbered variable $x_{2i+1}$ is the only variable equal to 1,
-and the assignment where a given even-numbered variable $x_{2j}$ is the only variable equal to 1,
-we can see that these assignments produce different values for $f_n$,
-so variables $x_{2i+1}$ and $x_{2j}$ must belong to different classes.
-If we compare the assignment where an odd-numbered variable $x_{2i+1}$  
-and the even-numbered variable $x_{2i+2}$ 
-are the only variables equal to 1,
-and the assignment where an odd-numbered variable of the form $x_{2k+1}$, $k >i$,
-and $x_{2i+2}$  are the only variables equal to 1,
-we can see that these assignments produce different values for $f_n$,
-so $x_{2i+1}$ and $x_{2k+1}$ must belong to different classes.
-Thus, each odd-numbered variable is the only member of its class.
-If we compare the assignment where an even-numbered variable $x_{2i}$  
-and the odd-numbered variable $x_{2i+1}$ 
-are the only variables equal to 1,
-and the assignment where an even-numbered variable of the form $x_{2k}$, $k >i$,
-and $x_{2i+1}$  are the only variables equal to 1,
-we can see that these assignments produce different values for $f_n$,
-so $x_{2i}$ and $x_{2k}$ must belong to different classes.
-Thus, each even-numbered variable is the only member of its class.
-\QED
-
-%%The following result shows another form of asymmetry for the function
-%%defined in the above theorem.  (This form of asymmetry may be stronger
-%%than the one established in the above theorem.)
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\fi
+\textbf{Part (ii):}~
+Suppose that symmetric function $f$ is also an NCF.  
+Since constant functions are not NCFs,
+$f$ has value $\overline{b}$ when none of its variables has value $a$.
+The four possible combinations of values for $a$ and $b$ correspond
+to the four functions OR, AND, NOR and NAND, each of which is an
+NCF.  \QED

### 2019-02-06 abstract.tex
ssr

@@
 Boolean functions. 
-A more general form of symmetric Boolean functions,
+A general form of symmetric Boolean functions,
 called $r$-symmetric functions (where $r$ is the symmetry level),
-has also been studied in the literature.
+has been studied in the literature.
 We observe that, in general,
@@
 the property that the symmetry level of $f$ is $n$.
-We also identify all the Boolean functions that NCFs 
-as well as symmetric.
+In addition, we identify all the Boolean functions that are NCFs 
+and symmetric.

### 2019-02-06 preliminaries.tex
ssr

@@
 Suppose lines $n-1$ and $n$ in the specification of $f$ 
-have complementary canalyzed outputs.
+have complementary canalyzed values.
 Then, the function remains unchanged 
-if the canalyzing value and canalyzed output in line $n$
+if the canalyzing value and canalyzed value in line $n$
 are both complemented. 
@@
 A {\bf layer} of a NCF representation is a maximal length sequence 
-of lines with the same canalyzed output.
+of lines with the same canalyzed value.
 \end{definition}
@@
 Layer~2 consisting of the lines with $x_4$ and $x_5$, and
-Layer~3 consisting of the line with $x_6$.  \qed
+Layer~3 consisting of the line with $x_6$.  %\qed
 
@@
 \begin{enumerate}
-  \item Lines $n-1$ and $n$ have the same canalyzed output ~and
+  \item Lines $n-1$ and $n$ have the same canalyzed value ~and
   \item each pair of consecutive lines
-        with the same canalyzed output occur in the lexicographic order 
+        with the same canalyzed value occur in the lexicographic order 
         of their canalyzing variables; that is, if two consecutive lines are \\
@@
 since the line with $x_2$ precedes the one 
-with $x_1$. \qed
+with $x_1$. %\qed
 
@@
 we note that in polynomial time, for any NCF representation with more than one line,
-we can first modify the last line if necessary so that it has the same canalyzed output as the preceding line,
+we can first modify the last line if necessary 
+so that it has the same canalyzed value as the preceding line,
 and then we can lexicographically sort
-each maximal sequence of consecutive lines with the same canalyzed output,
+each maximal sequence of consecutive lines with the same canalyzed value,
 thereby obtaining an equivalent normalized NCF representation.

### 2019-02-06 symmetry_properties.tex
ssr

@@
 In line $k$ of $f$, canalyzing value $a_k$ is either the same or
-the complement of canalyzing value $a_j$, and canalyzed output $b_k$
-is either the same or the complement of canalyzed output $b_j$,
+the complement of canalyzing value $a_j$, and canalyzed value $b_k$
+is either the same or the complement of canalyzed value $b_j$,
 Thus, there are four possible cases for the form of line $k$.  We
@@
 Since line $k$ has
-the same canalyzing value and canalyzed output as line $j$, line
+the same canalyzing value and canalyzed value as line $j$, line
 $k$ must occur in a lower layer than that of line $j$.  Thus, there
 is at least one line between line $j$ and line $k$ for which the
-canalyzed output is $\overline{b_j}$.  Let line $q$ be the first
+canalyzed value is $\overline{b_j}$.  Let line $q$ be the first
 such line.  Note that for all $i$ such that $1 \leq i < q$, the
@@
 canalyzing value of line $i$, but the value of variable $x_q$ does
-match the canalyzing value of line $q$.  Since canalyzed output
+match the canalyzing value of line $q$.  Since canalyzed value
 $b_q$ equals $\overline{b_j}$, we have that $f(\alpha') =
@@
 Let line $q$ be the first line such that $j < q \leq k$ and the canalyzed
-output of line $q$ is $\overline{b_j}$.  Note that $q$ might possibly
+value of line $q$ is $\overline{b_j}$.  Note that $q$ might possibly
 equal $k$.  For all $i$ such that $1 \leq i < q$, the value of
@@
 value of line $i$, but the value of variable $x_q$ does match the
-canalyzing value of line $q$.  Since canalyzed output $b_q$ equals
+canalyzing value of line $q$.  Since canalyzed value $b_q$ equals
 $\overline{b_j}$, we have that $f(\alpha') = \overline{b_j}$.
@@
 
-Suppose there is a line below line $j$ for which the canalyzed output is
+Suppose there is a line below line $j$ for which the canalyzed value is
 $\overline{b_j}$.  Let $q$ be the first such line.  Note that for
@@
 $i$, but the value of variable $x_q$ does match the canalyzing value
-of line $q$.  Since canalyzed output $b_q$ equals $\overline{b_j}$,
+of line $q$.  Since canalyzed value $b_q$ equals $\overline{b_j}$,
 we have that $f(\alpha') = \overline{b_j}$.
@@
 Now suppose there is no line below line $j$ for which the canalyzed
-output is $\overline{b_j}$.  Then lines $j$ and $k$ both occur in
+value is $\overline{b_j}$.  Then lines $j$ and $k$ both occur in
 the last layer of $f$.  Thus, for every line $i$, $1 \leq i \leq
@@
 match the canalyzing value of line $i$.  Consequently, $f(\alpha')
-= \overline{b_j}$, the complement of the canalyzed output in the
+= \overline{b_j}$, the complement of the canalyzed value in the
 last layer of $f$.
@@
 Suppose there is a line, other than line $k$, below line $j$ for
-which the canalyzed output is $\overline{b_j}$.  Let $q$ be the
+which the canalyzed value is $\overline{b_j}$.  Let $q$ be the
 first such line.  Note that for all $i$ such that $1 \leq i < q$,
@@
 the canalyzing value of line $i$, but the value of variable $x_q$
-does match the canalyzing value of line $q$.  Since canalyzed output
+does match the canalyzing value of line $q$.  Since canalyzed value
 $b_q$ equals $\overline{b_j}$, we have that $f(\alpha') =
@@
 Now suppose that line $k$ is the only line below line $j$ for which
-the canalyzed output is $\overline{b_j}$.  Since line $j$ has
-canalyzed output $b_j$, line $k$ is the only line in its layer.
+the canalyzed value is $\overline{b_j}$.  Since line $j$ has
+canalyzed value $b_j$, line $k$ is the only line in its layer.
 Since $f$ is normalized, the last layer of $f$ contains at least
 two lines.  Thus, there is a last layer following the layer containing
-line $k$.  The canalyzed output of all the lines in this last layer
+line $k$.  The canalyzed value of all the lines in this last layer
 is $b_j$.  Thus, for every line $i$, $1 \leq i \leq n$, the value
@@
 canalyzing value of line $i$.  Consequently, $f(\alpha') =
-\overline{b_j}$, the complement of the canalyzed output in the last
+\overline{b_j}$, the complement of the canalyzed value in the last
 layer of $f$.  \QED
@@
 $a_{k'}$ is either the same or the complement of $a_k$, and canalyzed
-output $b_{k'}$ is either the same or the complement of $b_k$, Thus,
+value $b_{k'}$ is either the same or the complement of $b_k$, Thus,
 there are four possible cases for the form of line $k'$ of $f$.  We
@@
 Since line $k'$
-has the same canalyzing value and canalyzed output as line $k$, and
+has the same canalyzing value and canalyzed value as line $k$, and
 $f$ is properly $n$-symmetric, line $k'$ must occur in a lower layer
 than that of line $k$.  Thus, there is at least one line between
-line $k$ and line $k'$ for which the canalyzed output is $\overline{b_k}$.
+line $k$ and line $k'$ for which the canalyzed value is $\overline{b_k}$.
 Let line $q$ be the first such line.  Note that for all $i$ such
@@
 value of line $i$, but $c_{\pi(q)}$ does match the canalyzing value
-of line $q$.  Since canalyzed output $b_q$ equals $\overline{b_k}$,
+of line $q$.  Since canalyzed value $b_q$ equals $\overline{b_k}$,
 we have that $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)}) =
@@
 Let line $q$ be the first line such that $k < q \leq k'$ and
-the canalyzed output of line $q$ is $\overline{b_k}$.
+the canalyzed value of line $q$ is $\overline{b_k}$.
 Note that for all $i$ such that $1 \leq i < q$, 
@@
 Suppose there is a line below line $k$ for which the canalyzed
-output is $\overline{b_k}$.  Let $q$ be the first such line.  Note
+value is $\overline{b_k}$.  Let $q$ be the first such line.  Note
 that for all $i$ such that $1 \leq i < q$, $c_{\pi(i)}$ does not
@@
 Now suppose there is no line below line $k$ for which the canalyzed
-output is $\overline{b_k}$.  Then line $k$ occurs in the last layer
+value is $\overline{b_k}$.  Then line $k$ occurs in the last layer
 of $f$.  Since $f$ is both normalized and properly $n$-symmetric,
@@
 Suppose there is a line, other than line $k'$, below line $k$ for
-which the canalyzed output is $\overline{b_k}$.  Let $q$ be the
+which the canalyzed value is $\overline{b_k}$.  Let $q$ be the
 first such line.  Note that for all $i$ such that $1 \leq i < q$,
@@
 Now suppose that line $k'$ is the only line below line $k$ for which
-the canalyzed output is $\overline{b_k}$.  Since line $k$ has
-canalyzed output $b_k$, line $k'$ is the only line in its layer.
+the canalyzed value is $\overline{b_k}$.  Since line $k$ has
+canalyzed value $b_k$, line $k'$ is the only line in its layer.
 Since $f$ is normalized, there is a last layer following the layer
@@
 representation, the last line would be changed to have canalyzing
-value 0, and the same canalyzed output as the preceding line.
+value 0, and the same canalyzed value as the preceding line.
 

### 2019-02-06 intro.tex
ssr

@@
 $f(1, x_2, x_3) ~=~ 0$, for all combinations of values for 
-$x_2$ and $x_3$. \qed
+$x_2$ and $x_3$. %\qed
 
@@
 \noindent
-\qed
+%\qed
 
+\iffalse
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \noindent
@@
 \noindent
-\qed
+%\qed
 
@@
 %% such that the value of the function is 1.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\fi
 
@@
 it is $r$-symmetric, but not $(r-1)$-symmetric.
-For example, the function $f_3(x_1, x_2, x_3) = (x_1 \wedge x_2) \vee\, \overline{x_3}$
+For example, the function $f_3(x_1, x_2, x_3) = (x_1 \wedge x_2) \vee\, \overline{x_3}$~
 is not 1-symmetric since $f_3(1, 0, 1) \neq f_3(1, 1, 0)$; however, 
@@
 
-In the literature (see, e.g., \cite{Crama-Hammer-2011,HT-2016,Toth-etal-1977}),
+In the literature (see e.g., \cite{Crama-Hammer-2011,HT-2016,Toth-etal-1977}),
 a 1-symmetric function $f$ is referred to simply
@@
 The class of $r$-symmetric functions have also been
-studied in the literature on discrete dynamical systems (see, e.g., 
+studied in the literature on discrete dynamical systems (see e.g., 
 \cite{Barrett-etal-2007,Rosenkrantz-etal-2015,MR-2007}).
@@
 the property that the symmetry level of $f$ is $n$.
-Finally, we identify all the Boolean functions that NCFs 
-as well as symmetric.
+Finally, we identify all the Boolean functions that are NCFs 
+and symmetric.
 

### 2019-02-06 preliminaries.tex
ssr

@@
 where $\pi$ is a permutation of $\{1, 2, \ldots, n\}$.
-We will refer to this as the \textbf{lexicographic} ordering
-of the variables in the specification.
-
-\medskip
-
-Throughout this paper, we assume that an NCF $f$ with $n$
+%% We will refer to this as the \textbf{lexicographic} ordering
+%% of the variables in the specification.
+Throughout this paper, we will assume that the variable
+ordering uses the identity permutation so that $x_i$ is the 
+variable tested in line $i$, $1 \leq i \leq n$. 
+We assume that an NCF $f$ with $n$
 variables is specified using the simplified representation with $n$ lines. 
@@
 in the results stated in this section.
-
-The following observation is from \cite{Stearns-etal-2018}.
+We will use the following result from \cite{Stearns-etal-2018}.
 
@@
 \medskip
-We can now define the notion of a \textbf{normalized} 
+We can now define the notion of a \textbf{default-normalized} 
 representation for NCFs.
@@
 the simplified representation with 
-lexicographic ordering given by  
-$\langle x_{\pi(1)}, x_{\pi(2)}, \ldots, x_{\pi(n)}\rangle$.
-We say that the representation is {\bf normalized} if
-both of the following conditions hold.
-\begin{enumerate}
-  \item Lines $n-1$ and $n$ have the same canalyzed value ~and
-  \item each pair of consecutive lines
-        with the same canalyzed value occur in the lexicographic order 
-        of their canalyzing variables; that is, if two consecutive lines are \\
-%\medskip
-\noindent
-\hspace*{1.1in} $x_i$:~ $a_i \longrightarrow b$ \\
-\hspace*{1.1in} $x_j$:~ $a_j \longrightarrow b$
-
-%\medskip
-\noindent
-then $x_i$ precedes $x_j$ in the lexicographic ordering.
-\end{enumerate}
+the variable ordering $\langle x_1, x_2, \ldots, x_n\rangle$.
+We say that the representation is {\bf default-normalized} if
+lines $n-1$ and $n$ have the same canalyzed value.
 \end{definition}
@@
 \textbf{Example 5:}~ 
-
-\medskip
-\noindent
-\textbf{Part 1:}~
 Consider the NCF $f$ shown in Example~4 with
@@
 value on the ``Default" line to 0, then we obtain 
-a normalized representation.
-
-\medskip
-\noindent
-\textbf{Part 2:}~ Consider the normalized version of Example~4 
-produced in Part~1 above. 
-In that representation, if we exchange Lines 1 and 2, 
-by Part~(a) of Observation~\ref{obs:ncf_transformations},
-the function does not change. 
-However, the resulting representation is \emph{not} normalized
-since the line with $x_2$ precedes the one 
-with $x_1$. %\qed
+a default-normalized representation.
 
@@
 so that it has the same canalyzed value as the preceding line,
-and then we can lexicographically sort
-each maximal sequence of consecutive lines with the same canalyzed value,
-thereby obtaining an equivalent normalized NCF representation.
+thereby obtaining an equivalent default-normalized NCF representation.
 We state this formally below.
@@
 \begin{observation}\label{obs:normalization_poly}
-Given an NCF representation for a function $f$, a normalized
+Given an NCF representation for a function $f$, a default-normalized
 representation for $f$ can be obtained in polynomial time. \QED
@@
 that approximates the symmetry level of a Boolean function
-specified as Boolean expression to within the factor $\rho$.
+specified as a Boolean expression to within the factor $\rho$.
 Without loss of generality, we can assume that $\rho$ is an integer.

### 2019-02-08 symmetry_properties.tex
ssr

@@
 
+\subsection{Number of Strongly Asymmetric NCFs}
+\label{sse:number_strongly_asymmetric}
+
+\begin{theorem}\label{thm:count_strongly_asymmetric}
+For any $n \geq 2$, the number of Boolean functions with $n$ variables 
+that are both strongly asymmetric and NCFs is $n! \, 2^n$.
+\end{theorem}
+
+\noindent
+\textbf{Proof:}~
+Consider a default normalized representation of a strongly asymmetric NCF $f$.
+The representation contains $n-1$ layers.
+The first $n-2$ layers each contain one line, and the last layer contains two lines.
+
+Consider the canalyzed values in $f$. The canalyzed value in the
+first layer can be either 0 or 1.  The canalyzed value in every
+other layer is the complement of the canalyzed value in the preceding
+layer.  Thus, there are 2 possibilities for the sequence of canalyzed
+values in $f$.
+
+The canalyzing variable in each of the first $n-2$ lines (one per
+layer) can be any variable that has not yet occurred in a preceding
+line.  The last layer contains the remaining two variables.  Thus,
+there are $n! / 2$ possibilities for the pattern of variable
+occurrences in $f$.
+
+For each of the $n$ variables, the canalyzing value for the line
+containing that variable can be either 0 or 1.  Thus, there are
+$2^n$ possibilities for canalyzing values.  \QED
+
 \subsection{Symmetric Canalyzing and Nested Canalyzing Functions}

### 2019-02-08 abstract.tex
ssr

@@
 Boolean functions. 
+Using a standard representation for NCFs, we develop a 
+characterization of when two variables of an NCF are symmetric.
 A general form of symmetric Boolean functions,
@@
 a Boolean function to within any factor, unless \textbf{P} = \cnp. 
-In contrast, we show that the symmetry level of an NCF $f$
+In contrast, using the characterization mentioned aboe, we show 
+that the symmetry level of an NCF $f$
 can be easily computed given a standard representation of $f$.
@@
 the property that the symmetry level of $f$ is $n$.
+We use this result to derive a closed form expression for the
+number of $n$-variable Boolean functions that are NCFs and strongly
+asymmetric.
 In addition, we identify all the Boolean functions that are NCFs 

### 2019-02-08 preliminaries.tex
ssr

@@
 \begin{observation}\label{obs:normalization_poly}
-Given an NCF representation for a function $f$, a default-normalized
+Given an NCF representation for a Boolean function $f$, a default-normalized
 representation for $f$ can be obtained in polynomial time. \QED

### 2019-02-08 abstract.tex
ssr

@@
 such as sensitivity and stability.
-In this note, we examine the relationships between NCFs and symmetric 
-Boolean functions. 
+In this note, we examine the relationships between NCFs, symmetric 
+Boolean functions and a more generl form of symmetric Boolean functions,
+called $r$-symmetric functions (where $r$ is the symmetry level).
 Using a standard representation for NCFs, we develop a 
 characterization of when two variables of an NCF are symmetric.
-A general form of symmetric Boolean functions,
-called $r$-symmetric functions (where $r$ is the symmetry level),
-has been studied in the literature.
-We observe that, in general,
-one cannot efficiently approximate the symmetry level of
-a Boolean function to within any factor, unless \textbf{P} = \cnp. 
-In contrast, using the characterization mentioned aboe, we show 
+Using this characterization, we show 
 that the symmetry level of an NCF $f$
 can be easily computed given a standard representation of $f$.
+(In general, we show that the symmetry level of
+a Boolean function cannot be efficiently approximated
+to within any factor, unless \textbf{P} = \cnp.)
 We also show that for any NCF $f$ with $n$ variables, the notion of
 strong asymmetry considered in the literature is equivalent to
-the property that the symmetry level of $f$ is $n$.
+the property that $f$ is $n$-symmetric. 
 We use this result to derive a closed form expression for the
-number of $n$-variable Boolean functions that are NCFs and strongly
-asymmetric.
+number of $n$-variable Boolean functions 
+that are NCFs and strongly asymmetric.
 In addition, we identify all the Boolean functions that are NCFs 

### 2019-02-08 abstract.tex
ssr

@@
 In this note, we examine the relationships between NCFs, symmetric 
-Boolean functions and a more generl form of symmetric Boolean functions,
+Boolean functions and a more general form of symmetric Boolean functions,
 called $r$-symmetric functions (where $r$ is the symmetry level).
@@
 can be easily computed given a standard representation of $f$.
-(In general, we show that the symmetry level of
+(In general, we prove that the symmetry level of
 a Boolean function cannot be efficiently approximated

### 2019-02-11 abstract.tex
ssr

@@
 can be easily computed given a standard representation of $f$.
-(In general, we prove that the symmetry level of
-a Boolean function cannot be efficiently approximated
-to within any factor, unless \textbf{P} = \cnp.)
+(In contrast, we prove that even approximating the symmetry level of
+a general Boolean function to within any factor $\geq 1$ is \cnp-hard.) 
 We also show that for any NCF $f$ with $n$ variables, the notion of

### 2019-02-11 intro.tex
ssr

@@
 
-In this note, our focus is on the relationship between 
+Our focus is on the relationships between 
 NCFs and symmetric Boolean functions.
-We observe that, in general,
-one cannot efficiently approximate the symmetry level of
-a Boolean function to within any factor, unless \textbf{P} = \cnp.
-In contrast, we show that the symmetry level of an NCF $f$
+Using a standard representation for NCFs, we develop a
+characterization of when two variables of an NCF are symmetric.
+Using this characterization, we show
+that the symmetry level of an NCF $f$
 can be easily computed given a standard representation of $f$.
+(In contrast, we show that 
+one cannot even efficiently approximate the symmetry level of
+a general Boolean function to within any factor $\geq 1$, 
+unless \textbf{P} = \cnp.)
 We also show that for any NCF $f$ with $n$ variables, the notion of
 strong asymmetry considered in the literature is equivalent to
-the property that the symmetry level of $f$ is $n$.
-Finally, we identify all the Boolean functions that are NCFs 
-and symmetric.
+the property that $f$ is $n$-symmetric.
+We use this result to derive a closed form expression for the
+number of $n$-variable Boolean functions
+that are NCFs and strongly asymmetric.
+In addition, we identify all the Boolean functions that are 
+canalyzing and symmetric as well as those that 
+are NCFs and symmetric.
 
@@
 In the literature on computational learning theory,
-generalized NCFs are referred to as $1$-\textbf{decision lists} \cite{KV-1994}.
+NCFs are referred to as $1$-\textbf{decision lists} \cite{KV-1994}.
 Many researchers have pointed out the usefulness of NCFs 
@@
 Li-etal-2011,Li-etal-2013,Klotz-etal-2013, Stearns-etal-2018}. 
-To our knowledge, the relationship between NCFs and symmetric
-Boolean functions has not been addressed in the literature.
+To our knowledge, relationships between NCFs and symmetric
+Boolean functions have not been addressed in the literature.

### 2019-02-11 preliminaries.tex
ssr

@@
 We also show that, in general, the symmetry level
-of a Boolean function cannot be approximated to within any
+of a Boolean function cannot be efficiently approximated to within any
 factor $\rho \geq 1$, unless \textbf{P} = \cnp.
@@
 %% of the variables in the specification.
-Throughout this paper, we will assume that the variable
-ordering uses the identity permutation so that $x_i$ is the 
-variable tested in line $i$, $1 \leq i \leq n$. 
+Throughout this paper, we will assume without loss of generality
+that the variable ordering uses the identity permutation 
+so that $x_i$ is the variable tested in line $i$, $1 \leq i \leq n$. 
 We assume that an NCF $f$ with $n$
@@
 
-\noindent
+%%\noindent
 Li et al. \cite{Li-etal-2013} have defined the concept of a 
@@
 \textbf{Example 5:}~ 
-Consider the NCF $f$ shown in Example~4 with
-the variable ordering $x_1$, $x_2$, $x_3$, $x_4$, $x_5$, $x_6$.
-This representation is \emph{not} normalized since lines 5 and 6
+Consider the NCF $f$ shown in Example~4. 
+This representation is \emph{not} default-normalized since lines 5 and 6
 have different canalyzed values.
@@
 
-\subsection{Complexity of Estimating the Symmetry Level of a 
-Boolean Function}
+In view of Observation~\ref{obs:normalization_poly},
+we will assume that any NCF is specified in default-normalized form.
+
+\subsection{Complexity of Approximating the Symmetry Level of a 
+General Boolean Function}
 \label{sse:symmetry_level_hardness}
@@
 
-Let $f$ be a CNF formula representing an instance of SAT.
+Let $g$ be a CNF formula representing an instance of SAT.
 Let $X$ = $\{x_1, x_2, \ldots, x_n\}$ denote the set of
-Boolean variables used in $f$.
-We create another CNF formula $f'$ as follows.
+Boolean variables used in $g$.
+We create another CNF formula $f$ as follows.
 Let $Y = \{y_1, y_2, \ldots, y_{\rho+1}\}$ and
@@
 new sets of variables, with each set containing $\rho+1$ variables.
-The expression for $f'$ is given by
+The expression for $f$ is given by
 \[
-f'(x_1, \ldots, x_n, y_1, \ldots, y_{\rho+1},
+f(x_1, \ldots, x_n, y_1, \ldots, y_{\rho+1},
 z_1, \ldots, z_{\rho+1}) ~=~ 
-    f(x_1, \ldots x_n) \wedge (y_1 \vee \overline{z_1}) \wedge \ldots \wedge
+    g(x_1, \ldots x_n) \wedge (y_1 \vee \overline{z_1}) \wedge \ldots \wedge
                               (y_{\rho+1} \vee \overline{z_{\rho+1}}).
 \]
-We note that $f'$ is also a CNF formula.
+We note that $f$ is also a CNF formula.
 We have the following claims.
@@
 \noindent
-\textbf{Claim 1:}~ If $f$ is \emph{not} satisfiable, 
-the symmetry level of $f'$ is 1.
+\textbf{Claim 1:}~ If $g$ is \emph{not} satisfiable, 
+the symmetry level of $f$ is 1.
 
@@
 \noindent
-\textbf{Proof of Claim 1:}~ If $f$ is not satisfiable, $f'$ is also
-not satisfiable; that is, for all inputs, the value of $f'$ is 0.
+\textbf{Proof of Claim 1:}~ If $g$ is not satisfiable, $f$ is also
+not satisfiable; that is, for all inputs, the value of $f$ is 0.
 Thus, all the variables in $X \cup Y \cup Z$ can be included in one
 symmetry group.
-The value of $f'$ is 0 regardless of how many variables in the
+The value of $f$ is 0 regardless of how many variables in the
 group have the value 1.  \hfill$\Box$
@@
 \noindent
-\textbf{Claim 2:}~ If $f$ \emph{is} satisfiable, 
-the symmetry level of $f'$ is at least $\rho+1$.
+\textbf{Claim 2:}~ If $g$ \emph{is} satisfiable, 
+the symmetry level of $f$ is at least $\rho+1$.
 
@@
 \noindent
-\textbf{Proof of Claim 2:}~ Suppose $f$ is satisfiable.
+\textbf{Proof of Claim 2:}~ Suppose $g$ is satisfiable.
 We argue that if $i \neq j$, variables $y_i$ and $y_j$
-cannot be in the same symmetry group for the function $f'$.
+cannot be in the same symmetry group for the function $f$.
 To see this, consider any two variables $y_i$ and $y_j$ with $i \neq j$,
 and construct an assignment $\alpha$ to
-the variables of $f'$ in the following manner.
+the variables of $f$ in the following manner.
 \begin{description}
 \item{(i)} For the variables in $X$, choose an assignment that
-satisfies $f$.
+satisfies $g$.
 \item{(ii)} Let $y_i = 0$, $z_i = 1$, $y_j = 1$ and $z_j = 0$.
@@
 Since this assignment $\alpha$ sets the clause $(y_i \vee \overline{z_i})$ to 0,
-the value of $f'$ under the assignment $\alpha$ is 0.
+the value of $f$ under the assignment $\alpha$ is 0.
 Now, consider the assignment $\alpha'$ that is obtained by interchanging
@@
 values assigned to the other variables. 
-It can be seen that under assignment $\alpha'$, the value of $f'$ is 1.
+It can be seen that under assignment $\alpha'$, the value of $f$ is 1.
 In both $\alpha$ and $\alpha'$, the number of 1-valued variables in 
 the set $\{y_i, y_j\}$ is 1; however, the two assignments lead to 
-different values for $f'$.
-Therefore, if $f$ is satisfiable, for $i \neq j$, $y_i$ and $y_j$ cannot
-be in the same symmetry group.
+different values for $f$.
+Therefore, if $g$ is satisfiable, for $i \neq j$, $y_i$ and $y_j$ cannot
+be in the same symmetry group for $f$.
 Since there are $\rho+1$ variables in $Y$, the number of symmetry groups 
-for $f'$ must be at least $\rho+1$. \hfill$\Box$
+for $f$ must be at least $\rho+1$. \hfill$\Box$
 
@@
 
-Now, suppose we execute Algorithm \cala{} on the function $f'$.
-If $f$ is \emph{not} satisfiable, then from Claim 1, 
-the symmetry level of $f'$ is 1. 
+Now, suppose we execute Algorithm \cala{} on the function $f$.
+If $g$ is \emph{not} satisfiable, then from Claim~1, 
+the symmetry level of $f$ is 1. 
 Since \cala{} is a $\rho$-approximation algorithm,
 it should produce at most $\rho$ symmetry groups.
-On the other hand, if $f$ is satisfiable, by Claim 2, the symmetry
-level of $f'$ is at least $\rho+1$; so, \cala{} will produce at least
+On the other hand, if $g$ is satisfiable, by Claim~2, the symmetry
+level of $f$ is at least $\rho+1$; so, \cala{} will produce at least
 $\rho+1$ groups.
-Thus, by examining the output of \cala, we can determine whether or not $f$
-is satisfiable.
+In other words, $g$ is not satisfiable iff the number
+of symmetry groups produced by \cala{} is at most $\rho$.
 Since \cala{} runs in polynomial time, we have an efficient 
-algorithm for SAT, contradicting the assumption that \textbf{P} $\neq$ \cnp.
-\QED
+algorithm for SAT, contradicting the assumption that 
+\textbf{P} $\neq$ \cnp.  \QED
 
@@
 we will show in the next section that the symmetry level 
-of an NCF (or that of a
-generalized NCF) can be computed efficiently.
+of an NCF can be computed efficiently.

### 2019-02-11 symmetry_properties.tex
ssr

@@
 for the symmetry level of an NCF specified using
-its normalized representation.
-Our next result (Theorem~\ref{thm:ncf_strong_asymmetry})
-shows that for any NCF $f$ with $n$ variables,
-the two statements ``$f$ is strongly asymmetric"
-``the symmetry level of $f$ is $n$" are equivalent. 
+its default-normalized representation.
+%% We also list some corollaries that follow from the
+%% characterization.
+Next, we show that for any NCF $f$ with $n$ variables,
+the two statements ``$f$ is strongly asymmetric" ~and~
+``the symmetry level of $f$ is $n$"~ are equivalent 
+(Theorem~\ref{thm:ncf_strong_asymmetry}).
+We use this result to derive a closed form expression 
+for the number of $n$-variable NCFs that are 
+also strongly asymmetric (Theorem~\ref{thm:count_strongly_asymmetric}).
 Our final result (Proposition~\ref{pro:ncf_symmetric})
@@
 Two variables of an NCF $f$ are symmetric iff
-in the normalized representation of $f$,
+in the default-normalized representation of $f$,
 they occur in the same layer and have the same canalyzing value.
@@
 Suppose that two variables of an NCF $f$ occur in the same layer of a
-normalized representation, and the lines containing these variables
+default-normalized representation, and the lines containing these variables
 have the same canalyzing value.  Then in any input assignment, the
@@
 
-For the converse, consider the normalized NCF representation of $f$. 
+For the converse, consider the default-normalized NCF representation of $f$. 
 For purposes of notational simplicity, we assume without loss of generality that the
@@
 canalyzed value $b_j$, line $k$ is the only line in its layer.
-Since $f$ is normalized, the last layer of $f$ contains at least
+Since $f$ is default-normalized, the last layer of $f$ contains at least
 two lines.  Thus, there is a last layer following the layer containing
@@
 \begin{theorem}\label{thm:ncf_r_symmetric}
-Suppose the normalized representation of a given NCF contains 
+Suppose the default-normalized representation of a given NCF contains 
 $r_1$ layers with only one distinct canalyzing value,
@@
 \begin{corollary}\label{cor:ncf_r_symmetric_layers}
-(i) An NCF with a normalized nested canalyzing representation consisting
+(i) An NCF with a default-normalized nested canalyzing representation consisting
 of $q$ layers is $2q$-symmetric, and is not $(q-1)$-symmetric.
-(ii) An $r$-symmetric NCF has a normalized nested canalyzing representation
+(ii) An $r$-symmetric NCF has a default-normalized nested canalyzing representation
 with at most $r-1$ layers, and has an unnormalized nested canalyzing
@@
 For the converse, consider an $n$ variable NCF that is properly $n$-symmetric.
-Let $f$ be the normalized NCF representation for the function.  For
+Let $f$ be the default-normalized NCF representation for the function.  For
 purposes of notational simplicity, without loss of generality,
@@
 value is $\overline{b_k}$.  Then line $k$ occurs in the last layer
-of $f$.  Since $f$ is both normalized and properly $n$-symmetric,
+of $f$.  Since $f$ is both default-normalized and properly $n$-symmetric,
 this last layer contains exactly two lines.  Thus $k = n-1$ and $k'
@@
 canalyzed value $b_k$, line $k'$ is the only line in its layer.
-Since $f$ is normalized, there is a last layer following the layer
+Since $f$ is default-normalized, there is a last layer following the layer
 containing line $k'$.  Thus, for all $i$, $c_{\pi(i)}$ does not
@@
 \textbf{Proof:}~
-Consider a default normalized representation of a strongly asymmetric NCF $f$.
+Consider a default-normalized representation of a strongly asymmetric NCF $f$.
 The representation contains $n-1$ layers.

### 2019-02-28 intro.tex
ssr: added refs and edited related work

@@
 Li-etal-2011,Li-etal-2013,Klotz-etal-2013, Stearns-etal-2018}.
+
+Over the years, a considerable amount of research on detecting symmetry properties 
+of Boolean functions has been reported in the literature
+(e.g., \cite{BS-1968,Biswas-1970,TM-1996,KS-2000,Darga-etal-2008,Maurer-2015}).
+Several references have observed the usefulness of detecting symmetries
+in automated logic synthesis (e.g., \cite{AP-2008,Hu-etal-2008,Darga-etal-2008}).
+Reference \cite{Maurer-2015} provides a thorough discussion of known
+symmetry detection algorithms and presents a new universal algorithm 
+for detecting any form of permutation-based symmetry in Boolean functions. 
 To our knowledge, relationships between NCFs and symmetric

### 2019-03-01 IEEE_TC_submission.d/intro.tex
ssr

--- 2019-02-28 intro.tex
@@
 
-The class of \textbf{canalyzing} Boolean functions, introduced 
+\IEEEPARstart{T}{he} class of \textbf{canalyzing} Boolean functions, introduced 
 in \cite{Kauffman-1969}, is defined as follows.
@@
 \]
-regardless of the values assigned to the variables in $X - \{x_i\}$.
+for all combinations of values assigned to the variables in $X - \{x_i\}$.
 \end{definition}
@@
 \textbf{Example 1:}~ Consider the function 
-$f(x_1, x_2, x_3) ~=~ \overline{x_1} \wedge (x_2 \vee \overline{x_3})$.
+$f(x_1, x_2, x_3)$ $~=~$ $\overline{x_1} \wedge (x_2 \vee \overline{x_3})$.
 This function is canalyzing since if we set to $x_1 = 1$,
@@
        \:b_1 & \mathrm{if~~} x_{\pi(1)} ~=~ a_1 \\
-       \:b_2 & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~}
-            x_{\pi(2)} ~=~ a_2 \\
-       \:\vdots & \vdots \\
-       \:b_n & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots~
-             x_{\pi(n-1)} ~\neq~ a_{n-1} \mathrm{~~and~~} x_{\pi(n)} ~=~ a_n \\
-       \:\overline{b_n} & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots~
-            x_{\pi(n)} ~\neq~ a_n \\
+       \:b_2 & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~}\\
+             & x_{\pi(2)} ~=~ a_2 \\
+       \:\vdots & \vdots \\ %\smallskip
+       \:b_n & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots \\
+             & x_{\pi(n-1)} ~\neq~ a_{n-1} \mathrm{~~and~~}\\
+             & x_{\pi(n)} ~=~ a_n \\ %\smallskip
+       \:\overline{b_n} & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots\\
+                        & x_{\pi(n)} ~\neq~ a_n \\
    \end{cases}
@@
 \noindent
-\hspace*{1.1in} $x_{\pi(i)}:~ a_i ~~\longrightarrow~~ b_i$
+\hspace*{0.5in} $x_{\pi(i)}:~ a_i ~~\longrightarrow~~ b_i$
 
@@
 \noindent
-\hspace*{1.1in} Default:~ $\overline{b_n}$
+\hspace*{0.5in} Default:~ $\overline{b_n}$
 
@@
        \:1 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~}
-            x_{2} ~\neq~ 1 \mathrm{~~and~~} x_{3} = 0 \\
+            x_{2} ~\neq~ 1 \mathrm{~~and~~}\\ 
+           &x_{3} = 0 \\
        \:0 & \mathrm{if~~} x_{1} ~\neq~ 1 \mathrm{~~and~~} 
-            x_{2} ~\neq~ 1 \mathrm{~~and~~} x_{3} ~\neq~ 0 \\
+             x_{2} ~\neq~ 1 \mathrm{~~and~~}\\ 
+           & x_{3} ~\neq~ 0 \\ \\ 
    \end{cases}
@@
 \begin{tabular}{ll}
-\hspace*{1.1in} & $x_1:~$  $1 ~\longrightarrow~ 0$ \\ [1ex]
-\hspace*{1.1in} & $x_2:~$  $1 ~\longrightarrow~ 1$ \\ [1ex]
-\hspace*{1.1in} & $x_3:~$  $0 ~\longrightarrow~ 1$ \\ [1ex]
-\hspace*{1.1in} & Default:~ $0$ \\
+\hspace*{0.5in} & $x_1:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
+\hspace*{0.5in} & $x_2:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
+\hspace*{0.5in} & $x_3:~$  $0 ~\longrightarrow~ 1$ \\ [0.5ex]
+\hspace*{0.5in} & Default:~ $0$ \\
 \end{tabular}

### 2019-03-01 IEEE_TC_submission.d/preliminaries.tex
ssr

--- 2019-02-11 preliminaries.tex
@@
-\section{Other Definitions and Preliminary Results}
+\section{Other Definitions and Preliminary\newline Results}
 \label{sec:prelim}
@@
 in the simplified representation. 
-\begin{description}
-\item{(a)} 
+%\begin{description}
+\begin{enumerate}
+\item 
 For any $q \geq 2$ and for any $i$, $1 \leq i \leq n-q+1$,
@@
 
-\item{(b)} 
+\item 
 Suppose lines $n-1$ and $n$ in the specification of $f$ 
@@
 \QED
-\end{description}
+\end{enumerate}
+%\end{description}
 \end{observation}
@@
 \medskip
-From Part~(a) of Observation~\ref{obs:ncf_transformations},
+From Part~1 of Observation~\ref{obs:ncf_transformations},
 it follows that the lines within the same layer of an NCF
 representation can be permuted without changing the function.
-Part~(b) of Observation~\ref{obs:ncf_transformations} points
+Part~2 of Observation~\ref{obs:ncf_transformations} points
 out that for any NCF $f$ with $n \geq 2$ variables, there is a simplified 
@@
 new sets of variables, with each set containing $\rho+1$ variables.
-The expression for $f$ is given by
+The expression for $f(x_1, \ldots, x_n, y_1, \ldots, y_{\rho+1})$
+is 
 \[
-f(x_1, \ldots, x_n, y_1, \ldots, y_{\rho+1},
-z_1, \ldots, z_{\rho+1}) ~=~ 
     g(x_1, \ldots x_n) \wedge (y_1 \vee \overline{z_1}) \wedge \ldots \wedge
@@
 the variables of $f$ in the following manner.
-\begin{description}
-\item{(i)} For the variables in $X$, choose an assignment that
+\begin{enumerate}
+\item For the variables in $X$, choose an assignment that
 satisfies $g$.
-\item{(ii)} Let $y_i = 0$, $z_i = 1$, $y_j = 1$ and $z_j = 0$.
-\item{(iii)} For $1 \leq p \leq \rho+1$, $p \neq i$ and $p \neq j$,
+\item Let $y_i = 0$, $z_i = 1$, $y_j = 1$ and $z_j = 0$.
+\item For $1 \leq p \leq \rho+1$, $p \neq i$ and $p \neq j$,
 let $y_p = 1$ and $z_p$ = 0.
-\end{description}
+\end{enumerate}
 Since this assignment $\alpha$ sets the clause $(y_i \vee \overline{z_i})$ to 0,

### 2019-03-01 IEEE_TC_submission.d/symmetry_properties.tex
ssr

--- 2019-02-11 symmetry_properties.tex
@@
-\section{Symmetry of Nested Canalyzing Functions}
+\section{Symmetry of Nested Canalyzing\newline Functions}
 \label{sec:ncf_and_symmetry}
@@
 to the variables of $f$.
-\begin{description}
-\item{(i)} For $1 \leq i < j$, set $c_i = \overline{a_i}$. 
-\item{(ii)} For $i = j$, set $c_j =a_j$. 
-\item{(iii)} For $i = k$, set $c_k =\overline{a_j}$. 
-\item{(iv)} For $j < i < k$, and for  $k < i \leq n$, 
+\begin{enumerate}
+\item For $1 \leq i < j$, set $c_i = \overline{a_i}$. 
+\item For $i = j$, set $c_j =a_j$. 
+\item For $i = k$, set $c_k =\overline{a_j}$. 
+\item For $j < i < k$, and for  $k < i \leq n$, 
 if $b_i = b_j$, then set $c_i =  \overline{a_i}$,
 else set $c_i =  a_i$.
-\end{description}
+\end{enumerate}
 Note that $f(\alpha) = b_j$, since the variables in all lines above
@@
 Assignment $(c_1, c_2, \ldots,  c_n)$ is constructed as follows.
-\begin{description}
-\item{(i)} For $1 \leq i < k$, set $c_i = \overline{a_i}$.  
-\item{(ii)} For $i = k$, set $c_k =a_k$. 
-\item{(iii)} For $i > k$, let $i' = \pi^{-1}(i)$, so that $\pi(i')
+\begin{enumerate}
+\item For $1 \leq i < k$, set $c_i = \overline{a_i}$.  
+\item For $i = k$, set $c_k =a_k$. 
+\item For $i > k$, let $i' = \pi^{-1}(i)$, so that $\pi(i')
 = i$.  Thus, after the permutation of values, variable $x_{i'}$
@@
 \overline{a_{i'}}$, else set $c_i =  a_{i'}$.
-\end{description}
+\end{enumerate}
 Since $c_k$ matches the canalyzing value of line $k$ of $f$, and
@@
 Now suppose there is no line below line $k$ for which the canalyzed
-value is $\overline{b_k}$.  Then line $k$ occurs in the last layer
-of $f$.  Since $f$ is both default-normalized and properly $n$-symmetric,
+value is $\overline{b_k}$.  
+Then line $k$ occurs in the last layer of $f$.  
+Since $f$ is both default-normalized and properly $n$-symmetric,
 this last layer contains exactly two lines.  Thus $k = n-1$ and $k'
@@
 
+\noindent
 Examples of properly $n$-symmetric $n$-variable NCFs satisfying the
@@
 
-\hspace*{1in} $x_1 \vee (\overline{x}_2 \wedge ( x_3 \vee (\overline{x}_4 \wedge (\cdots ))$
+\hspace*{0.5in} $x_1 \vee (\overline{x}_2 \wedge ( x_3 \vee (\overline{x}_4 \wedge (\cdots ))$
 
@@
 For instance,
-\[
-f_6 = x_1 \vee (\overline{x}_2 \wedge (x_3 \vee (\overline{x}_4 \wedge (x_5 \vee \overline{x}_6))
-~~~~\mathrm{and}~~~~
-f_7 = x_1 \vee (\overline{x}_2 \wedge (x_3 \vee (\overline{x}_4 \wedge (x_5 
+\begin{align*}
+f_6 &~=~  x_1 \vee (\overline{x}_2 \wedge (x_3 \vee (\overline{x}_4 \wedge 
+      (x_5 \vee \overline{x}_6)) \\
+f_7 &~=~  x_1 \vee (\overline{x}_2 \wedge (x_3 \vee (\overline{x}_4 \wedge (x_5 
           \vee (\overline{x}_6 \wedge  x_7)).
-\]
+\end{align*}
 Function $f_n(x_1, x_2, \ldots, x_n)$ is the NCF corresponding to

### 2019-03-04 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@
 
-\hspace*{1in}
+\hspace*{0.25in}
 $x_i : a_i ~\longrightarrow~ b_i$.
@@
 
-\hspace*{1in} $x_i : a_i ~\longrightarrow~ b_i$.
+\hspace*{0.25in} $x_i : a_i ~\longrightarrow~ b_i$.
 
@@
 \noindent
-Examples of properly $n$-symmetric $n$-variable NCFs satisfying the
-conditions of Theorem \ref{thm:ncf_r_symmetric} are the following.
-For each $n > 1$, let $f_n$ be the function defined by the following
-formula:
+We now provide
+examples of properly $n$-symmetric $n$-variable NCFs satisfying the
+conditions of Theorem \ref{thm:ncf_r_symmetric}.
+For each $n > 1$, let $f_n$ be the function of $n$ variables, namely
+$x_1, \ldots, x_n$, defined by the following formula:
 
@@
 
-\smallskip
+\medskip
 
 \noindent
-\hspace*{0.5in}
+\hspace*{0.25in}
 $x_1:~ 1 ~\longrightarrow~ 1$ \\
-\hspace*{0.5in}
+\hspace*{0.25in}
 $x_2:~ 1 ~\longrightarrow~ 0$ \\
-\hspace*{0.5in}
+\hspace*{0.25in}
 $x_3:~ 1 ~\longrightarrow~ 1$ \\
@@
 
+\smallskip
+
 \noindent
@@
 
+\smallskip
+
 \noindent
-\hspace*{0.5in}
+\hspace*{0.25in}
 $x_n:~ 1 ~\longrightarrow~ 1$
+
+\smallskip
 
@@
 
+\smallskip
+
 \noindent
-\hspace*{0.5in}
+\hspace*{0.25in}
 $x_n:~ 1 ~\longrightarrow~ 0$.

### 2019-03-05 IEEE_TC_submission.d/abstract.tex
ssr

--- 2019-02-11 abstract.tex
@@
 \begin{center}
-\textbf{Abstract}  %% ~~ \textcolor{red}{\textbf{(To be revised)}}
+\textbf{Abstract}  
 \end{center}
@@
 such as sensitivity and stability.
-In this note, we examine the relationships between NCFs, symmetric 
+The purpose of this note is to bring out some relationships between NCFs, symmetric 
 Boolean functions and a more general form of symmetric Boolean functions,
-called $r$-symmetric functions (where $r$ is the symmetry level).
+which we call $r$-symmetric functions (where $r$ is the symmetry level).
 Using a standard representation for NCFs, we develop a 

### 2019-03-05 IEEE_TC_submission.d/intro.tex
ssr

@@
 \noindent 
-with $x_{\pi(i)}$ being the \textbf{canalyzing variable} that is
+where $x_{\pi(i)}$ is the \textbf{canalyzing variable} that is
 \textbf{tested} in line $i$, 
-and $a_i$ and $b_i$ being respectively the canalyzing and 
+and $a_i$ and $b_i$ are respectively the canalyzing and 
 canalyzed values in line $i$,~ $1 \leq i \leq n$.
@@
 (without loss of generality) that each NCF is specified in this manner.
-%% The simplified representation provides the following 
-%% convenient computational view of an NCF.~
-%% Lines defining an NCF are 
-%% considered sequentially in a top-down manner.
-%% The computation stops at the first line where the 
-%% specified condition is satisfied, and the value of the function
-%% is the canalyzed value on that line. 
+The simplified representation provides the following simple
+computational view of an NCF. 
+The lines that define an NCF are 
+considered sequentially in a top-down manner.
+The computation stops at the first line where the 
+specified condition is satisfied, and the value of the function
+is the canalyzed value on that line. 
 We now present an example of an NCF using the two representations
@@
              x_{2} ~\neq~ 1 \mathrm{~~and~~}\\ 
-           & x_{3} ~\neq~ 0 \\ \\ 
+           & x_{3} ~\neq~ 0 \\ 
    \end{cases}
@@
 \begin{tabular}{ll}
-\hspace*{0.5in} & $x_1:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
-\hspace*{0.5in} & $x_2:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
-\hspace*{0.5in} & $x_3:~$  $0 ~\longrightarrow~ 1$ \\ [0.5ex]
-\hspace*{0.5in} & Default:~ $0$ \\
+\hspace*{0.25in} & $x_1:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
+\hspace*{0.25in} & $x_2:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
+\hspace*{0.25in} & $x_3:~$  $0 ~\longrightarrow~ 1$ \\ [0.5ex]
+\hspace*{0.25in} & Default:~ $0$ \\
 \end{tabular}
@@
 As a simple example, each pair of variables in the function 
-$f_2(x_1, x_2, x_3)$ = $x_1 \oplus x_2 \oplus x_3$ is symmetric.
+$f_2(x_1, x_2, x_3)$ = $x_1 \oplus x_2 \oplus x_3$ is symmetric,
+where `$\oplus$' represents the Exclusive-Or operator.
 This form of interchange symmetry partitions the set of variables into a set of
@@
 Other forms of symmetry which can be more general than the permutations 
-corresponding to symmetry groups have been considered \cite{KS-2000}.
+corresponding to symmetry groups have been considered (see e.g.,
+\cite{Maurer-2015,KS-2000}).
 A Boolean function $f(x_1, x_2, \ldots, x_n)$
@@
 NCFs and symmetric Boolean functions.
-Using a standard representation for NCFs, we develop a
+Using a normalized representation for NCFs 
+(defined in Section~\ref{sec:prelim}), we develop a
 characterization of when two variables of an NCF are symmetric.
@@
 that the symmetry level of an NCF $f$
-can be easily computed given a standard representation of $f$.
+can be easily computed given a normalized representation of $f$.
 (In contrast, we show that 

### 2019-03-05 IEEE_TC_submission.d/preliminaries.tex
ssr

@@
 \label{def:layer}
-A {\bf layer} of a NCF representation is a maximal length sequence 
+A {\bf layer} of an NCF representation is a maximal length sequence 
 of lines with the same canalyzed value.
@@
 \begin{tabular}{ll}
-\hspace*{1.1in} & $x_1:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
-\hspace*{1.1in} & $x_2:~$  $0 ~\longrightarrow~ 0$ \\ [0.5ex]
-\hspace*{1.1in} & $x_3:~$  $0 ~\longrightarrow~ 0$ \\ [0.5ex]
-\hspace*{1.1in} & $x_4:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
-\hspace*{1.1in} & $x_5:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
-\hspace*{1.1in} & $x_6:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
-\hspace*{1.1in} & Default:~ $1$ \\
+\hspace*{0.25in} & $x_1:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
+\hspace*{0.25in} & $x_2:~$  $0 ~\longrightarrow~ 0$ \\ [0.5ex]
+\hspace*{0.25in} & $x_3:~$  $0 ~\longrightarrow~ 0$ \\ [0.5ex]
+\hspace*{0.25in} & $x_4:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
+\hspace*{0.25in} & $x_5:~$  $1 ~\longrightarrow~ 1$ \\ [0.5ex]
+\hspace*{0.25in} & $x_6:~$  $1 ~\longrightarrow~ 0$ \\ [0.5ex]
+\hspace*{0.25in} & Default:~ $1$ \\
 \end{tabular}
@@
 In view of Observation~\ref{obs:normalization_poly},
-we will assume that any NCF is specified in default-normalized form.
+we will assume from now on that any NCF is specified in default-normalized form.
 
@@
 new sets of variables, with each set containing $\rho+1$ variables.
-The expression for $f(x_1, \ldots, x_n, y_1, \ldots, y_{\rho+1})$
-is 
+The expression for $f$, which is a function of $n+2\rho+2$
+variables, namely $x_1, \ldots, x_n$, $y_1, \ldots, y_{\rho+1}$,
+$z_1, \ldots, z_{\rho+1}$, is the following:
 \[
-    g(x_1, \ldots x_n) \wedge (y_1 \vee \overline{z_1}) \wedge \ldots \wedge
+    g(x_1, \ldots x_n) \wedge (y_1 \vee \overline{z_1}) 
+                       \wedge (y_2 \vee \overline{z_2}) 
+                       \wedge \ldots \wedge
                               (y_{\rho+1} \vee \overline{z_{\rho+1}}).
 \]
-We note that $f$ is also a CNF formula.
+Since $g$ is a CNF formula, so is $f$.
 We have the following claims.
@@
 Since there are $\rho+1$ variables in $Y$, the number of symmetry groups 
-for $f$ must be at least $\rho+1$. \hfill$\Box$
+for $f$ must be at least $\rho+1$, and this 
+completes our proof of Claim~2. \hfill$\Box$
 
@@
 
-Now, suppose we execute Algorithm \cala{} on the function $f$.
+We now continue with the proof of Theorem~\ref{thm:approx_sym_level_hard}.
+Suppose we execute the approximation algorithm \cala{} 
+on the function $f$ defined above.
 If $g$ is \emph{not} satisfiable, then from Claim~1, 

### 2019-03-05 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@
 For the converse, consider the default-normalized NCF representation of $f$. 
-For purposes of notational simplicity, we assume without loss of generality that the
+As mentioned earlier, we assume without loss of generality that the
 canalyzing variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.
@@
 \noindent
-Consider the following input assignment $\alpha = (c_1, c_2, \ldots,  c_n)$ 
-to the variables of $f$.
+Consider the following assignment $\alpha$ = $(c_1, c_2, \ldots,  c_n)$ 
+to the variables $x_1$, $x_2$, $\ldots$, $x_n$ of $f$.
 \begin{enumerate}
@@
 \item For $j < i < k$, and for  $k < i \leq n$, 
-if $b_i = b_j$, then set $c_i =  \overline{a_i}$,
+if $b_i = b_j$,\\ then set $c_i =  \overline{a_i}$,
 else set $c_i =  a_i$.
@@
 by interchanging the values of variables $x_j$ and $x_k$, so that
-after the interchange variable $x_j$ has value $\overline{a_j}$ and
+after the interchange, variable $x_j$ has value $\overline{a_j}$ and
 variable $x_k$ has value $a_j$.  We will show that $f(\alpha') =
@@
 \textbf{Proof:}~ 
-If an NCF is $(n-1)$-symmetric, then consider a permutation than
+If an NCF is $(n-1)$-symmetric, then consider a permutation that
 interchanges two variables from a symmetry group with at least two
@@
 $\pi(i) = i$,~ $\pi(k) > k$,~ and for $k < i \leq n$,~ $\pi(i) \geq k$.
+
+\smallskip
 
@@
 containing line $k'$.  Thus, for all $i$, $c_{\pi(i)}$ does not
-match the canalyzing value of line $i$, so $f(c_{\pi(1)}, c_{\pi(2)},
-\ldots, c_{\pi(n)}) = \overline{b_k}$.  \QED
+match the canalyzing value of line $i$; therefore, 
+$f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)})$ ~=~ $\overline{b_k}$.  \QED
 
@@
 Consider a default-normalized representation of a strongly asymmetric NCF $f$.
-The representation contains $n-1$ layers.
+The representation has $n-1$ layers.
 The first $n-2$ layers each contain one line, and the last layer contains two lines.
@@
 containing that variable can be either 0 or 1.  Thus, there are
-$2^n$ possibilities for canalyzing values.  \QED
+$2^n$ possibilities for canalyzing values.
+
+Hence, the number of Boolean functions with $n$ variables that
+are both strongly asymmetric and NCFs is equal to \\ 
+$2 \times (n!/2) \times 2^n$ ~=~ $n!\,2^n$.  \QED
 

### 2019-03-05 IEEE_TC_submission.d/summary.tex
ssr

\section{Summary}
\label{sec:summary}

We presented a characterization of when two variables
of an NCF are symmetric and used that characterization
to show that the symmetry level of an NCF can be easily computed.
We also showed that an $n$-variable NCF is strongly asymmetric
iff its symmetry level is $n$.
We presented several corollaries of this result, including
a closed form expression for the number of NCFs which are also
strongly asymmetric.
Thus, our brings out several interesting relationships 
between NCFs and symmetric Boolean functions.

### 2019-03-05 IEEE_TC_submission.d/abstract.tex
ssr

@@
 has been used to model certain biological phenomena.
-Researchers have studied some properties of these functions
+Researchers have studied various properties of these functions
 such as sensitivity and stability.
 The purpose of this note is to bring out some relationships between NCFs, symmetric 
-Boolean functions and a more general form of symmetric Boolean functions,
+Boolean functions and a generalization of symmetric Boolean functions,
 which we call $r$-symmetric functions (where $r$ is the symmetry level).
-Using a standard representation for NCFs, we develop a 
+Using a normalized representation for NCFs, we develop a 
 characterization of when two variables of an NCF are symmetric.
@@
 that are NCFs and strongly asymmetric.
-In addition, we identify all the Boolean functions that are NCFs 
+We also identify all the Boolean functions that are NCFs 
 and symmetric.

### 2019-03-05 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@
 For any $n \geq 2$, the number of Boolean functions with $n$ variables 
-that are both strongly asymmetric and NCFs is $n! \, 2^n$.
+that are both strongly asymmetric and NCFs is $n! \, 2^{n-1}$.
 \end{theorem}
@@
 line.  The last layer contains the remaining two variables.  Thus,
-there are $n! / 2$ possibilities for the pattern of variable
+there are $n! /[n-(n-2)]!$ = $n!/2$ possibilities for the pattern of variable
 occurrences in $f$.
 
-For each of the $n$ variables, the canalyzing value for the line
-containing that variable can be either 0 or 1.  Thus, there are
-$2^n$ possibilities for canalyzing values.
+For each of the first $n-2$ variables, the canalyzing value for the line
+containing that variable can be either 0 or 1.  
+%% The last two lines, the canalyzed values on those two lines must be
+%% the same because of the default-normalized representation. 
+For the last two lines, the canalyzing values for the variables 
+in those lines must be complements of each other; otherwise, by 
+Theorem~\ref{thm:ncf_symmetric_variables}, those two variables 
+will be symmetric.
+Thus, the number of possible canalyzing values over all the $n$
+lines is $2^{n-2} \times 2$ = $2^{n-1}$.
 
@@
 are both strongly asymmetric and NCFs is equal to \\ 
-$2 \times (n!/2) \times 2^n$ ~=~ $n!\,2^n$.  \QED
+$2 \times (n!/2) \times 2^{n-1}$ ~=~ $n!\,2^{n-1}$.  \QED
 

### 2019-03-22 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@
 \end{corollary}
+
+\subsection{Testing Whether an $r$-Symmetric Function is an NCF}
+\label{sse:rsym_to_ncf}
+
+Theorem~\ref{thm:ncf_r_symmetric} shows that given an NCF $f$,
+the problem of finding the smallest integer $r$ such that
+$f$ is $r$-symmetric can be solved efficiently.
+We now show that the converse problem, that is, determining whether a given
+$r$-symmetric Boolean function is an NCF,
+can also be solved efficiently.
+When the answer is ``yes", our algorithm also constructs a
+default-normalized representation of $f$.
+
+Let $f(x_1, x_2, \ldots, x_n)$ be an $r$-symmetric Boolean function of $n$
+Boolean variables.
+We assume that the symmetry groups $g_i$, $1 \leq i \leq r$,
+of $f$ are given.
+Let $m_i$ denote the number of variables
+in the group containing variable $x_i$, $1 \leq i \leq m$.
+Thus, in the group containing $x_i$, the number
+of variables which can take on the value 1 varies from 0
+to $m_i$, $1 \leq i \leq n$.
+We also assume that $f$ is
+given as a table $T$ with $\mu = \prod_{i=1}^r (m_i+1)$ rows.
+Each row of $T$ specifies an $r$-tuple $(c_1, c_2, \ldots, c_r)$,
+where $c_i$ is the number of variables of group $g_i$ which
+have the value 1, along with the \{0,1\} value of $f$ for that $r$-tuple.
+As will be seen, our algorithm for determining whether $f$ is
+an NCF runs in $O(r^2\mu)$ time.
+The algorithm relies on the following observation.
+
+\begin{observation}\label{obs:rsym_canalyzing}
+Suppose $f(x_1, x_2, \ldots, x_n)$ is an $r$-symmetric Boolean function.
+Consider any variable $x_i$ and suppose
+the number of variables in the group
+containing $x_i$ is $m_i$, $1 \leq i \leq r$.
+\begin{enumerate}
+\item The function $f$ is canalyzing in variable $x_i$
+with canalyzing value 1 iff all the table entries where the number
+of 1's in the group containing $x_i$ is nonzero have the same value
+of $f$. (This value of $f$ is the canalyzed value when $x_i = 1$.)
+\item The function $f$ is canalyzing in variable $x_i$ with
+canalyzing value 0 iff all the table entries where the number of
+1's in the group containing $x_i$ is less than $m_i$ have the same
+value of $f$. (This value of $f$ is the canalyzed value when $x_i = 0$.)
+\end{enumerate}
+\end{observation}
+We now specify an iterative algorithm for determining whether 
+$f$ is an NCF, and if so, constructing a simplified 
+NCF representation (defined in Section~\ref{sec:intro}) for $f$.
+In iteration $j$, the algorithm uses a set $X_j$ of remaining variables,
+table $T_j$ with $\mu_j$ rows, and the number of groups $g_j$.
+Initially (i.e., $j = 1$), $X_1$ = $\{x_1, \ldots, x_n\}$,
+$T_1$ is the given table $T$ for $f$, 
+$\mu_1 = \mu$ (the number of rows of $T$) and $g_j = r$ (the number of
+symmetry groups of $f$) and the NCF representation for $f$ is empty.
+The algorithm executes as long as one or more symmetry groups remain.
+In each iteration, the algorithm carries out the following steps.
+
+\smallskip
+\noindent
+I. Use Observation~\ref{obs:rsym_canalyzing} to determine whether there is
+a variable $x_i \in X_j$ is canalyzing.
+
+\smallskip
+\noindent
+II. If yes, perform the following steps. 
+\begin{enumerate}
+\item Let $\alpha$ and $\beta$ be the respective canalyzing and 
+canalyzed values for $x_i$ found in Step~I. 
+\item For each variable $x_p$ in the group containing $x_i$, add the rule~
+``$x_p~:~\alpha ~\longrightarrow~ \beta$"~ to the NCF representation of $f$.
+\item Obtain $X_{j+1}$ by deleting from $X_j$ all the variables which are 
+in the same group as $x_i$.
+\item Set $g_{j+1} = g_j - 1$.
+\item Obtain $T_{j+1}$ by retaining only those rows of $T_j$ where all the
+variables in the group containing $x_i$ are set to $\overline{\alpha}$. 
+\item Start the next iteration.
+\end{enumerate}
+
+\smallskip
+\noindent
+III. (Here, Step~I didn't find a canalyzing variable.)~ 
+Output ``$f$ is not an NCF" and \textbf{stop}.
+
+\smallskip
+\noindent
+The correctness of the algorithm is a direct consequence of
+Observation~\ref{obs:rsym_canalyzing}. If the algorithm is successful,
+it produces a simplified NCF representation of $f$; this can then be
+efficiently converted into the default-normalized representation as
+discussed in Section~\ref{sse:ncf_layer}.
+
+To estimate the running time, we note that in iteration $j$, we can
+determine in $O(g_j\,\mu_j)$ = $O(r\,\mu)$ time using 
+Observation~\ref{obs:rsym_canalyzing}
+whether there is a variable that is canalyzing.
+(This can be done by considering one arbitrary variable 
+from each group.) 
+Once a variable is identified, the other steps in that iteration
+can be completed in $O(\mu_j)$ = $O(\mu)$ time.
+Thus, the time for each iteration is $O(r\,\mu)$.
+Since there are at most $r$ iterations, the running time of the
+algorithm is $O(r^2\,\mu)$.
+The following theorem summarizes this result.
+
+\begin{theorem}\label{thm:rsym_canalyzing}
+Given an $r$-symmetric function $f$, the problem of testing whether
+$f$ is an NCF can be solved efficiently. \QED
+\end{theorem}
 
@@
 line.  The last layer contains the remaining two variables.  Thus,
-there are $n! /[n-(n-2)]!$ = $n!/2$ possibilities for the pattern of variable
-occurrences in $f$.
+there are $n! /[n-(n-2)]!$ = $n!/2$ possibilities for the pattern of 
+canalyzing variable occurrences in $f$.
 

### 2019-03-22 IEEE_TC_submission.d/abstract.tex
ssr

@@
-\vspace*{-0.05in}
-\begin{center}
-\textbf{Abstract}  
-\end{center}
-
-\smallskip
-
 A class of Boolean functions, 
@@
 a general Boolean function to within any factor $\geq 1$ is \cnp-hard.) 
-We also show that for any NCF $f$ with $n$ variables, the notion of
+We also present an efficient algorithm for testing whether 
+a given $r$-symmetric function is an NCF.
+Further, we show that for any NCF $f$ with $n$ variables, the notion of
 strong asymmetry considered in the literature is equivalent to

### 2019-03-22 IEEE_TC_submission.d/intro.tex
ssr

@@
 unless \textbf{P} = \cnp.)
-We also show that for any NCF $f$ with $n$ variables, the notion of
+We also present an efficient algorithm to test whether a given
+$r$-symmetric function $f$ is an NCF. 
+Further, we show that for any NCF $f$ with $n$ variables, the notion of
 strong asymmetry considered in the literature is equivalent to

### 2019-03-22 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@
 its default-normalized representation.
-%% We also list some corollaries that follow from the
-%% characterization.
+We also present an efficient algorithm (Theorem~\ref{thm:rsym_canalyzing})
+for the converse problem, that is,
+testing whether a given $r$-symmetric function $f$ is an NCF;
+if so, our algorithm constructs a default-normalized 
+representation for $f$.
 Next, we show that for any NCF $f$ with $n$ variables,
@@
 $f$ is $r$-symmetric can be solved efficiently.
-We now show that the converse problem, that is, determining whether a given
+We now show that the converse problem, that is, testing whether a given
 $r$-symmetric Boolean function is an NCF,
@@
 Let $f(x_1, x_2, \ldots, x_n)$ be an $r$-symmetric Boolean function of $n$
-Boolean variables.
+variables.
 We assume that the symmetry groups $g_i$, $1 \leq i \leq r$,
@@
 \end{observation}
-We now specify an iterative algorithm for determining whether 
-$f$ is an NCF, and if so, constructing a simplified 
-NCF representation (defined in Section~\ref{sec:intro}) for $f$.
-In iteration $j$, the algorithm uses a set $X_j$ of remaining variables,
-table $T_j$ with $\mu_j$ rows, and the number of groups $g_j$.
-Initially (i.e., $j = 1$), $X_1$ = $\{x_1, \ldots, x_n\}$,
-$T_1$ is the given table $T$ for $f$, 
-$\mu_1 = \mu$ (the number of rows of $T$) and $g_j = r$ (the number of
+We now explain how Observation~\ref{obs:rsym_canalyzing} can be
+used to develop an iterative algorithm for determining whether 
+$f$ is an NCF, and if so, constructing a default-normalized 
+representation (defined in Section~\ref{sec:prelim}) for $f$.
+At beginning of iteration $j$, the algorithm uses a set $X_j$ of remaining variables,
+table $T_j$ with $\mu_j$ rows and the number of remaining groups is $g_j$.
+Initially (i.e., $j = 1$), $X_1$ = $X$ = $\{x_1, \ldots, x_n\}$,
+$T_1 = T$ (the given table $T$ for $f$), 
+$\mu_1 = \mu$ (the number of rows of $T$), $g_j = r$ (the number of
 symmetry groups of $f$) and the NCF representation for $f$ is empty.
-The algorithm executes as long as one or more symmetry groups remain.
-In each iteration, the algorithm carries out the following steps.
+The algorithm carries out iteration $j$ if $g_j \geq 1$; 
+in that iteration, the algorithm performs Steps I, II and III
+as described below.
 
@@
 I. Use Observation~\ref{obs:rsym_canalyzing} to determine whether there is
-a variable $x_i \in X_j$ is canalyzing.
+a canalyzing variable $x_i \in X_j$.
 
@@
 
-\smallskip
+\medskip
 \noindent
@@
 To estimate the running time, we note that in iteration $j$, we can
-determine in $O(g_j\,\mu_j)$ = $O(r\,\mu)$ time using 
+determine in $O(g_j \mu_j)$ = $O(r \mu)$ time using 
 Observation~\ref{obs:rsym_canalyzing}
@@
 can be completed in $O(\mu_j)$ = $O(\mu)$ time.
-Thus, the time for each iteration is $O(r\,\mu)$.
+Thus, the time for each iteration is $O(r \mu)$.
 Since there are at most $r$ iterations, the running time of the
-algorithm is $O(r^2\,\mu)$.
+algorithm is $O(r^2 \mu)$.
 The following theorem summarizes this result.
@@
 \begin{theorem}\label{thm:rsym_canalyzing}
-Given an $r$-symmetric function $f$, the problem of testing whether
-$f$ is an NCF can be solved efficiently. \QED
+The problem of testing whether a given 
+$r$-symmetric function $f$ is an NCF can 
+be solved efficiently. \QED
 \end{theorem}

### 2019-03-22 IEEE_TC_submission.d/intro.tex
ssr

@@
 
-\IEEEPARstart{T}{he} class of \textbf{canalyzing} Boolean functions, introduced 
-in \cite{Kauffman-1969}, is defined as follows.
+\IEEEPARstart{R}{search} on the symmetry properties of Boolean
+functions has received a lot of attention from the
+automated logic synthesis community. 
+(Additional discussion on this is provided in Section~\ref{sse:related}.)
+In this paper, we focus on the symmetry properties of
+a class of Boolean functions (called \textbf{nested canalyzing functions})
+which are used to model biological phenomena.
+To define this class, we start with a simpler
+class of functions, called \textbf{canalyzing} Boolean functions.
+This class, introduced in \cite{Kauffman-1969}, is defined as follows.
 
@@
 
+Over the years, a considerable amount of research on detecting symmetry properties 
+of Boolean functions has been reported in the literature
+(e.g., \cite{BS-1968,Biswas-1970,TM-1996,KS-2000,Darga-etal-2008,Maurer-2015}).
+Several references have observed the usefulness of detecting symmetries
+in automated logic synthesis (e.g., \cite{AP-2008,Hu-etal-2008,Darga-etal-2008}).
+Reference \cite{Maurer-2015} provides a thorough discussion of known
+symmetry detection algorithms and presents a new universal algorithm 
+for detecting any form of permutation-based symmetry in Boolean functions.
+
 The term \textbf{canalization}, coined by
@@
 
-Over the years, a considerable amount of research on detecting symmetry properties 
-of Boolean functions has been reported in the literature
-(e.g., \cite{BS-1968,Biswas-1970,TM-1996,KS-2000,Darga-etal-2008,Maurer-2015}).
-Several references have observed the usefulness of detecting symmetries
-in automated logic synthesis (e.g., \cite{AP-2008,Hu-etal-2008,Darga-etal-2008}).
-Reference \cite{Maurer-2015} provides a thorough discussion of known
-symmetry detection algorithms and presents a new universal algorithm 
-for detecting any form of permutation-based symmetry in Boolean functions. 
 To our knowledge, relationships between NCFs and symmetric

### 2019-03-22 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@
 \begin{corollary}\label{cor:ncf_not_rsymm}
-For every $n \geq 2$, there is an $n$-variable NCF that is not $n-1$ symmetric.
+For every $n \geq 2$, there is an $n$-variable NCF that is not $n-1$ symmetric. \QED
 \end{corollary}
@@
 with at most $r-1$ layers, and has an unnormalized nested canalyzing
-representation with at most $r$ layers.  
+representation with at most $r$ layers.   \QED
 \end{corollary}
@@
 As will be seen, our algorithm for determining whether $f$ is
-an NCF runs in $O(r^2\mu)$ time.
+an NCF runs in $O(r\mu)$ time.
 The algorithm relies on the following observation.
@@
 value of $f$. (This value of $f$ is the canalyzed value when $x_i = 0$.)
+\QED
 \end{enumerate}
@@
 To estimate the running time, we note that in iteration $j$, we can
-determine in $O(g_j \mu_j)$ = $O(r \mu)$ time using 
+determine in $O(g_j \mu_j)$ = $O(r \mu_j)$ time using 
 Observation~\ref{obs:rsym_canalyzing}
-whether there is a variable that is canalyzing.
-(This can be done by considering one arbitrary variable 
+whether there is a canalyzing variable.
+(This can be done by considering an arbitrary variable 
 from each group.)
 Once a variable is identified, the other steps in that iteration
-can be completed in $O(\mu_j)$ = $O(\mu)$ time.
-Thus, the time for each iteration is $O(r \mu)$.
-Since there are at most $r$ iterations, the running time of the
-algorithm is $O(r^2 \mu)$.
+can be completed in $O(\mu_j)$ time.
+Thus, the time for iteration $j$ is $O(r \mu_j)$.
+We note that in each iteration, the number of rows in the table
+is reduced by at least a factor of 2; that is, $\mu_{j+1} \leq \mu_j/2$.
+To see this, suppose the canalyzing variable $x_i$ found in iteration $j$
+is in group $\alpha_p$, which has $m_p$ variables. 
+Thus, in $T_j$, there are $m_p+1 \geq 2$ possibilities for the number of 1's 
+in $\alpha_p$. 
+As indicated in Substep~5 of Step~II above, in $T_{j+1}$, 
+we only retain those rows of $T_j$ where all the variables 
+in $\alpha_p$ have the same value.
+In other words, there is only one possibility for the number of
+1's in group $\alpha_p$.
+Thus, $\mu_{j+1}$, the number of rows in $T_{j+1}$,
+is at most $\mu_j/(m_p+1)$ $\leq$ $\mu_j/2$.
+It follows by simple induction that $\mu_j \leq \mu/2^{j-1}$.
+Thus, the running time over all the $r$ iterations is bounded by
+$O(r [\sum_{j=1}^{r}(\mu/2^{j-1})])$ = $O(r\mu)$.
 The following theorem summarizes this result.

### 2019-03-22 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@
 
-Let $f(x_1, x_2, \ldots, x_n)$ be an $r$-symmetric Boolean function of $n$
-variables.
+Let $f(x_1, x_2, \ldots, x_n)$ be an $r$-symmetric 
+Boolean function of $n$ variables.
 We assume that the symmetry groups $g_i$, $1 \leq i \leq r$,
 of $f$ are given.
-Let $m_i$ denote the number of variables
-in the group containing variable $x_i$, $1 \leq i \leq m$.
-Thus, in the group containing $x_i$, the number
+Let $m_i = |g_i|$, $1 \leq i \leq m$.
+Thus, in group $g_i$, the number
 of variables which can take on the value 1 varies from 0
-to $m_i$, $1 \leq i \leq n$.
+to $m_i$, $1 \leq i \leq r$.
 We also assume that $f$ is
-given as a table $T$ with $\mu = \prod_{i=1}^r (m_i+1)$ rows.
-Each row of $T$ specifies an $r$-tuple $(c_1, c_2, \ldots, c_r)$,
+given as a table $T$ of the following form:
+each row of $T$ specifies an $r$-tuple $(c_1, c_2, \ldots, c_r)$,
 where $c_i$ is the number of variables of group $g_i$ which
 have the value 1, along with the \{0,1\} value of $f$ for that $r$-tuple.
+Thus, $\mu$, the number of rows in $T$, 
+is given by  $\mu = \prod_{i=1}^r (m_i+1)$.
 As will be seen, our algorithm for determining whether $f$ is
@@
 the number of variables in the group
-containing $x_i$ is $m_i$, $1 \leq i \leq r$.
+containing $x_i$ is $\alpha_i$, $1 \leq i \leq n$.
 \begin{enumerate}
@@
 used to develop an iterative algorithm for determining whether 
-$f$ is an NCF, and if so, constructing a default-normalized 
+$f$ is an NCF; if so, the algorithm also constructs a default-normalized 
 representation (defined in Section~\ref{sec:prelim}) for $f$.
-At beginning of iteration $j$, the algorithm uses a set $X_j$ of remaining variables,
-table $T_j$ with $\mu_j$ rows and the number of remaining groups is $g_j$.
+At beginning of iteration $j$, the algorithm has the set $X_j$ of remaining variables,
+table $T_j$ with $\mu_j$ rows and $\ell_j$ remaining groups.
 Initially (i.e., $j = 1$), $X_1$ = $X$ = $\{x_1, \ldots, x_n\}$,
 $T_1 = T$ (the given table $T$ for $f$), 
-$\mu_1 = \mu$ (the number of rows of $T$), $g_j = r$ (the number of
+$\mu_1 = \mu$ (the number of rows of $T$), $\ell_j = r$ (the number of
 symmetry groups of $f$) and the NCF representation for $f$ is empty.
-The algorithm carries out iteration $j$ if $g_j \geq 1$; 
+The algorithm carries out iteration $j$ if $\ell_j \geq 1$; 
 in that iteration, the algorithm performs Steps I, II and III
@@
 canalyzed values for $x_i$ found in Step~I. 
-\item For each variable $x_p$ in the group containing $x_i$, add the rule~
+\item For each variable $x_p$ in the group containing $x_i$, append the rule~
 ``$x_p~:~\alpha ~\longrightarrow~ \beta$"~ to the NCF representation of $f$.
@@
 in the same group as $x_i$.
-\item Set $g_{j+1} = g_j - 1$.
+\item Set $\ell_{j+1} = \ell_j - 1$.
 \item Obtain $T_{j+1}$ by retaining only those rows of $T_j$ where all the
@@
 To estimate the running time, we note that in iteration $j$, we can
-determine in $O(g_j \mu_j)$ = $O(r \mu_j)$ time using 
-Observation~\ref{obs:rsym_canalyzing}
-whether there is a canalyzing variable.
+determine whether there is a canalyzing variable in $X_j$
+in $O(\ell_j \mu_j)$ = $O(r \mu_j)$ time using 
+Observation~\ref{obs:rsym_canalyzing}.
 (This can be done by considering an arbitrary variable 
@@
 is reduced by at least a factor of 2; that is, $\mu_{j+1} \leq \mu_j/2$.
-To see this, suppose the canalyzing variable $x_i$ found in iteration $j$
-is in group $\alpha_p$, which has $m_p$ variables. 
+To see why, suppose the canalyzing variable $x_i$ found in iteration $j$
+is in group $g_p$ with $m_p$ variables. 
 Thus, in $T_j$, there are $m_p+1 \geq 2$ possibilities for the number of 1's 
-in $\alpha_p$. 
-As indicated in Substep~5 of Step~II above, in $T_{j+1}$, 
-we only retain those rows of $T_j$ where all the variables 
-in $\alpha_p$ have the same value.
+in group $g_p$. 
+As indicated in Substep~5 of Step~II above, $T_{j+1}$ 
+contains only those rows of $T_j$ where all the variables 
+in $g_p$ have the same value.
 In other words, there is only one possibility for the number of
-1's in group $\alpha_p$.
+1's in group $g_p$.
 Thus, $\mu_{j+1}$, the number of rows in $T_{j+1}$,
@@
 It follows by simple induction that $\mu_j \leq \mu/2^{j-1}$.
-Thus, the running time over all the $r$ iterations is bounded by
+Thus, the running time over all the $r$ iterations is given by
 $O(r [\sum_{j=1}^{r}(\mu/2^{j-1})])$ = $O(r\mu)$.

### 2019-03-24 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@
 can also be solved efficiently.
-When the answer is ``yes", our algorithm also constructs a
-default-normalized representation of $f$.
+When the answer is ``yes", a
+default-normalized representation (defined in Section~\ref{sec:prelim}
+of $f$ can also be constructed efficiently.
 
@@
 \begin{enumerate}
-\item The function $f$ is canalyzing in variable $x_i$
+\item Function $f$ is canalyzing in variable $x_i$
 with canalyzing value 1 iff all the table entries where the number
@@
 of $f$. (This value of $f$ is the canalyzed value when $x_i = 1$.)
-\item The function $f$ is canalyzing in variable $x_i$ with
+\item Function $f$ is canalyzing in variable $x_i$ with
 canalyzing value 0 iff all the table entries where the number of
-1's in the group containing $x_i$ is less than $m_i$ have the same
+1's in the group containing $x_i$ is less than $\alpha_i$ have the same
 value of $f$. (This value of $f$ is the canalyzed value when $x_i = 0$.)
@@
 $f$ is an NCF; if so, the algorithm also constructs a default-normalized 
-representation (defined in Section~\ref{sec:prelim}) for $f$.
+representation for $f$.
 At beginning of iteration $j$, the algorithm has the set $X_j$ of remaining variables,
@@
 \item For each variable $x_p$ in the group containing $x_i$, append the rule~
-``$x_p~:~\alpha ~\longrightarrow~ \beta$"~ to the NCF representation of $f$.
+``$x_p\,:\,\alpha ~\longrightarrow~ \beta$"~ to the NCF representation of $f$.
 \item Obtain $X_{j+1}$ by deleting from $X_j$ all the variables which are 
@@
 from each group.)
-Once a variable is identified, the other steps in that iteration
+Once a canalyzing variable is identified, the other steps in that iteration
 can be completed in $O(\mu_j)$ time.
@@
 Thus, the running time over all the $r$ iterations is given by
-$O(r [\sum_{j=1}^{r}(\mu/2^{j-1})])$ = $O(r\mu)$.
+$O(r [\sum_{j=1}^{r}(\mu/2^{j-1})])$ = $O(r\mu)$, since the
+geometric sum is bounded by $2\mu$.
 The following theorem summarizes this result.

### 2019-03-25 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@
 When the answer is ``yes", a
-default-normalized representation (defined in Section~\ref{sec:prelim}
+default-normalized representation (defined in Section~\ref{sec:prelim})
 of $f$ can also be constructed efficiently.
@@
 Boolean function of $n$ variables.
-We assume that the symmetry groups $g_i$, $1 \leq i \leq r$,
-of $f$ are given.
+We assume that the symmetry groups 
+of $f$, denoted by $g_1$, $g_2$, $\ldots$, $g_r$,
+are given.
 Let $m_i = |g_i|$, $1 \leq i \leq m$.
@@
 the number of variables in the group
-containing $x_i$ is $\alpha_i$, $1 \leq i \leq n$.
+containing $x_i$ is $\nu_i$, $1 \leq i \leq n$.
 \begin{enumerate}
@@
 canalyzing value 0 iff all the table entries where the number of
-1's in the group containing $x_i$ is less than $\alpha_i$ have the same
+1's in the group containing $x_i$ is less than $\nu_i$ have the same
 value of $f$. (This value of $f$ is the canalyzed value when $x_i = 0$.)
@@
 representation for $f$.
-At beginning of iteration $j$, the algorithm has the set $X_j$ of remaining variables,
-table $T_j$ with $\mu_j$ rows and $\ell_j$ remaining groups.
+We use the following notation.
+At beginning of iteration $j$, $X_j$ denotes the set of remaining variables,
+$T_j$ denotes the table which provides values for the function,
+$\mu_j$ denotes the number of rows of $T_j$ and $\ell_j$ 
+denotes the number of remaining groups.
 Initially (i.e., $j = 1$), $X_1$ = $X$ = $\{x_1, \ldots, x_n\}$,
 $T_1 = T$ (the given table $T$ for $f$), 
-$\mu_1 = \mu$ (the number of rows of $T$), $\ell_j = r$ (the number of
-symmetry groups of $f$) and the NCF representation for $f$ is empty.
+$\mu_1 = \mu$ (the number of rows of $T$) and $\ell_j = r$ (the number of
+symmetry groups of $f$); further, the NCF representation for $f$ is empty.
 The algorithm carries out iteration $j$ if $\ell_j \geq 1$; 

### 2019-03-27 IEEE_TC_submission.d/intro.tex
ssr

@@
 functions has received a lot of attention from the
-automated logic synthesis community. 
+logic synthesis community. 
 (Additional discussion on this is provided in Section~\ref{sse:related}.)
@@
 a 1-symmetric function $f$ is referred to simply
-as a \textbf{symmetric} function.
+as a \textbf{symmetric} function or 
+as a \textbf{totally symmetric function} \cite{Biswas-1970,BS-1968}.
 Thus, in any symmetric function, each pair of variables is symmetric.
@@
 Over the years, a considerable amount of research on detecting symmetry properties 
-of Boolean functions has been reported in the literature
-(e.g., \cite{BS-1968,Biswas-1970,TM-1996,KS-2000,Darga-etal-2008,Maurer-2015}).
+of Boolean functions has been reported in the literature.
+For example, references \cite{Biswas-1970,TM-1996} present methods 
+to determine whether a given Boolean function is symmetric.
+Born and Scidmore \cite{BS-1968} show how certain Boolean functions
+can be converted into symmetric functions by introducing additional
+variables. 
+%% (e.g., \cite{BS-1968,Biswas-1970,TM-1996,KS-2000,Darga-etal-2008,Maurer-2015}).
 Several references have observed the usefulness of detecting symmetries

### 2019-03-27 IEEE_TC_submission.d/intro.tex
ssr

@@
 
-Over the years, a considerable amount of research on detecting symmetry properties 
+The study of symmetry properties of Boolean functions
+was initiated by Shannon \cite{Shannon-1938}.
+Over the years, a considerable amount of research on detecting 
+and exploiting symmetry properties 
 of Boolean functions has been reported in the literature.
@@
 %% (e.g., \cite{BS-1968,Biswas-1970,TM-1996,KS-2000,Darga-etal-2008,Maurer-2015}).
-Several references have observed the usefulness of detecting symmetries
-in automated logic synthesis (e.g., \cite{AP-2008,Hu-etal-2008,Darga-etal-2008}).
+Several researchers have developed techniques for exploiting symmetries
+for automated logic synthesis 
+(e.g., \cite{AP-2008,Hu-etal-2008,KS-2000,Darga-etal-2008}).
 Reference \cite{Maurer-2015} provides a thorough discussion of known

### 2019-03-27 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@
 in the same group as $x_i$.
-\item Set $\ell_{j+1} = \ell_j - 1$.
 \item Obtain $T_{j+1}$ by retaining only those rows of $T_j$ where all the
 variables in the group containing $x_i$ are set to $\overline{\alpha}$. 
-\item Start the next iteration.
+\item Set $\ell_{j+1} = \ell_j - 1$.
+\item If $\ell_{j+1} \geq 1$, start the next iteration; 
+      otherwise, \textbf{stop}.
 \end{enumerate}
@@
 Observation~\ref{obs:rsym_canalyzing}.
-(This can be done by considering an arbitrary variable 
+(This is done by considering an arbitrary variable 
 from each group.)
@@
 We note that in each iteration, the number of rows in the table
-is reduced by at least a factor of 2; that is, $\mu_{j+1} \leq \mu_j/2$.
+is reduced by a factor of at least 2; that is, $\mu_{j+1} \leq \mu_j/2$.
 To see why, suppose the canalyzing variable $x_i$ found in iteration $j$
@@
 in group $g_p$. 
-As indicated in Substep~5 of Step~II above, $T_{j+1}$ 
+As indicated in Substep~4 of Step~II above, $T_{j+1}$ 
 contains only those rows of $T_j$ where all the variables 

### 2019-03-28 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@
 We use the following notation.
-At beginning of iteration $j$, $X_j$ denotes the set of remaining variables,
-$T_j$ denotes the table which provides values for the function,
-$\mu_j$ denotes the number of rows of $T_j$ and $\ell_j$ 
-denotes the number of remaining groups.
-Initially (i.e., $j = 1$), $X_1$ = $X$ = $\{x_1, \ldots, x_n\}$,
-$T_1 = T$ (the given table $T$ for $f$), 
-$\mu_1 = \mu$ (the number of rows of $T$) and $\ell_j = r$ (the number of
-symmetry groups of $f$); further, the NCF representation for $f$ is empty.
+At beginning of iteration $j$,~
+$\ell_j$ denotes the number of remaining groups,
+$X_j$ denotes a set of variables with
+exactly one variable from each remaining group,
+$T_j$ denotes the table which provides values for the function and
+$\mu_j$ denotes the number of rows of $T_j$.
+Initially (i.e., $j = 1$),~ $\ell_1 = r$ (the number of symmetry groups of $f$),
+$X_1$ is constructed by choosing one 
+variable (arbitrarily) from each of the $r$ groups,
+$T_1 = T$ (the given table $T$ for $f$) and 
+$\mu_1 = \mu$ (the number of rows of $T$);
+further, the NCF representation for $f$ is empty.
 The algorithm carries out iteration $j$ if $\ell_j \geq 1$; 
@@
 \item For each variable $x_p$ in the group containing $x_i$, append the rule~
-``$x_p\,:\,\alpha ~\longrightarrow~ \beta$"~ to the NCF representation of $f$.
-\item Obtain $X_{j+1}$ by deleting from $X_j$ all the variables which are 
-in the same group as $x_i$.
+``$x_p : \alpha ~\longrightarrow~ \beta$"~ to the NCF representation of $f$.
+\item Set $X_{j+1}$ = $X_j - \{x_i\}$.  
 \item Obtain $T_{j+1}$ by retaining only those rows of $T_j$ where all the
@@
 Observation~\ref{obs:rsym_canalyzing}.
-(This is done by considering an arbitrary variable 
-from each group.)
+(This is done by considering each variable in $X_j$.)
 Once a canalyzing variable is identified, the other steps in that iteration

### 2019-03-29 IEEE_TC_submission.d/abstract.tex
ssr

@@
+Many researchers have studied symmetry properties of 
+various Boolean functions. 
 A class of Boolean functions, 
@@
 has been used to model certain biological phenomena.
-Researchers have studied various properties of these functions
-such as sensitivity and stability.
-The purpose of this note is to bring out some relationships between NCFs, symmetric 
+The purpose of this brief contribution is to bring out 
+some interesting relationships between NCFs, symmetric 
 Boolean functions and a generalization of symmetric Boolean functions,
@@
 can be easily computed given a standard representation of $f$.
-(In contrast, we prove that even approximating the symmetry level of
-a general Boolean function to within any factor $\geq 1$ is \cnp-hard.) 
+%% (In contrast, we prove that even approximating the symmetry level of
+%% a general Boolean function to within any factor $\geq 1$ is \cnp-hard.) 
 We also present an efficient algorithm for testing whether 

### 2019-03-29 IEEE_TC_submission.d/intro.tex
ssr

@@
 functions has received a lot of attention from the
-logic synthesis community. 
+logic design and automated synthesis communities. 
 (Additional discussion on this is provided in Section~\ref{sse:related}.)
@@
        \:\overline{b_n} & \mathrm{if~~} x_{\pi(1)} ~\neq~ a_1 \mathrm{~~and~~} \ldots\\
-                        & x_{\pi(n)} ~\neq~ a_n \\
+                        & x_{\pi(n)} ~\neq~ a_n. \\
    \end{cases}
@@
 (without loss of generality) that each NCF is specified in this manner.
-The simplified representation provides the following simple
+The simplified representation provides the following 
 computational view of an NCF. 
@@
 For a Boolean function $f$, the integer $r$ such that $f$ is properly $r$-symmetric
-will be referred to as the \textbf{symmetry level} of $f$.
+is referred to as the \textbf{symmetry level} of $f$.
 Thus, the symmetry level of $f$ is the smallest integer $r$ such that
@@
 $f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
-The class of $r$-symmetric functions have also been
+For $r \geq 1$, the class of $r$-symmetric functions has been
 studied in the literature on discrete dynamical systems (see e.g., 
@@
 Other forms of symmetry which can be more general than the permutations 
-corresponding to symmetry groups have been considered (see e.g.,
+corresponding to symmetry groups have also been considered (see e.g.,
 \cite{Maurer-2015,KS-2000}).
@@
 
-The study of symmetry properties of Boolean functions
-was initiated by Shannon \cite{Shannon-1938}.
+The usefulness of symmetry properties in synthesizing 
+combinational logic functions was first noted 
+by Shannon \cite{Shannon-1938}.
 Over the years, a considerable amount of research on detecting 
@@
 (e.g., \cite{AP-2008,Hu-etal-2008,KS-2000,Darga-etal-2008}).
-Reference \cite{Maurer-2015} provides a thorough discussion of known
+Maurer \cite{Maurer-2015} provides a thorough discussion of known
 symmetry detection algorithms and presents a new universal algorithm 

### 2019-03-29 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@
 
-\subsection{Symmetric Pairs of Variables in an NCF: A Characterization
-and Its Applications}
+\subsection{Symmetric Pairs of Variables in an NCF:\newline 
+A Characterization and its Applications}
 \label{sse:ncf_strong_sym}
@@
 
-\subsection{Testing Whether an $r$-Symmetric Function is an NCF}
+\subsection{Testing Whether an $r$-Symmetric Function\newline is an NCF}
 \label{sse:rsym_to_ncf}
@@
 At beginning of iteration $j$,~
-$\ell_j$ denotes the number of remaining groups,
+$r_j$ denotes the number of remaining groups,
 $X_j$ denotes a set of variables with
@@
 $\mu_j$ denotes the number of rows of $T_j$.
-Initially (i.e., $j = 1$),~ $\ell_1 = r$ (the number of symmetry groups of $f$),
+Initially (i.e., $j = 1$),~ $r_1 = r$ (the number of symmetry groups of $f$),
 $X_1$ is constructed by choosing one 
@@
 further, the NCF representation for $f$ is empty.
-The algorithm carries out iteration $j$ if $\ell_j \geq 1$; 
+The algorithm carries out iteration $j$ if $r_j \geq 1$; 
 in that iteration, the algorithm performs Steps I, II and III
@@
 variables in the group containing $x_i$ are set to $\overline{\alpha}$. 
-\item Set $\ell_{j+1} = \ell_j - 1$.
-\item If $\ell_{j+1} \geq 1$, start the next iteration; 
+\item Set $r_{j+1} = r_j - 1$.
+\item If $r_{j+1} \geq 1$, start the next iteration; 
       otherwise, \textbf{stop}.
@@
 determine whether there is a canalyzing variable in $X_j$
-in $O(\ell_j \mu_j)$ = $O(r \mu_j)$ time using 
+in $O(r_j \mu_j)$ = $O(r \mu_j)$ time using 
 Observation~\ref{obs:rsym_canalyzing}.
@@
 value of $f_n$ is 0 if $n$ is odd and 1 if $n$ is even.  Also, note
-that the above representation is unnormalized. To normalize the
+that the above representation is not default-normalized. 
+To obtain a default-normalized 
 representation, the last line would be changed to have canalyzing
@@
 For any $n \geq 2$, the number of Boolean functions with $n$ variables 
-that are both strongly asymmetric and NCFs is $n! \, 2^{n-1}$.
+that are both strongly asymmetric and NCFs is~ $n! \, 2^{n-1}$.
 \end{theorem}

### 2019-03-29 IEEE_TC_submission.d/preliminaries.tex
ssr

@@
 problems considered in later sections.
-We also show that, in general, the symmetry level
+We also show (Theorem~\ref{thm:approx_sym_level_hard}) 
+that, in general, the symmetry level
 of a Boolean function cannot be efficiently approximated to within any
@@
 where $\pi$ is a permutation of $\{1, 2, \ldots, n\}$.
-%% We will refer to this as the \textbf{lexicographic} ordering
-%% of the variables in the specification.
 Throughout this paper, we will assume without loss of generality
@@
 representation in which lines $n-1$ and $n$ are in the same layer.
-
-%% It can be verified that the number of layers in the 
-%% extended monomial formula for a NCF (as defined in \cite{Li-etal-2013})
-%% is the same as the number of layers in the normalized representation.
-
-\medskip
 We can now define the notion of a \textbf{default-normalized} 

### 2019-06-06 dmtcs_submission.d/intro.tex
ssr: creating the version for DMTCS

--- 2019-03-29 IEEE_TC_submission.d/intro.tex
@@
 
-\IEEEPARstart{R}{search} on the symmetry properties of Boolean
+Rsearch on the symmetry properties of Boolean
 functions has received a lot of attention from the

### 2019-06-06 dmtcs_submission.d/symmetry_properties.tex
ssr: creating the version for DMTCS

--- 2019-03-29 IEEE_TC_submission.d/symmetry_properties.tex
@@
-\section{Symmetry of Nested Canalyzing\newline Functions}
+\section{Symmetry of Nested Canalyzing Functions}
 \label{sec:ncf_and_symmetry}
@@
 
-\subsection{Symmetric Pairs of Variables in an NCF:\newline 
-A Characterization and its Applications}
+\subsection{Symmetric Pairs of Variables in an NCF:
+A Characterization and\newline its Applications}
 \label{sse:ncf_strong_sym}
@@
 
-\subsection{Testing Whether an $r$-Symmetric Function\newline is an NCF}
+\subsection{Testing Whether an $r$-Symmetric Function is an NCF}
 \label{sse:rsym_to_ncf}
@@
 Hence, the number of Boolean functions with $n$ variables that
-are both strongly asymmetric and NCFs is equal to \\ 
+are both strongly asymmetric and NCFs is equal to 
 $2 \times (n!/2) \times 2^{n-1}$ ~=~ $n!\,2^{n-1}$.  \QED

### 2019-06-06 dmtcs_submission.d/abstract.tex
ssr

--- 2019-03-29 IEEE_TC_submission.d/abstract.tex
@@
 has been used to model certain biological phenomena.
-The purpose of this brief contribution is to bring out 
-some interesting relationships between NCFs, symmetric 
+We identify some interesting relationships between NCFs, symmetric 
 Boolean functions and a generalization of symmetric Boolean functions,

### 2019-06-06 dmtcs_submission.d/intro.tex
ssr

@@
 
-Rsearch on the symmetry properties of Boolean
+Research on the symmetry properties of Boolean
 functions has received a lot of attention from the
@@
 \textbf{tested} in line $i$, 
-and $a_i$ and $b_i$ are respectively the canalyzing and 
-canalyzed values in line $i$,~ $1 \leq i \leq n$.
+and $a_i$ and $b_i$ are respectively the \emph{canalyzing} and 
+\emph{canalyzed} values in line $i$,~ $1 \leq i \leq n$.
 Each such line is called a \textbf{rule}.
@@
 as a \textbf{symmetric} function or 
-as a \textbf{totally symmetric function} \cite{Biswas-1970,BS-1968}.
+as a \textbf{totally symmetric function} \citep{Biswas-1970,BS-1968}.
 Thus, in any symmetric function, each pair of variables is symmetric.
@@
 In general, there are functions with $n$ variables that are properly
-$n$-symmetric, but not strongly asymmetric \cite{KS-2000}.
+$n$-symmetric, but not strongly asymmetric \citep{KS-2000}.
 However, in the case of NCFs with $n$ variables, we will show 
@@
 combinational logic functions was first noted 
-by Shannon \cite{Shannon-1938}.
+by \citet{Shannon-1938}.
 Over the years, a considerable amount of research on detecting 
@@
 of Boolean functions has been reported in the literature.
-For example, references \cite{Biswas-1970,TM-1996} present methods 
+For example, \citet{Biswas-1970} and \citet{TM-1996} present methods 
 to determine whether a given Boolean function is symmetric.
-Born and Scidmore \cite{BS-1968} show how certain Boolean functions
+\citet*{BS-1968} show how certain Boolean functions
 can be converted into symmetric functions by introducing additional
@@
 (e.g., \cite{AP-2008,Hu-etal-2008,KS-2000,Darga-etal-2008}).
-Maurer \cite{Maurer-2015} provides a thorough discussion of known
+\citet{Maurer-2015} provides a thorough discussion of known
 symmetry detection algorithms and presents a new universal algorithm 
@@
 The term \textbf{canalization}, coined by
-Waddington \cite{Waddington-1942}, is generally used to describe
+\citet{Waddington-1942}, is generally used to describe
 the stability of a biological system with changes
 in external conditions.
-In 1969, Kauffman \cite{Kauffman-1969} introduced canalyzing Boolean functions
+\citet{Kauffman-1969} introduced canalyzing Boolean functions
 to explain the stability of gene regulatory networks.
 The subclass of NCFs was
-introduced later by Kauffman et al. \cite{Kauffman-etal-2003} 
+introduced later by \citet{Kauffman-etal-2003} 
 to facilitate a rigorous analysis of the Boolean network model
@@
 It is known that the class of NCFs coincides with that
-of unate cascade Boolean functions \cite{Jarrah-etal-2007}.
+of unate cascade Boolean functions \citep{Jarrah-etal-2007}.
 In the literature on computational learning theory,
-NCFs are referred to as $1$-\textbf{decision lists} \cite{KV-1994}.
+NCFs are referred to as $1$-\textbf{decision lists} \citep{KV-1994}.
 Many researchers have pointed out the usefulness of NCFs 
@@
 stability have also been studied in the 
-literature \cite{Kauffman-etal-2004, Layne-2011,Layne-etal-2012,
-Li-etal-2011,Li-etal-2013,Klotz-etal-2013, Stearns-etal-2018}.
+literature (e.g., \cite{Kauffman-etal-2004, Layne-2011,Layne-etal-2012,
+Li-etal-2011,Li-etal-2013,Klotz-etal-2013, Stearns-etal-2018}).
 

### 2019-06-07 dmtcs_submission.d/symmetry_properties.tex
ssr

@@
 in Section~\ref{sec:prelim}.
-In general, there are functions with $n$ variables that are properly
-$n$-symmetric, but not strongly asymmetric \cite{KS-2000}.
+%% In general, there are functions with $n$ variables that are properly
+%% $n$-symmetric, but not strongly asymmetric \cite{KS-2000}.
+For general Boolean functions, the notions of ``$n$-symmetry" and
+``strong asymmetry" are not equivalent. 
+We illustrate this by presenting an example of a Boolean function
+which is properly $n$-symmetric but not strongly asymmetric.
+
+\medskip
+
+\noindent
+\textbf{Example 6:}
+Consider the Boolean function $f$ of four variables, 
+namely $a$, $b$, $c$ and $d$,
+whose truth table is shown in Table~\ref{tab:not_str_ssym_ex}.
+Function $f$ is symmetric under the permutation $cdab$.
+For convenience, we give the truth table for $f$ in a form
+that makes its symmetry under the permutation $cdab$ clear.
+We now argue that the function is
+not pairwise symmetric for any pair of variables.
+
+\medskip
+
+\noindent
+\begin{minipage}{0.05\textwidth}
+\hspace*{0.08in}
+\end{minipage}
+\begin{minipage}{0.85\textwidth}
+\begin{description}
+\item{(i)} $f(0100) \neq f(1000)$, so $a$ and $b$ are not symmetric.
+\item{(ii)} $f(0011) \neq f(1001)$, so $a$ and $c$ are not symmetric.
+\item{(iii)} $f(0001) \neq f(1000)$, so $a$ and $d$ are not symmetric.
+\item{(iv)} $f(0100) \neq f(0010)$, so $b$ and $c$ are not symmetric.
+\item{(v)} $f(0011) \neq f(0110)$, so $b$ and $d$ are not symmetric.
+\item{(vi)} $f(0001) \neq f(0010)$, so $c$ and $d$ are not symmetric.
+\end{description}
+\end{minipage}
+
+%%%% Truth table for the Boolean function in Example 6.
+
+\begin{table}[tb]
+\begin{center}
+\begin{tabular}{|l|l|c||l|l|c|}\hline
+$a~b$ & $c~d$ & $f$ & $a~b$ & $c~d$ & $f$ \\ \hline\hline
+0~0 & 0~1 & 0 & 0~1 & 1~1 & 1 \\ \hline
+0~1 & 0~0 & 0 & 1~1 & 0~1 & 1  \\ \hline
+0~0 & 1~0 & 1 & 1~0 & 1~1 & 1 \\ \hline
+1~0 & 0~0 & 1 & 1~1 & 1~0 & 1 \\ \hline
+0~0 & 1~1 & 0 & 0~0 & 0~0 & 0 \\ \hline
+1~1 & 0~0 & 0 & 0~1 & 0~1 & 0 \\ \hline
+0~1 & 1~0 & 1 & 1~0 & 1~0 & 0 \\ \hline
+1~0 & 0~1 & 1 & 1~1 & 1~1 & 0 \\ \hline
+\end{tabular}
+\end{center}
+\caption{An Example of a Boolean Function which is
+properly $n$-symmetric but not strongly asymmetric.}
+\label{tab:not_str_ssym_ex}
+\end{table}
+
+%%%%% End of truth table.
+
 We now show that any $n$ variable NCF that is properly $n$-symmetric

### 2019-06-07 dmtcs_submission.d/symmetry_properties.tex
ssr

@@
 that makes its symmetry under the permutation $cdab$ clear.
-We now argue that the function is
-not pairwise symmetric for any pair of variables.
+So, the function $f$ is not strongly asymmetric.
+We now demonstrate that the function is properly $4$-symmetric
+by observing that it is not symmetric with respect to any pair of variables.
 
@@
 \noindent
-\begin{minipage}{0.05\textwidth}
-\hspace*{0.08in}
+\begin{minipage}{0.01\textwidth}
+%\hspace*{0.03in}
 \end{minipage}
@@
 \end{center}
-\caption{An Example of a Boolean Function which is
-properly $n$-symmetric but not strongly asymmetric.}
+\caption{An example of a Boolean function with 4 variables which is
+properly $4$-symmetric but not\newline strongly asymmetric.}
 \label{tab:not_str_ssym_ex}

### 2019-09-25 revised_dmtcs.d/intro.tex
DMTCS paper: revised version

--- 2019-06-06 dmtcs_submission.d/intro.tex
@@
 (e.g., \cite{Layne-2011,
-Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}).
+Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}.
 Properties of NCFs such as sensitivity and 
@@
 literature (e.g., \cite{Kauffman-etal-2004, Layne-2011,Layne-etal-2012,
-Li-etal-2011,Li-etal-2013,Klotz-etal-2013, Stearns-etal-2018}).
+Li-etal-2011,Li-etal-2013,Klotz-etal-2013, Stearns-etal-2018,
+Paul-etal-2019,2017b-Kadelka-etal}).
+Generalized versions of nested canalyzing functions where the 
+variables and function values can be from a domain 
+of size three or more have also been studied 
+(e.g., \cite{2017a-Kadelka-etal}).
 

### 2019-09-25 revised_dmtcs.d/preliminaries.tex
DMTCS paper: revised version

--- 2019-03-29 IEEE_TC_submission.d/preliminaries.tex
@@
-\section{Other Definitions and Preliminary\newline Results}
+\section{Other Definitions and Preliminary Results}
 \label{sec:prelim}
@@
 %%\noindent
-Li et al. \cite{Li-etal-2013} have defined the concept of a 
+\citet{Li-etal-2013} defined the concept of a 
 {\bf layer} of an NCF in terms of
@@
 the symmetry level of $f$ to within any factor $\rho \geq 1$.
+This result holds even when the function is given as an expression
+in conjunctive normal form (CNF), that is, in the form of a conjunction
+of clauses where each clause is a disjunction of literals.
 
@@
 the CNF Satisfiability problem (SAT) which is known 
-to be \cnp-hard \cite{GJ-1979}.
+to be \cnp-hard \citep{GJ-1979}.
 

### 2019-09-25 revised_dmtcs.d/symmetry_properties.tex
ssr: changes suggested by reviewer

--- 2019-06-07 dmtcs_submission.d/symmetry_properties.tex
@@
 Thus the function is at most $(r_1 + 2 r_2)$-symmetric.
-Moreover, any pair of variables from different layers, or with different canalyzing values
+Moreover, any pair of variables from different layers, 
+or with different canalyzing values
 are not symmetric, so there are at least $(r_1 + 2 r_2)$ symmetry groups.
@@
 of $q$ layers is $2q$-symmetric, and is not $(q-1)$-symmetric.
-(ii) An $r$-symmetric NCF has a default-normalized nested canalyzing representation
-with at most $r-1$ layers, and has an unnormalized nested canalyzing
-representation with at most $r$ layers.   \QED
+(ii) An $r$-symmetric NCF has a default-normalized 
+nested canalyzing representation with at most $r$ layers. \QED
 \end{corollary}
@@
 are given.
-Let $m_i = |g_i|$, $1 \leq i \leq m$.
+Let $m_i = |g_i|$, $1 \leq i \leq r$.
 Thus, in group $g_i$, the number
@@
 is given by  $\mu = \prod_{i=1}^r (m_i+1)$.
+Since there are $\mu$ rows and each row is an $r$-tuple,
+the size of the table $T$ is $O(r\mu)$.
 As will be seen, our algorithm for determining whether $f$ is
@@
 geometric sum is bounded by $2\mu$.
+Recall that the size of input table $T$ 
+(which specifies the $r$-symmetric function $f$)
+is $O(r\mu)$. 
+Thus, the running time of the algorithm
+is linear in the size of the input.
 The following theorem summarizes this result.
@@
 \begin{theorem}\label{thm:rsym_canalyzing}
-The problem of testing whether a given 
-$r$-symmetric function $f$ is an NCF can 
-be solved efficiently. \QED
+Suppose an $r$-symmetric function $f$ is given as a table where
+each row has an $r$-tuple of the form $(c_1, c_2, \ldots, c_r)$,
+with $c_i$ being the number of variables in group $g_i$ that have value 1,
+along with the \{0,1\} value of the function for that row.
+The problem of testing whether $f$ 
+is an NCF can be solved in time that 
+is linear in the size of the input. \QED
 \end{theorem}
+
+When an $r$-symmetric function $f$ with $n$ variables is
+specified using the table representation described in the
+statement of Theorem~\ref{thm:rsym_canalyzing},
+the size of the table is $O(n^r)$.
+When $r$ is \emph{fixed}, the size of 
+this representation and the running time 
+of the above algorithm are both polynomial functions of $n$.
 
@@
 \begin{tabular}{|l|l|c||l|l|c|}\hline
-$a~b$ & $c~d$ & $f$ & $a~b$ & $c~d$ & $f$ \\ \hline\hline
+$a~b$ & $c~d$ & {Value of $f$} & $a~b$ & $c~d$ & {Value of $f$} \\ \hline\hline
 0~0 & 0~1 & 0 & 0~1 & 1~1 & 1 \\ \hline
@@
 
-\hspace*{0.5in} $x_1 \vee (\overline{x}_2 \wedge ( x_3 \vee (\overline{x}_4 \wedge (\cdots ))$
+\hspace*{0.5in} $x_1 \vee (\overline{x}_2 \wedge 
+    ( x_3 \vee (\overline{x}_4 \wedge (\cdots ))))$
 
@@
 f_6 &~=~  x_1 \vee (\overline{x}_2 \wedge (x_3 \vee (\overline{x}_4 \wedge 
-      (x_5 \vee \overline{x}_6)) \\
+      (x_5 \vee \overline{x}_6))))~~~ \mathrm{and}\\
 f_7 &~=~  x_1 \vee (\overline{x}_2 \wedge (x_3 \vee (\overline{x}_4 \wedge (x_5 
-          \vee (\overline{x}_6 \wedge  x_7)).
+          \vee (\overline{x}_6 \wedge  x_7))))).
 \end{align*}

### 2019-09-25 revised_dmtcs.d/symmetry_properties.tex
ssr: minor changes

@@
 this representation and the running time 
-of the above algorithm are both polynomial functions of $n$.
+of the above algorithm are both polynomials functions of $n$. 
+However, when $r$ is a function of $n$, the size
+of the input and the running time of the algorithm are
+not polynomial functions of $n$.
 

### 2019-09-25 revised_dmtcs.d/response_to_reviwer.tex
ssr: Response to reviewer document

\documentclass[11pt]{article}

\setlength{\textheight}{9.0 in}
\setlength{\textwidth}{6.5 in}
\setlength{\oddsidemargin}{0 in}
\setlength{\evensidemargin}{0 in}
\setlength{\topmargin}{-0.5 in}

\setlength{\parskip}{4pt}

%% \usepackage[nofillcomment,ruled,linesnumbered]{algorithm2e}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}

%% \pagestyle{empty}
\begin{document}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{fact}{Fact}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{observation}{Observation}[section]
\newtheorem{claim}{Claim}[section]

\newcommand{\cnp}{\textbf{NP}}
\newcommand{\true}{\texttt{True}}
\newcommand{\false}{\texttt{False}}

\newcommand{\QED}{\hfill\rule{2mm}{2mm}}

\newcommand{\irange}{\mbox{$1 \leq i \leq n$}}
\newcommand{\jrange}{\mbox{$1 \leq j \leq m$}}

\newcommand{\dunder}[1]{\underline{\underline{#1}}}

\setlength{\parskip}{3pt}

\baselineskip=1.2\normalbaselineskip

\begin{center}
\dunder{\Large{\textbf{Response to the Reviewer}}}
\end{center}

\medskip

\hspace*{0.25in}
\mbox{
\begin{tabular}{ll}
\textsf{Title:} &  Symmetry Properties of Nested Canalyzing Functions \\ [2ex]
\textsf{Authors:} &  D. J. Rosenkrantz, M. V. Marathe, S. S. Ravi and 
                     R. E. Stearns \\ [2ex]
\textsf{Journal:} & Discrete Mathematics and Theoretical Computer Science
\end{tabular}
}

\bigskip\bigskip

\noindent
We thank the reviewer for reading the manuscript carefully and
providing very helpful suggestions.
We have incorporated all the corrections under the section entitled
``Minor" in the review.
Below, we discuss how the two items mentioned under the section
entitled ``Major"  in the review
have been addressed in the revised version.

\begin{enumerate}
\item The statement of Part (ii) of Corollary 3.4 has been corrected. 
It now reads as follows: 
``An $r$-symmetric NCF has a default-normalized
nested canalyzing representation with at most $r$ layers."

\item We have changed the statement of Theorem 3.6 to point out that
the algorithm runs in time that is linear in the \emph{size of the input}.
After the statement of the theorem, we have pointed out that when
$r$ is fixed, the size of the input and the running time of the 
algorithm are both polynomial functions of $n$, 
the number of variables.
If $r$ depends on $n$, then the input size and the running time
of the algorithm are not polynomial functions of $n$.
\end{enumerate}

\end{document}

### 2019-09-25 revised_dmtcs.d/response_to_reviewer.tex
ssr: adding response to reviewer document

--- 2019-09-25 revised_dmtcs.d/response_to_reviwer.tex
@@
 
-\medskip
+\bigskip
 
@@
 \textsf{Title:}   & Symmetry Properties of Nested Canalyzing Functions \\ [2ex]
-\textsf{Authors:} &  D. J. Rosenkrantz, M. V. Marathe, S. S. Ravi and 
-                     R. E. Stearns \\ [2ex]
-\textsf{Journal:} & Discrete Mathematics and Theoretical Computer Science
+\textsf{Authors:} & D. J. Rosenkrantz, M. V. Marathe, S. S. Ravi 
+                    and R. E. Stearns \\ [2ex]
+\textsf{Journal:} & Discrete Mathematics and Theoretical Computer Science \\
 \end{tabular}
@@
 
-\noindent
 We thank the reviewer for reading the manuscript carefully and
-providing very helpful suggestions.
-We have incorporated all the corrections under the section entitled
-``Minor" in the review.
-Below, we discuss how the two items mentioned under the section
-entitled ``Major"  in the review
-have been addressed in the revised version.
+providing very helpful suggestions. We have incorporated all the
+corrections indicated in the section entitled ``Minor" in the review. 
+Below, we mention how the two items mentioned in the 
+``Major" section of the review have been 
+addressed in the revised version.
+
+\medskip
 
 \begin{enumerate}
-\item The statement of Part (ii) of Corollary 3.4 has been corrected. 
-It now reads as follows: 
-``An $r$-symmetric NCF has a default-normalized
+\item
+The statement of Part (ii) of Corollary 3.4 has been corrected. It
+now reads as follows: ``An $r$-symmetric NCF has a default-normalized
 nested canalyzing representation with at most $r$ layers."
 
-\item We have changed the statement of Theorem 3.6 to point out that
-the algorithm runs in time that is linear in the \emph{size of the input}.
-After the statement of the theorem, we have pointed out that when
-$r$ is fixed, the size of the input and the running time of the 
-algorithm are both polynomial functions of $n$, 
-the number of variables.
-If $r$ depends on $n$, then the input size and the running time
-of the algorithm are not polynomial functions of $n$.
+\item We have changed the statement of Theorem 3.6 to point out
+that the algorithm runs in time that is linear in the \emph{size of the
+input.} 
+After the statement of the theorem, we have pointed out that
+when $r$ is fixed, the size of the input and the running time of the
+algorithm are both polynomial functions of $n$, the number of variables.
+If $r$ depends on $n$, then the input size and the running time of the
+algorithm are not polynomial functions of $n$.
 \end{enumerate}

### 2019-09-26 revised_dmtcs.d/response_to_reviewer.tex
ssr: minor change to response document

@@
 input.} 
-After the statement of the theorem, we have pointed out that
+After the statement of the theorem, we have indicated that
 when $r$ is fixed, the size of the input and the running time of the
 algorithm are both polynomial functions of $n$, the number of variables.
-If $r$ depends on $n$, then the input size and the running time of the
-algorithm are not polynomial functions of $n$.
+If $r$ is not fixed, then the input size and the running time of the
+algorithm are not necessarily polynomial functions of $n$; however,
+the running time of the algorithm remains linear in the size of the input.
 \end{enumerate}

### 2019-09-27 revised_dmtcs.d/symmetry_properties.tex
ssr: minor changes

@@
 When an $r$-symmetric function $f$ with $n$ variables is
-specified using the table representation described in the
+specified by giving the table representation described in the
 statement of Theorem~\ref{thm:rsym_canalyzing},
-the size of the table is $O(n^r)$.
+the size of the given table is $O(n^r)$.
 When $r$ is \emph{fixed}, the size of 
@@
 of the above algorithm are both polynomials functions of $n$. 
-However, when $r$ is a function of $n$, the size
+When $r$ is not fixed, while the size
 of the input and the running time of the algorithm are
-not polynomial functions of $n$.
+not necessarily polynomial functions of $n$, 
+the running time remains linear in the input size.
 
@@
 
+\medskip
+
+\noindent
 We now show that any $n$ variable NCF that is properly $n$-symmetric

### 2019-09-27 revised_dmtcs.d/intro.tex
ssr: minor changes

@@
 class of functions, called \textbf{canalyzing} Boolean functions.
-This class, introduced in \cite{Kauffman-1969}, is defined as follows.
+This class, introduced by \cite{Kauffman-1969}, is defined as follows.
 
@@
 Another class of Boolean functions, called \textbf{nested canalyzing functions} (NCFs),
-was introduced later in \cite{Kauffman-etal-2003} to carry out a detailed
+was introduced later by \cite{Kauffman-etal-2003} to carry out a detailed
 analysis of the behavior of certain biological systems.
@@
 $f(a_1, a_2, \ldots, a_n)$ = $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
-For $r \geq 1$, the class of $r$-symmetric functions has been
+For fixed $r \geq 1$, the class of $r$-symmetric functions has been
 studied in the literature on discrete dynamical systems (see e.g., 
@@
 $f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
-In general, there are functions with $n$ variables that are properly
+In general, there are Boolean functions with $n$ variables that are properly
 $n$-symmetric, but not strongly asymmetric \citep{KS-2000}.
@@
 unless \textbf{P} = \cnp.)
-We also present an efficient algorithm to test whether a given
+We also present an algorithm to test whether a given
 $r$-symmetric function $f$ is an NCF. 
@@
 (e.g., \cite{Layne-2011,
-Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}.
+Layne-etal-2012,Li-etal-2011,Li-etal-2012,Li-etal-2013}).
 Properties of NCFs such as sensitivity and 

### 2019-09-27 revised_dmtcs.d/preliminaries.tex
ssr: minor changes

@@
 
+\noindent
 In view of Observation~\ref{obs:normalization_poly},
-we will assume from now on that any NCF is specified in default-normalized form.
+we assume henceforth that any NCF is specified in default-normalized form.
 
@@
 \begin{enumerate}
-\item For the variables in $X$, choose an assignment that
+\item For the variables in $X$, choose any assignment that
 satisfies $g$.

### 2019-09-27 revised_dmtcs.d/symmetry_properties.tex
ssr: minor changes

@@
 its default-normalized representation.
-We also present an efficient algorithm (Theorem~\ref{thm:rsym_canalyzing})
+We also present an algorithm (Theorem~\ref{thm:rsym_canalyzing})
 for the converse problem, that is,
@@
 \item For $j < i < k$, and for  $k < i \leq n$, 
-if $b_i = b_j$,\\ then set $c_i =  \overline{a_i}$,
+if $b_i = b_j$,~ then set $c_i =  \overline{a_i}$,
 else set $c_i =  a_i$.

### 2019-09-27 revised_dmtcs.d/symmetry_properties.tex
ssr: final set of minor changes

@@
 $f$ is $r$-symmetric can be solved efficiently.
-We now show that the converse problem, that is, testing whether a given
-$r$-symmetric Boolean function is an NCF,
-can also be solved efficiently.
-When the answer is ``yes", a
-default-normalized representation (defined in Section~\ref{sec:prelim})
-of $f$ can also be constructed efficiently.
+We now consider the converse problem, that is, testing whether a given
+$r$-symmetric Boolean function is an NCF.
+For this problem, we present an algorithm whose running time
+is linear in the size of the representation 
+of the $r$-symmetric function $f$.
+When the answer is ``yes", the algorithm also constructs a
+default-normalized representation 
+(defined in Section~\ref{sec:prelim}) of $f$.
 
@@
 Thus, the time for iteration $j$ is $O(r \mu_j)$.
-We note that in each iteration, the number of rows in the table
+We claim that in each iteration, the number of rows in the table
 is reduced by a factor of at least 2; that is, $\mu_{j+1} \leq \mu_j/2$.
@@
 \textbf{Part (i):}~
-Suppose symmetric function $f$ is also a canalyzing function. 
+Suppose that symmetric function $f$ is also a canalyzing function. 
 Then there is a variable $x$, and values $a$ and $b$ 
